### 2021-08-10 paper/pmih_paper.Rmd
recoding changes

---
title: "Paper summary: police demand and persons with mental ill-health"
output: html_document
tags:
- demand
- reactive demand
- mental health
- calls for service
- text mining
date: '2021-28-04'
---

## Police demand

There is a common misconception that police spend most of their time responding to and dealing with crime. Instead, the evidence suggests that public demand for police services originates from a diverse array of incidents -- many of which involve vulnerable people and require specific training and numerous

Estimating the scale of demand of police services is not a simple task, particularly when it comes to complex sources of demand, such as those originating from persons with mental ill-health. One common method has been to use emergency calls for service data. For instance, we can say _how many_ emergency calls were classified as involving mental ill-health in any given time period. We can take this one step further by calculating the _amount of time_ police spend dealing with such incidents. We can then make claims such as: "Police officers spend x% of their time responding to and resolving incidents involving mental ill-health".

Gathering such evidence -- which simply _describes_ the composition of public demand for the police -- is important for a number of reasons. Among them, it can help guide training to ensure that officers are suitably prepared for their duties, maximising well-being outcomes for the public and mitigating against unnecessary strain on officers. Understanding police demand can also contribute to how funds from central and local governments are decided and distributed to police forces and other public services. This has become particularly pertinent following recent 'defund the police' campaigns. With police performing a variety of non-crime-related duties, any changes to their role will likely have a knock-on effect to other sectors (and vice-versa).

The question which follows is then: how can we measure police demand? Are we measuring it reliabily

## Mental health

In a recent paper published in _Policing: A Journal of Policy and Practice_, we focused on describing

Police contact with persons experiencing mental ill-health might be the most salient example of these challenges.

 Inevitably, this is not straightforward.

## Mental health

Since my article in The Conversation, I have become increasingly intrigued by the complexities of understanding and measuring police demand. This has largely driven by some of the amazing research that has been published in recent years (see readings below).

In many countries, one such demand type is particularly concerning; those originating from incidents involving people with mental ill-health. In England and Wales, forces have reported that around 2% of incidents relate to mental health, but this is widely understood to be an **underestimate** due to inconsistencies in how incidents are flagged. Not only this, but the proportion of emergency calls for service does not necessarily reflect the **time** police actually consume dealing with such incidents.

These two issues represented a key motivation for the study. We sought to (1) provide an improved, data-driven estimate of the proportion of calls for service involving mental ill-health, and (2) estimate the amount of time officers spend resolving such incidents.

### 2021-08-10 scripts/calls_detroit.R

@@ -20,171 +20,114 @@ library(tidyr)
 # Load.
 detroit_df <- read_csv("data/detroit_16_21.csv")
 
-# Date checks.
-detroit_df %>% 
-  slice(1:5) %>% 
-  select(call_timestamp)
-
 # Subset for 2019. We want pre-COVID.
 detroit19_df <- detroit_df %>% 
   filter(str_detect(call_timestamp, "2019/"))
 
 # Date checks.
 detroit19_df %>% 
-  slice(1:5) %>% 
-  select(call_timestamp)
-
-detroit19_df %>% 
   mutate(year = as.character(str_extract_all(call_timestamp, "^.{4}"))) %>% 
   group_by(year) %>% 
   summarise(counts = n())
 
-# Remove original to free memory if needed.
-rm(detroit_df)
+# Check time variables variables.
+detroit19_df %>% 
+  select(calldescription, dispatchtime, traveltime, totalresponsetime, time_on_scene, totaltime)
 
-# Missings.
+# Totaltime = totalresponsetime + time_on_scene
+# Makes sense.
+
+# Missings in key variables.
 sum(is.na(detroit19_df$calldescription))
-sum(is.na(detroit19_df$totaltime))
+sum(is.na(detroit19_df$totaltime)) # ~121k
 
 # Remove missings, zeros and negatives.
 detroit19_sub_df <- detroit19_df %>% 
   drop_na(totaltime) %>% 
-  filter(totaltime != 0, totaltime >= 0)
+  filter(totaltime >0)
 
-# What's the distribution now?
-# ggplot(data = detroit19_sub_df) +
-#   geom_histogram(mapping = aes(x = totaltime), bins = 100)
+# What is the distribution of total time?
+ggplot(data = detroit19_sub_df) +
+  geom_histogram(mapping = aes(x = totaltime), bins = 30)
 
-# Outliers are massively skewing this. For now, arbitrary cut-off.
-detroit19_sub_no_df <- detroit19_sub_df %>% 
-  filter(totaltime < 300)
-
-# What's the distribution now?
-# ggplot(data = detroit19_sub_no_df) +
-#   geom_histogram(mapping = aes(x = totaltime), bins = 100)
-
-# What take the most time?
-freqs_df <- detroit19_sub_no_df %>% 
-  group_by(calldescription) %>% 
-  summarise(tt_cd = sum(totaltime)) %>% 
-  arrange(tt_cd)
-
-# Calculate sum times by category, filter out admin/role categories.
[…]

### 2021-08-20 scripts/calls_detroit.R
visual tweaks

@@ -84,7 +84,7 @@ detroit19_deploy_df <- detroit19_deploy_df %>%
          calldescription2 = if_else(str_detect(calldescription, "AUTO X")                                  ,"AUTO ACCIDENT"              , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "ASSAULT AND BATTERY")                     ,"ASSAULT & BATTERY"          , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "ASSAULT DANGEROUSOR SERIOUS|ASSAULT OR")  ,"OTHER ASSAULT"              , calldescription2),
-         calldescription2 = if_else(str_detect(calldescription,"FELONIOUS ASSAULT")                        ,"FELONIOUS ASSAULT"          , calldescription2),
+         calldescription2 = if_else(str_detect(calldescription, "FELONIOUS ASSAULT")                       ,"FELONIOUS ASSAULT"          , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "OVERDOSE|DRUG OD")                        ,"OVERDOSE"                   , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "ROBBERY ARMED")                           ,"ARMED ROBBERY"              , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "ROBBERY NOT ARMED")                       ,"UNARMED ROBBERY"            , calldescription2),
@@ -98,7 +98,7 @@ detroit19_deploy_df <- detroit19_deploy_df %>%
          calldescription2 = if_else(str_detect(calldescription, "NARCOTICS")                               ,"NARCOTICS"                  , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "LARCENY")                                 ,"LARCENY"                    , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "RAPE")                                    ,"RAPE"                       , calldescription2),
-         calldescription2 = if_else(str_detect(calldescription, "LEWD AND LASCIVIOUS")                     ,"LEWD & LASCIVIOUS CONDUCT"  , calldescription2),
[…]

### 2021-08-22 scripts/calls_detroit.R
descriptive tables

@@ -151,6 +151,12 @@ calldesc2_df <- detroit19_deploy_df %>%
   distinct() %>% 
   arrange(calldescription)
 
+# Save for reader reference.
+calldesc2_df %>% 
+  rename(calldescription_original = calldescription,
+         calldescription_new      = calldescription2) %>% 
+write_csv(file = "results/categorisation_summary.csv")
+
 # Aggregate.
 detroit_19_times_df <- detroit19_deploy_df %>% 
   group_by(calldescription2) %>% 
@@ -290,11 +296,12 @@ detroit_19_times_df <- detroit_19_times_df %>%
 # Check missings.
 sum(is.na(detroit_19_times_df)) # 0
 
-# Create 'other' categories, aggregate.
+# Create 'other' categories for incident <0.2% deployed time, then aggregate.
 detroit_19_times_agg_df <- detroit_19_times_df %>%
   mutate(calldescription2 = if_else(prop_time < 0.2, true  = "OTHER", false = calldescription2)) %>% 
   group_by(calldescription2, type) %>% 
-  summarise(prop_time  = sum(prop_time),
+  summarise(freq       = sum(sum_counts),
+            prop_time  = sum(prop_time),
             prop_count = sum(prop_counts)) %>% 
   ungroup() %>% 
   arrange(type)
@@ -303,14 +310,22 @@ detroit_19_times_agg_df <- detroit_19_times_df %>%
 sum(is.na(detroit_19_times_agg_df)) # 0
 
 # Descriptive stats. 
-detroit_19_times_agg_df %>% 
+des_stats_df <- detroit_19_times_agg_df %>% 
   group_by(type) %>% 
-  summarise(type_prop_count  = sum(prop_count),
-            type_prop_time   = sum(prop_time)) %>% 
+  summarise(`Frequency`       = sum(freq),
+            `Percentage`      = round(sum(prop_count), 2),
+            `Percentage time` = round(sum(prop_time ), 2)) %>%
+  rename(`Demand category` = type) %>% 
   ungroup()
+
+# Save.
+write_csv(x = des_stats_df, file = "results/des_stats.csv")
 
 # Create categorical colour scheme for future use. Colourblind friendly.
 # col_vec <- c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")
+
+# Save data used for proportional breakdown.
[…]

### 2021-08-22 paper.Rmd
paper start

--- 2021-08-10 paper/pmih_paper.Rmd
@@
 ---
-title: "Paper summary: police demand and persons with mental ill-health"
-output: html_document
-tags:
-- demand
-- reactive demand
-- mental health
-- calls for service
-- text mining
-date: '2021-28-04'
+title: "Describing the scale and composition of police demand in Detroit."
+output: pdf_document
+header-includes:
+  - \usepackage{lscape}
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = F, message=F, comment=F, warning=F)
+```
 
@@
 
-The question which follows is then: how can we measure police demand? Are we measuring it reliabily
+\newpage
 
-## Mental health
+\begin{landscape}
 
-In a recent paper published in _Policing: A Journal of Policy and Practice_, we focused on describing
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category."}
+knitr::include_graphics("visuals/fig1_time_ann.png")
+```
 
-Police contact with persons experiencing mental ill-health might be the most salient example of these challenges.
+\end{landscape}
 
- Inevitably, this is not straightforward.
+\newpage
 
-## Mental health
+```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and week by demand type.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig2_time_heat.png")
+```
 
-Since my article in The Conversation, I have become increasingly intrigued by the complexities of understanding and measuring police demand.
-This has largely driven by some of the amazing research that has been published in recent years (see readings below).
-
-In many countries, one such demand type is particularly concerning; those originating from incidents involving people with mental ill-health.
-In England and Wales, forces have reported that around 2% of incidents relate to mental health, but this is widely understood to be an **underestimate** due to inconsistencies in how incidents are flagged.
-Not only this, but the proportion of emergency calls for service does not necessarily reflect the **time** police actually consume dealing with such incidents.
-
-These two issues represented a key motivation for the study.
-We sought to (1) provide an improved, data-driven estimate of the proportion of calls for service involving mental ill-health, and (2) estimate the amount of time officers spend resolving such incidents.
+```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts by demand type.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig3_maps.png")
+```

### 2021-08-22 scripts/calls_detroit.R

@@ -352,7 +352,7 @@ time_ann_gg <- time_gg +
   annotate(geom = "text", label = "community", family = "sans",
            size = 16, x = 0.93, y = 1.08) +
   annotate(geom = "text", label = "quality of life", family = "sans",
-           size = 16, x = 1.07, y = 0.57, angle = 90) 
+           size = 16, x = 1.07, y = 0.57, angle = -90) 
 
 # Save.
 ggsave(filename = "visuals/fig1_time_ann.png", height = 42, width = 62, unit = "cm", dpi = 300)

### 2021-08-22 scripts/phoenix.R

@@ -1,47 +1,0 @@
-
-# # Download open calls-for-service data for 2020 in Phoenix.
-# download.file(url = "https://www.phoenixopendata.com/dataset/64a60154-3b2d-4583-8fb5-6d5e1b469c28/resource/3c0ae3ec-456f-45f4-801d-b8d6699ba32e/download/callsforsrvc2020.csv",
-#               destfile = "data/cfs_phoenix_2021.csv")
-
-# Load.
-cfs_df <- read_csv("data/cfs_phoenix_2021.csv", col_types = cols(INCIDENT_NUM = col_character()))
-# cri_df <- read_csv("data/crime_phoenix.csv", col_types = cols(`INC NUMBER` = col_character()))
-
-# Frequencies. Pull out those that are 'infrequent' (arbitary <= 20 counts per year).
-infreqs <- cfs_df %>% 
-  group_by(FINAL_CALL_TYPE) %>% 
-  summarise(counts = n()) %>% 
-  ungroup() %>% 
-  arrange(counts) %>% # visual check
-  filter(counts <= 100) %>% 
-  pull(FINAL_CALL_TYPE)
-
-# Remove these call types from the individual records.
-# cfs_sub_df <- cfs_df %>% 
-#   filter(FINAL_CALL_TYPE %nin% infreqs)
-
-# Code these call types as 'other'
-cfs_sub_df <- cfs_df %>%
-  mutate(FINAL_CALL_TYPE2 = if_else(FINAL_CALL_TYPE %in% infreqs,
-                                    "OTHER",
-                                    FINAL_CALL_TYPE))
-
-
-# Now we are dealing with slightly less categories.
-length(unique(cfs_sub_df$FINAL_CALL_TYPE2))
-
-# Proportion frequencies.
-freqs_df <- cfs_sub_df %>% 
-  drop_na(FINAL_CALL_TYPE2) %>% 
-  filter()
-  group_by(FINAL_CALL_TYPE2, DISP_CODE) %>% 
-  summarise(counts = n()) %>% 
-  mutate(total = nrow(cfs_sub_df),
-         props = 100*(counts/total)) %>% 
-  ungroup()
-
-# Visualize it.
-ggplot(data = freqs_df,
-       mapping = aes(area = props, label = FINAL_CALL_TYPE2, group = DISP_CODE)) +
-  geom_treemap(fill = "white", colour = "black", size = 2) +
-  geom_treemap_text()

### 2021-08-23 paper.Rmd
prop graphic labels

@@
 
-\begin{landscape}
+<!-- \begin{landscape} -->
 
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category."}
-knitr::include_graphics("visuals/fig1_time_ann.png")
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig1_time_port.png")
 ```
 
-\end{landscape}
+<!-- \end{landscape} -->
 
@@
 
-```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and week by demand type.", out.width = "\\textwidth"}
+```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and week, for each demand type.", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig2_time_heat.png")
@@
 
-```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts by demand type.", out.width = "\\textwidth"}
+```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts for each demand type.", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig3_maps.png")

### 2021-08-23 scripts/calls_detroit.R

@@ -79,10 +79,10 @@ detroit19_deploy_df <- detroit19_deploy_df %>%
   mutate(calldescription2 = if_else(str_detect(calldescription, "DV")                                      ,"DOMESTIC INCIDENT"          , calldescription),
          calldescription2 = if_else(str_detect(calldescription, "MENTAL")                                  ,"MENTAL HEALTH"              , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "SUICIDE")                                 ,"SUICIDE OR SUICIDE THREAT"  , calldescription2),
-         calldescription2 = if_else(str_detect(calldescription, "HIT & RUN|HIT& RUN|H&R")                  ,"HIT & RUN"                  , calldescription2),
+         calldescription2 = if_else(str_detect(calldescription, "HIT & RUN|HIT& RUN|H&R")                  ,"HIT AND RUN"                  , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "LOST PROPERTY|RECOVERED / FOUND PROPERTY"),"LOST PROPERTY"              , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "AUTO X")                                  ,"AUTO ACCIDENT"              , calldescription2),
-         calldescription2 = if_else(str_detect(calldescription, "ASSAULT AND BATTERY")                     ,"ASSAULT & BATTERY"          , calldescription2),
+         calldescription2 = if_else(str_detect(calldescription, "ASSAULT AND BATTERY")                     ,"ASSAULT AND BATTERY"          , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "ASSAULT DANGEROUSOR SERIOUS|ASSAULT OR")  ,"OTHER ASSAULT"              , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "FELONIOUS ASSAULT")                       ,"FELONIOUS ASSAULT"          , calldescription2),
          calldescription2 = if_else(str_detect(calldescription, "OVERDOSE|DRUG OD")                        ,"OVERDOSE"                   , calldescription2),
@@ -98,10 +98,10 @@ detroit19_deploy_df <- detroit19_deploy_df %>%
          calldescription2 = if_else(str_detect(calldescription, "NARCOTICS")                               ,"NARCOTICS"                  , calldescription2),
[…]

### 2021-08-23 paper.Rmd
fonts

--- 2021-08-22 paper.Rmd
@@
 title: "Describing the scale and composition of police demand in Detroit."
+author: "Samuel Langton (Netherlands Institute for the Study of Crime and Law Enforcement)"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message=F, comment=F, warning=F)
+knitr::opts_chunk$set(echo = F, message=F, comment=F, warning=F, fig.pos = 'H')
+library(readr)
+library(dplyr)
+library(kableExtra)
 ```
 
-## Police demand
+## Introduction
 
@@
 Instead, the evidence suggests that public demand for police services originates from a diverse array of incidents -- many of which involve vulnerable people and require specific training and numerous
+
+## Police demand
 
@@
 \newpage
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig1_time_port.png")
+```
+\newpage
 
-\begin{landscape}
+```{r table1}
+table1_df <- read_csv("results/table1_des_stats.csv")
 
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category."}
-knitr::include_graphics("visuals/fig1_time_ann.png")
+kable(table1_df, caption = "Frequency and proportional breakdown (counts and time) for each demand type.",
+      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
+    kable_styling(latex_options = c("striped"),
+                  font_size = 10, full_width = F)
 ```
 
-\end{landscape}
+## Temporal patterning
 
 \newpage
-
-```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and week by demand type.", out.width = "\\textwidth"}
+```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and week, for each demand type.", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig2_time_heat.png")
 ```
+\newpage
 
-```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts by demand type.", out.width = "\\textwidth"}
+## Spatial patterning
+
+\newpage
+```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts for each demand type.", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig3_maps.png")
 ```
+\newpage
+
+## Conclusion

### 2021-08-23 scripts/calls_detroit.R

@@ -311,15 +311,16 @@ sum(is.na(detroit_19_times_agg_df)) # 0
 
 # Descriptive stats. 
 des_stats_df <- detroit_19_times_agg_df %>% 
+  # mutate(type = str_to_title(type)) %>% 
   group_by(type) %>% 
-  summarise(`Frequency`       = sum(freq),
-            `Percentage`      = round(sum(prop_count), 2),
-            `Percentage time` = round(sum(prop_time ), 2)) %>%
-  rename(`Demand category` = type) %>% 
+  summarise(`Count`      = sum(freq),
+            `Count (%)`  = round(sum(prop_count), 2),
+            `Time (%)`   = round(sum(prop_time ), 2)) %>%
+  rename(`Demand type` = type) %>% 
   ungroup()
 
 # Save.
-write_csv(x = des_stats_df, file = "results/des_stats.csv")
+write_csv(x = des_stats_df, file = "results/table1_des_stats.csv")
 
 # Create categorical colour scheme for future use. Colourblind friendly.
 # col_vec <- c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")
@@ -336,7 +337,7 @@ time_gg <- detroit_19_times_agg_df %>%
   filter(type != "unclassified") %>%
   ggplot(mapping = aes(area = prop_time, label = calldescription2_labs, subgroup = type)) +
   geom_treemap(fill = "snow", colour = "darkgrey", size = 2, alpha = 0.5) +
-  geom_treemap_text(padding.y = unit(0.3, "cm"), grow = FALSE, family = "sans") +
+  geom_treemap_text(padding.y = unit(0.3, "cm"), grow = FALSE) +
   geom_treemap_subgroup_border(colour = "dodgerblue3", size = 4) +
   theme_void() +
   theme(panel.border  = element_rect(colour = "dodgerblue3", fill = "transparent", size = 1.5))
@@ -345,17 +346,17 @@ time_gg <- detroit_19_times_agg_df %>%
 time_ann_gg <- time_gg +
   theme(plot.margin = unit(c(3,3,3,3), "cm")) +
   coord_cartesian(xlim = c(0,1), ylim = c(0,1), clip = "off") +
-  annotate(geom = "text", label = "crime"    , family = "sans",
+  annotate(geom = "text", label = "crime"    ,
            size = 9, x = 0.27, y = -0.065) +
[…]

### 2021-08-24 paper.Rmd
officer initiated change, writing

@@
 ---
-title: "Describing the scale and composition of police demand in Detroit."
+title: "Describing the scale and composition of calls for service in Detroit"
 author: "Samuel Langton (Netherlands Institute for the Study of Crime and Law Enforcement)"
-output: pdf_document
-header-includes:
-  - \usepackage{lscape}
+bibliography: refs.bib
+output:
+  # word_document: default
+  pdf_document: default
+header-includes: \usepackage{lscape}
+csl: apa.csl
+abstract: In the spirit of a recent contribution by Ratcliffe (2021) published in
+  _Crime Science_ this short paper describes the scale and composition of public demand
+  for police services in Detroit, United States, during 2019. Findings broadly confirm
+  those by Ratcliffe, namely, that a considerable proportion of police time is spent
+  resolving a diverse array of (non-criminal) calls for service, including public
+  health incidents. The temporal and spatial patterning of different types of public
+  demand for police services are often distinct. The data and code to replicate analyses
+  are made openly available.
 ---
+
+Keywords: police, calls for service, 911, demand, reactive.
 
@@
 
-There is a common misconception that police spend most of their time responding to and dealing with crime.
-Instead, the evidence suggests that public demand for police services originates from a diverse array of incidents -- many of which involve vulnerable people and require specific training and numerous
+Amidst austerity measures, growing public expectation, and scrutiny, understanding the public demand for police services has become a priority among evidence-based policing researchers and practitioners [@boulton2017calls].
+Without a grasp on the scale ('how much?') and composition ('what type?') of police demand, we are likely to observe sub-optimal and inequitable outcomes for the public, the inappropriate distribution of public funds, and unnecessary strain on officers [@ellison2021understanding; @lum2021can].
+Understanding the characteristics public demand for the police has become particularly pertinent following recent calls to rethink, and some cases, radically reform, the role and reach of contemporary police forces [@lum2021can].
 
-## Police demand
+A study from @ratcliffe2021policing in _Crime Science_ sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States.
+The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside the traditional crime-fighting role of the police.
+Instead, the origins of public demand for the police can often be attributed to a lack of supply and/or accessibility failure in other organisations, such as those providing support for people requiring (mental) health assistance [@wood2021can; @van2017law].
+The study's findings provided data-driven evidence to suggest that the police indeed spend considerable amounts of time dealing with incidents that could be resolved by other -- often more appropriate -- agencies.
+Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can].
+In advance of recent contributions, this short paper aims to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
 
-Estimating the scale of demand of police services is not a simple task, particularly when it comes to complex sources of demand, such as those originating from persons with mental ill-health.
-One common method has been to use emergency calls for service data.
-For instance, we can say _how many_ emergency calls were classified as involving mental ill-health in any given time period.
-We can take this one step further by calculating the _amount of time_ police spend dealing with such incidents.
-We can then make claims such as: "Police officers spend x% of their time responding to and resolving incidents involving mental ill-health".
+## Measuring demand
 
-Gathering such evidence -- which simply _describes_ the composition of public demand for the police -- is important for a number of reasons.
-Among them, it can help guide training to ensure that officers are suitably prepared for their duties, maximising well-being outcomes for the public and mitigating against unnecessary strain on officers.
-Understanding police demand can also contribute to how funds from central and local governments are decided and distributed to police forces and other public services.
-This has become particularly pertinent following recent 'defund the police' campaigns.
-With police performing a variety of non-crime-related duties, any changes to their role will likely have a knock-on effect to other sectors (and vice-versa).
+Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for service in a typical year, defined as 2019 (pre-COVID-19).
+Incident descriptions (e.g., investigate person) were classified into general categories, namely, community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+Incidents were subset for those originating from the public, and thus not initiated by an officer.
+Recognising that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by each incident, raw counts were supplemented by the time consumed by police officers dealing with each incident type.
+Findings indicated that around 55% of officer shift activity time was spent on crime incidents, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%).
+Focusing on medical/public health incidents, findings indicated that calls concentrated in particular areas of the city and during particular hours of the day.
+
+The data used for this study covers the city of Detroit in Michigan, United States.
+The City of Detroit publish comparable calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/].
+The raw data includes both citizen-initiated 911 calls to request police services and officer-initiated calls spanning back to September 2016.
+In alignment with @ratcliffe2021policing, the data used here is subset for the year 2019 and excludes those calls initiated by an officer.
+Each incident has supplementary information which describes the type of incident, the latitude and longitude coordinates, and a timestamp.
+For each incident, the response time and time on the scene are reported.
+In sum, these comprise the total time officers spent on the incident.
 
 \newpage
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category.", out.width = "\\textwidth"}
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category. Graphic inspired by Ratcliffe (2012).", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig1_time_port.png")
@@
 
-## Conclusion
+## Discussion

### 2021-08-24 scripts/calls_detroit.R

@@ -12,7 +12,6 @@ library(treemapify)
 library(ggplot2)
 library(sf)
 
-
 # Useful function.
 `%nin%` <- Negate(`%in%`)
 
@@ -26,9 +25,12 @@ library(sf)
 # Load.
 detroit_df <- read_csv("data/detroit_16_21.csv")
 
-# Subset for 2019. We want pre-COVID.
+# Initial filter.
+sum(is.na(detroit_df$officerinitiated)) # 0
+
 detroit_df <- detroit_df %>% 
-  filter(str_detect(call_timestamp, "2019"))
+  filter(str_detect(call_timestamp, "2019"), # 2019 only.
+         officerinitiated == "No")           # exclude officer initiated calls.
 
 # Date checks.
 detroit_df %>% 
@@ -37,16 +39,13 @@ detroit_df %>%
   summarise(counts = n()) %>% 
   ungroup()
 
-# Check time variables variables.
+# Check time variables variables. Totaltime = totalresponsetime + time_on_scene. Makes sense.
 detroit_df %>% 
   select(calldescription, dispatchtime, traveltime, totalresponsetime, time_on_scene, totaltime)
 
-# Totaltime = totalresponsetime + time_on_scene
-# Makes sense.
-
 # Missings in key variables.
 sum(is.na(detroit_df$calldescription))
-sum(is.na(detroit_df$totaltime)) # ~121k
+sum(is.na(detroit_df$totaltime)) 
 
 # Remove missings, zeros and negatives.
 detroit19_sub_df <- detroit_df %>% 
@@ -57,18 +56,18 @@ detroit19_sub_df <- detroit_df %>%
 ggplot(data = detroit19_sub_df) +
   geom_histogram(mapping = aes(x = totaltime), bins = 30)
 
-# First, filter out calldescriptions which are admin, completely unknown or do not involve deployment.
-# We retain 'REPORT' (or similar e.g. RPT).
+# First, filter out calldescriptions which are admin, completely unknown or do not appear to
+# involve deployment.
 detroit19_deploy_df <- detroit19_sub_df %>%
   filter(calldescription != "REMARKS", 
          calldescription != "START OF SHIFT INFORMATION",
          calldescription != "UNKNOWN PROBLEM",
-         calldescription != "VIRTUAL SPECIAL ATTENTION",
+         calldescription != "VIRTUAL SPECIAL ATTENTION", 
          calldescription != "EMPLOYEE CALL IN / TIME OFF",
[…]

### 2021-08-24 paper.Rmd
writing

@@
 
-Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for service in a typical year, defined as 2019 (pre-COVID-19).
-Incident descriptions (e.g., investigate person) were classified into general categories, namely, community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
-Incidents were subset for those originating from the public, and thus not initiated by an officer.
-Recognising that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by each incident, raw counts were supplemented by the time consumed by police officers dealing with each incident type.
+Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for service in 2019.
+CAD codes, used the describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were classified into six broad categories: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type.
 Findings indicated that around 55% of officer shift activity time was spent on crime incidents, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%).
-Focusing on medical/public health incidents, findings indicated that calls concentrated in particular areas of the city and during particular hours of the day.
+Focusing on medical/public health incidents, findings also indicated that calls concentrated in particular areas of the city and during particular hours of the day.
 
 The data used for this study covers the city of Detroit in Michigan, United States.
-The City of Detroit publish comparable calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/].
+The City of Detroit authority publish calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/].
 The raw data includes both citizen-initiated 911 calls to request police services and officer-initiated calls spanning back to September 2016.
 In alignment with @ratcliffe2021policing, the data used here is subset for the year 2019 and excludes those calls initiated by an officer.
-Each incident has supplementary information which describes the type of incident, the latitude and longitude coordinates, and a timestamp.
 For each incident, the response time and time on the scene are reported.
 In sum, these comprise the total time officers spent on the incident.
+Incidents which had a total allocated time of zero minutes (or negative values) were excluded, leaving 301,490 calls for service throughout the year.
+Each incident has a _calldescription_ variable which describes the nature of the call.
+There were 207 unique call description categories which in the interests of parsimony and ease of interpretation were recoded into 101 categories.
+^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (xxxxxxxxxxxxxxxxxxxxxxxxxx).
+Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.
+Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing].
+For each of the six demand types, an 'other' classification was generated for those incident types were consumed less than 0.2% of police time during the year.
+This left a total of 48 categories for the reported breakdown.
+
+Each incident also has corresponding latitude and longitude coordinates and a call timestamp.
+Due to irregularities in geocoding, likely due unknown locations for some calls, 2% of incidents were excluded from the spatial descriptive analyses.
 

### 2021-08-24 scripts/calls_detroit.R

@@ -32,6 +32,9 @@ detroit_df <- detroit_df %>%
   filter(str_detect(call_timestamp, "2019"), # 2019 only.
          officerinitiated == "No")           # exclude officer initiated calls.
 
+# How many?
+
+
 # Date checks.
 detroit_df %>% 
   mutate(year = as.character(str_extract_all(call_timestamp, "^.{4}"))) %>% 
@@ -353,7 +356,7 @@ time_ann_gg <- time_gg +
            size = 9, x = 0.75, y = -0.065) +
   annotate(geom = "text", label = "proactive"      , colour = "dodgerblue4", 
            size = 9, x = 0.62, y = 1.065) +
-  annotate(geom = "text", label = "community"      ,  colour = "dodgerblue4", 
+  annotate(geom = "text", label = "community"      , colour = "dodgerblue4", 
            size = 9, x = 0.92, y = 1.065) +
   annotate(geom = "text", label = "health"         , colour = "dodgerblue4", 
            size = 9, x = 1.073, y = 0.58, angle = -90)
@@ -405,7 +408,7 @@ dh_agg_hm_list <- lapply(dh_agg_list, function(x){
   ggplot(data = x) +
     geom_tile(mapping = aes(x = time_lr, y = week_day, fill = mean_count)) +
     scale_x_discrete(labels = 1:24) +
-    scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue4") +
+    scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue3") +
     guides(fill = guide_colourbar(barwidth = 0.5, barheight = 4)) +
     labs(fill = NULL, x = NULL, y = NULL) +
     theme_minimal() +
@@ -526,7 +529,7 @@ grid_maps_list <- lapply(detroit19_grid_list, function(x){
   ggplot() +
     geom_sf(data = x, mapping = aes(fill = call_count), colour = "transparent") +
     geom_sf(data = detroit_uni_sf, fill = "transparent") +
-    scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue4", n.breaks = 5) +
+    scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue3", n.breaks = 5) +
     labs(fill = NULL) +
     guides(fill = guide_colourbar(barwidth = 0.5, barheight = 8)) +
     theme_void() +

### 2021-08-25 paper.Rmd
writing

@@
   those by Ratcliffe, namely, that a considerable proportion of police time is spent
-  resolving a diverse array of (non-criminal) calls for service, including public
+  resolving a diverse array of calls for service, including public
   health incidents. The temporal and spatial patterning of different types of public
@@
 
-Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for service in 2019.
-CAD codes, used the describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were classified into six broad categories: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
+CAD codes, used to describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were classified into six broad categories: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
 Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type.
-Findings indicated that around 55% of officer shift activity time was spent on crime incidents, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%).
+Findings indicated that around 55% of officer shift activity time was spent on calls relating to crime, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%).
 Focusing on medical/public health incidents, findings also indicated that calls concentrated in particular areas of the city and during particular hours of the day.
@@
 For each incident, the response time and time on the scene are reported.
-In sum, these comprise the total time officers spent on the incident.
-Incidents which had a total allocated time of zero minutes (or negative values) were excluded, leaving 301,490 calls for service throughout the year.
+In sum, these comprise the total time officers spend on the incident.
+Incidents which had a total allocated time of zero minutes (or negative values) were excluded.
+
 Each incident has a _calldescription_ variable which describes the nature of the call.
-There were 207 unique call description categories which in the interests of parsimony and ease of interpretation were recoded into 101 categories.
+Call descriptions involving administrative duties (e.g., "start of shift information"), completely unknown problems, and non-deployment (e.g., "employee call in / time off") were removed.
+This left 207 unique call description categories.
+In the interests of parsimony and ease of interpretation these categories were recoded into 101 broader call descriptions.
 ^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (xxxxxxxxxxxxxxxxxxxxxxxxxx).
-Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.
-Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing].
-For each of the six demand types, an 'other' classification was generated for those incident types were consumed less than 0.2% of police time during the year.
-This left a total of 48 categories for the reported breakdown.
+Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing.
+2.8% of incidents were deemed classifiable.
+For each of the six demand types, an 'other' classification was generated for those incident types which consumed less than 0.2% of police time during the year.
+This left a total of 48 categories for the reported breakdowns, capturing 267 thousand individual public calls for service during 2019.
 
-Each incident also has corresponding latitude and longitude coordinates and a call timestamp.
-Due to irregularities in geocoding, likely due unknown locations for some calls, 2% of incidents were excluded from the spatial descriptive analyses.
+In the spirit of @ratcliffe2021policing, the frequency counts and proportional breakdowns of demand type classifications are reported in Table 1.
+This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of issues.
+In Detroit, just 46% of police time was spent dealing with calls for service relating to crime (55% in Philadelphia).
+The remainder of time was spent dealing with community issues (6%), public and mental health (7%), proactive activity (7%), quality of life incidents (19%) and traffic duties (13%).
+The proportional breakdown of time consumed for each incident, grouped by demand type classification, are visualised in Figure 1.
+Here, it is clear that a small number incident types tend to dominate each demand classification.
+While incident descriptions are not always comparable, this visualisation confirms the diversity and complexity of public demand for police services reported in @ratcliffe2021policing.
+
+```{r table1}
+table1_df <- read_csv("results/table1_des_stats.csv")
+
+kable(table1_df, caption = "Breakdown of frequency counts and time committed for each demand type.",
+      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
+    kable_styling(latex_options = c("striped"),
+                  font_size = 10, full_width = F)
+```
 
@@
 
-```{r table1}
-table1_df <- read_csv("results/table1_des_stats.csv")
-
-kable(table1_df, caption = "Frequency and proportional breakdown (counts and time) for each demand type.",
-      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
-    kable_styling(latex_options = c("striped"),
-                  font_size = 10, full_width = F)
-```
+Each incident also has corresponding latitude and longitude coordinates and a call timestamp.
+Due to irregularities in geocoding, likely due unknown locations for some calls, 2% of incidents were excluded from the spatial descriptive analyses.
 

### 2021-08-25 scripts/calls_detroit.R

@@ -311,6 +311,10 @@ detroit_19_times_agg_df <- detroit_19_times_df %>%
 # Check missings.
 sum(is.na(detroit_19_times_agg_df)) # 0
 
+# How many incidents in total?
+sum(detroit_19_times_agg_df$freq) # 267737
+
+
 # Descriptive stats. 
 des_stats_df <- detroit_19_times_agg_df %>% 
   # mutate(type = str_to_title(type)) %>% 

### 2021-08-28 paper.Rmd
writing

@@
 Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can].
-In advance of recent contributions, this short paper aims to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
+In advance of recent contributions, this short paper uses comparable descriptive statistics to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
 
@@
 Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing.
-2.8% of incidents were deemed classifiable.
+2.8% of incidents were deemed unclassifiable.
 For each of the six demand types, an 'other' classification was generated for those incident types which consumed less than 0.2% of police time during the year.
@@
 
-In the spirit of @ratcliffe2021policing, the frequency counts and proportional breakdowns of demand type classifications are reported in Table 1.
+In the spirit of @ratcliffe2021policing, the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1.
 This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of issues.
@@
 The remainder of time was spent dealing with community issues (6%), public and mental health (7%), proactive activity (7%), quality of life incidents (19%) and traffic duties (13%).
+
 The proportional breakdown of time consumed for each incident, grouped by demand type classification, are visualised in Figure 1.
@@
 
-Each incident also has corresponding latitude and longitude coordinates and a call timestamp.
-Due to irregularities in geocoding, likely due unknown locations for some calls, 2% of incidents were excluded from the spatial descriptive analyses.
-
 ## Temporal patterning
@@
 
+Each incident also has corresponding latitude and longitude coordinates to describe the incident location.
+2% of incidents were excluded from the spatial descriptive analyses due to incomplete or unknown geographic information.
+Individual points were aggregated to 1000ft$^2$ grid cells laid over the city to generate basic descriptive heatmaps of counts during the year.
+These are visualised for each demand classification in Figure 3.
+
 \newpage

### 2021-08-30 paper.Rmd
writing

@@
 
-## Introduction
+# Introduction
 
@@
 Without a grasp on the scale ('how much?') and composition ('what type?') of police demand, we are likely to observe sub-optimal and inequitable outcomes for the public, the inappropriate distribution of public funds, and unnecessary strain on officers [@ellison2021understanding; @lum2021can].
-Understanding the characteristics public demand for the police has become particularly pertinent following recent calls to rethink, and some cases, radically reform, the role and reach of contemporary police forces [@lum2021can].
+Understanding the characteristics of public demand for the police has become particularly pertinent following recent calls to rethink, and some cases, radically reform, the role and reach of contemporary police forces [@lum2021can].
 
@@
 Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can].
-In advance of recent contributions, this short paper uses comparable descriptive statistics to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
+In advance of recent contributions, this short paper uses comparable analyses to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
 
-## Measuring demand
+# Ratcliffe (2021)
 
 Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
-CAD codes, used to describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were classified into six broad categories: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+CAD codes, used to describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were categorised into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
 Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type.
@@
 Focusing on medical/public health incidents, findings also indicated that calls concentrated in particular areas of the city and during particular hours of the day.
+
+# Public demand in Detroit
 
@@
 2.8% of incidents were deemed unclassifiable.
-For each of the six demand types, an 'other' classification was generated for those incident types which consumed less than 0.2% of police time during the year.
-This left a total of 48 categories for the reported breakdowns, capturing 267 thousand individual public calls for service during 2019.
+For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of police time during the year.
+This left a total of 48 categories for the reported breakdowns, capturing ~267,000 individual public calls for service during 2019.
 
-In the spirit of @ratcliffe2021policing, the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1.
-This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of issues.
-In Detroit, just 46% of police time was spent dealing with calls for service relating to crime (55% in Philadelphia).
+## Scale and composition
+
+In alignment with @ratcliffe2021policing, the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1.
+This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
+In Detroit, just 46% of police time was spent dealing with calls for service relating to crime.
 The remainder of time was spent dealing with community issues (6%), public and mental health (7%), proactive activity (7%), quality of life incidents (19%) and traffic duties (13%).
-
-The proportional breakdown of time consumed for each incident, grouped by demand type classification, are visualised in Figure 1.
-Here, it is clear that a small number incident types tend to dominate each demand classification.
-While incident descriptions are not always comparable, this visualisation confirms the diversity and complexity of public demand for police services reported in @ratcliffe2021policing.
+The proportional breakdown of time consumed for each incident type, grouped by the six demand classifications, are visualised in Figure 1.
 

### 2021-09-01 paper.Rmd
writing

@@
 Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can].
-In advance of recent contributions, this short paper uses comparable analyses to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
+In advance of recent contributions, this short paper describes the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
 
@@
 Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
-CAD codes, used to describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were categorised into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+CAD codes, discrete descriptions of the incident based on information available to the dispatcher, of which eighty unique descriptions originated from the public and required officer dispatch, were categorised into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
 Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type.
@@
 For each incident, the response time and time on the scene are reported.
-In sum, these comprise the total time officers spend on the incident.
+In sum, these comprise the total deployed time officers spend on the incident.
 Incidents which had a total allocated time of zero minutes (or negative values) were excluded.
@@
 For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of police time during the year.
-This left a total of 48 categories for the reported breakdowns, capturing ~267,000 individual public calls for service during 2019.
+This left a total of 48 categories for the reported breakdowns, capturing ~267,000 individual public calls for service in Detroit during 2019.
 
@@
 \newpage
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category. Graphic inspired by Ratcliffe (2012).", out.width = "\\textwidth"}
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category. Graphic inspired by Ratcliffe (2021).", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig1_time_port.png")
@@
 
+The mean incident counts by day of the week and by hour, for each demand classification, are visualised in Figure 2.
+This demonstrates that the different sources of public demand for police services, as indicated by the six demand classifications, can have distinct temporal patterning.
+Community calls appear to concentrate on weekdays, around lunchtime.
+Crime and quality of life-related calls both appear to occur in the evening and early hours of the morning, although crime incidents are more impacted by weekend activity stretching into Saturday and Sunday mornings.
+Traffic incidents concentrate in the evenings, particularly on a Friday.
+Proactive policing does not have a distinct hourly patterning, but this could be attributed to the low hourly counts.
+Calls relating to health issues, the primary interest of @ratcliffe2021policing, appear to largely take place in the evening and early morning, aligning with findings from Philadelphia.
+
+## Spatial patterning
+
+Individual incident location coordinates were aggregated to synthetic grid cells laid over the city to generate heatmaps of counts throughout the year.
+^[Grid cells were defined as 1000$^2$, a size deemed appropriate based on a balance between capturing localised variation and interpretable visualisations.
+2% of incidents were excluded from the spatial visualisations due to incomplete geographic information.]These are visualised for each demand classification in Figure 3.
+
 \newpage
@@
 ```
-\newpage
 
-## Spatial patterning
-
-Each incident also has corresponding latitude and longitude coordinates to describe the incident location.
-2% of incidents were excluded from the spatial descriptive analyses due to incomplete or unknown geographic information.
-Individual points were aggregated to 1000ft$^2$ grid cells laid over the city to generate basic descriptive heatmaps of counts during the year.
-These are visualised for each demand classification in Figure 3.
-
-\newpage
 ```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts for each demand type.", out.width = "\\textwidth"}
@@
 
-## Discussion
+# Discussion
+
+# References

### 2021-09-01 scripts/calls_detroit.R

@@ -355,15 +355,19 @@ time_ann_gg <- time_gg +
   annotate(geom = "text", label = "crime"          , colour = "dodgerblue4", 
            size = 9, x = 0.22, y = 1.065) +
   annotate(geom = "text", label = "traffic"        , colour = "dodgerblue4", 
-           size = 9, x = 0.67, y = 0.36) +
+           size = 9, x = 0.82, y = 0.58, angle = -90) +
   annotate(geom = "text", label = "quality of life", colour = "dodgerblue4", 
            size = 9, x = 0.75, y = -0.065) +
   annotate(geom = "text", label = "proactive"      , colour = "dodgerblue4", 
            size = 9, x = 0.62, y = 1.065) +
   annotate(geom = "text", label = "community"      , colour = "dodgerblue4", 
            size = 9, x = 0.92, y = 1.065) +
-  annotate(geom = "text", label = "health"         , colour = "dodgerblue4", 
-           size = 9, x = 1.073, y = 0.58, angle = -90)
+  annotate(geom = "text", label = "health"         , colour = "dodgerblue4",
+           size = 9, x = 1.07, y = 0.58, angle = -90) +
+  annotate(geom = "text", label = "PPO = Personal Protection Order",
+           size = 4, x = -0.045, y = -0.03, hjust = 0) +
+  annotate(geom = "text", label = "UDAA = Unlawfully Driving Away of an Automobile",
+           size = 4, x = -0.045, y = -0.04, hjust = 0)
 
 # Save portrait version.
 ggsave(filename = "visuals/fig1_time_port.png", height = 48, width = 40, unit = "cm", dpi = 300)
@@ -418,7 +422,8 @@ dh_agg_hm_list <- lapply(dh_agg_list, function(x){
     theme_minimal() +
     theme(legend.text = element_text(size = 5),
           axis.text   = element_text(size = 6), 
-          legend.text.align = 0.5)
+          legend.text.align = 0.5,
+          legend.text = element_text(size = 11))
 })
 
 # Arrange and annotate graphic.
@@ -535,18 +540,19 @@ grid_maps_list <- lapply(detroit19_grid_list, function(x){
     geom_sf(data = detroit_uni_sf, fill = "transparent") +
     scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue3", n.breaks = 5) +
     labs(fill = NULL) +
-    guides(fill = guide_colourbar(barwidth = 0.5, barheight = 8)) +
[…]

### 2021-09-01 paper.Rmd
writing and maps

@@
 Individual incident location coordinates were aggregated to synthetic grid cells laid over the city to generate heatmaps of counts throughout the year.
-^[Grid cells were defined as 1000$^2$, a size deemed appropriate based on a balance between capturing localised variation and interpretable visualisations.
-2% of incidents were excluded from the spatial visualisations due to incomplete geographic information.]These are visualised for each demand classification in Figure 3.
+^[Grid cells were defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localised variation and interpretable visualisations.
+2% of incidents were excluded from the spatial visualisations due to incomplete geographic information.] These are visualised for each demand classification in Figure 3.
+
+- Community: Wayne State University buildings.
+- Crime: Henry Ford Hospital, including the emergency department.
+         The hotspot in the far east is also a hosptial, as is the north west one.
+- Proactive: Wayne State University buildings, Midtown neighbourhood, just outside
+  downtown.
+- Quality of life: university disticts, downtown detroit and financial district.
+- Health: downtown, universities, schools and detroit rescue mission ministries
 

### 2021-09-01 scripts/calls_detroit.R

@@ -1,4 +1,5 @@
 # Load libraries.
+library(ggspatial)
 library(nngeo)
 library(cowplot)
 library(readr)
@@ -422,8 +423,7 @@ dh_agg_hm_list <- lapply(dh_agg_list, function(x){
     theme_minimal() +
     theme(legend.text = element_text(size = 5),
           axis.text   = element_text(size = 6), 
-          legend.text.align = 0.5,
-          legend.text = element_text(size = 11))
+          legend.text.align = 0.5)
 })
 
 # Arrange and annotate graphic.
@@ -519,7 +519,7 @@ detroit19_deploy_clip_list <- detroit19_deploy_clip_sf %>%
   filter(type != "unclassified") %>% 
   group_split(type)
 
-# Create list of the duplicate grid sf objects to match. Not ideal!
+# Create list of the duplicate grid sf objects to match. NOT IDEAL! Forgive me.
 grids_list <- c(detroit_grid_sf, detroit_grid_sf, detroit_grid_sf,
                 detroit_grid_sf, detroit_grid_sf, detroit_grid_sf)
 
@@ -533,17 +533,44 @@ p2p_fun <- function(x, y){
 # Run p2p through lists of incidents and duplicate grids.
 detroit19_grid_list <- map2(grids_list, detroit19_deploy_clip_list, p2p_fun)
 
+# Check counts for each demand type. Note these counts as slightly lower than
+# the Table 1 counts because we lost 2% due to incomplete coordinates.
+# This confirms that the list remains in alphabetical order.
+lapply(detroit19_grid_list, function(x){sum(x$call_count)})
+
+# Save specific maps for exploration in QGIS (optional).
+names(detroit19_grid_list) <- unique(dh_agg_df$type)
+
+for (i in 1:length(detroit19_grid_list)) {
+  st_write(obj = detroit19_grid_list[[i]],
+           dsn = paste("results/", names(detroit19_grid_list[i]), "_map.shp", sep = ""))
+}
+
 # Generate maps of incident counts by type.
 grid_maps_list <- lapply(detroit19_grid_list, function(x){
   ggplot() +
     geom_sf(data = x, mapping = aes(fill = call_count), colour = "transparent") +
     geom_sf(data = detroit_uni_sf, fill = "transparent") +
-    scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue3", n.breaks = 5) +
[…]

### 2021-09-01 paper.Rmd
map legend and writing

@@
 output:
-  # word_document: default
-  pdf_document: default
+  word_document: default
+  # pdf_document: default
 header-includes: \usepackage{lscape}
@@
 2% of incidents were excluded from the spatial visualisations due to incomplete geographic information.] These are visualised for each demand classification in Figure 3.
+While there is overlap, call hotspots and the degree of concentration differs between types of public demand.
+Calls about community issues concentrate overwhelmingly in and around the Wayne State University (WSU).
+Crime calls tend to cluster near hospitals, most notably, the Henry Ford Hospital in the city centre.
+Proactive call activity originates from the WSU area and Midtown, and quality of life around WSU and downtown.
+Health calls concentrate in downtown, but also around WSU and in the vicinity of the Detroit Rescue Mission Ministries, a large facility for the vulnerable, abused and homeless people in the city.
 
-- Community: Wayne State University buildings.
-- Crime: Henry Ford Hospital, including the emergency department.
-         The hotspot in the far east is also a hosptial, as is the north west one.
-- Proactive: Wayne State University buildings, Midtown neighbourhood, just outside
-  downtown.
-- Quality of life: university disticts, downtown detroit and financial district.
-- Health: downtown, universities, schools and detroit rescue mission ministries
+# Discussion
 
@@
 
-# Discussion
-
 # References

### 2021-09-01 scripts/calls_detroit.R

@@ -568,8 +568,8 @@ grid_maps_list <- lapply(detroit19_grid_list, function(x){
 grid_maps_list[[5]] <- grid_maps_list[[5]] +
   scale_fill_continuous(guide = "colourbar", low = "snow", high = "dodgerblue3", n.breaks = 2,
                         limits = c(0,20+max(detroit19_grid_list[[5]]$call_count))) +
-  annotation_scale(pad_x = unit(2, "cm"), pad_y = unit(4, "cm"), line_width = 1) +
-  annotation_north_arrow(pad_x = unit(3.8, "cm"), pad_y = unit(1, "cm"),
+  annotation_scale(pad_x = unit(1, "cm"), pad_y = unit(0.1, "cm"), line_width = 1, text_cex = 1, style = "ticks") +
+  annotation_north_arrow(pad_x = unit(2.8, "cm"), pad_y = unit(1.5, "cm"),
                          style = north_arrow_fancy_orienteering)
 
 # Arrange maps.
@@ -581,4 +581,3 @@ maps_gg <-  plot_grid(plotlist = grid_maps_list,
                       scale = 1.05)
 # Save.
 ggsave(filename = "visuals/fig3_maps.png", height = 48, width = 40, unit = "cm", dpi = 300)
-  

### 2021-09-16 paper.Rmd
map annotations

--- 2021-09-01 paper.Rmd
@@
 title: "Describing the scale and composition of calls for service in Detroit"
-author: "Samuel Langton (Netherlands Institute for the Study of Crime and Law Enforcement)"
+author: Samuel Langton \dag
 bibliography: refs.bib
 output:
-  # word_document: default
   pdf_document: default
@@
 csl: apa.csl
-abstract: In the spirit of a recent contribution by Ratcliffe (2021) published in
-  _Crime Science_ this short paper describes the scale and composition of public demand
-  for police services in Detroit, United States, during 2019. Findings broadly confirm
-  those by Ratcliffe, namely, that a considerable proportion of police time is spent
-  resolving a diverse array of calls for service, including public
-  health incidents. The temporal and spatial patterning of different types of public
-  demand for police services are often distinct. The data and code to replicate analyses
-  are made openly available.
+abstract: This short paper describes the scale and composition of emergency public
+  demand for police services in Detroit, United States. The contribution is made in
+  replication of analyses reported by Ratcliffe (2021), which used data from Philadelphia.
+  Findings confirm those of Ratcliffe, namely, that requests for police services often
+  involve incidents which do not involve crime. Instead, police spend a considerable
+  proportion of deployment time performing a social service function, such as resolving
+  emergency incidents involving public (mental) health. The spatial and temporal patterning
+  of public demand for police services is distinct. In advance of the Philadelphia
+  study, the data and code used here are openly available for reproduction and re-use.
 ---
 
+\begin{center}
+
 Keywords: police, calls for service, 911, demand, reactive.
+
+\end{center}
+
+\dag Netherlands Institute for the Study of Crime and Law Enforcement, Amsterdam, Netherlands.
 
@@
 A study from @ratcliffe2021policing in _Crime Science_ sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States.
-The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside the traditional crime-fighting role of the police.
+The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside of the traditional crime-fighting role of the police.
 Instead, the origins of public demand for the police can often be attributed to a lack of supply and/or accessibility failure in other organisations, such as those providing support for people requiring (mental) health assistance [@wood2021can; @van2017law].
@@
 Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can].
-In advance of recent contributions, this short paper describes the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).
+In advance of recent contributions, this short paper describes the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
 
-# Ratcliffe (2021)
+<!-- # Ratcliffe (2021) -->
 
-Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
-CAD codes, discrete descriptions of the incident based on information available to the dispatcher, of which eighty unique descriptions originated from the public and required officer dispatch, were categorised into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+<!-- Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
+CAD codes, discrete descriptions of incidents based on information available to the dispatcher, were categorised into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
 Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type.
@@
 Focusing on medical/public health incidents, findings also indicated that calls concentrated in particular areas of the city and during particular hours of the day.
+-->
 
@@
 The data used for this study covers the city of Detroit in Michigan, United States.
-The City of Detroit authority publish calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/].
-The raw data includes both citizen-initiated 911 calls to request police services and officer-initiated calls spanning back to September 2016.
-In alignment with @ratcliffe2021policing, the data used here is subset for the year 2019 and excludes those calls initiated by an officer.
+The City of Detroit authority publish calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with @ratcliffe2021policing the data used here is subset for the year 2019 and excludes those calls initiated by an officer.
 For each incident, the response time and time on the scene are reported.
@@
 This left 207 unique call description categories.
-In the interests of parsimony and ease of interpretation these categories were recoded into 101 broader call descriptions.
-^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (xxxxxxxxxxxxxxxxxxxxxxxxxx).
-Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing.
+In the interests of parsimony and ease of interpretation these categories were recoded into 99 broader call descriptions.^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
+Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] This included the removal of incidents involving transporting prisoners and executing warrants, which were deemed to have been erroneously flagged as 'officer initiated'.
+Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing.
 2.8% of incidents were deemed unclassifiable.
 For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of police time during the year.
-This left a total of 48 categories for the reported breakdowns, capturing ~267,000 individual public calls for service in Detroit during 2019.
+This left a total of 46 categories for the reported breakdowns, capturing ~265,000 individual public calls for service in Detroit during 2019.
 
@@
 
-In alignment with @ratcliffe2021policing, the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1.
+In alignment with @ratcliffe2021policing the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1.
 This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
-In Detroit, just 46% of police time was spent dealing with calls for service relating to crime.
-The remainder of time was spent dealing with community issues (6%), public and mental health (7%), proactive activity (7%), quality of life incidents (19%) and traffic duties (13%).
+In Detroit, just 47% of police time was spent dealing with calls for service relating to crime.
+The remainder of time was spent dealing with community issues (7%), public and mental health (7%), proactive policing (5%), quality of life incidents (19%) and traffic duties (13%).
 The proportional breakdown of time consumed for each incident type, grouped by the six demand classifications, are visualised in Figure 1.
@@
 \newpage
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category. Graphic inspired by Ratcliffe (2021).", out.width = "\\textwidth"}
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category.", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig1_time_port.png")
@@
 
-## Temporal patterning
+## Spatio-temporal patterning
 
@@
 This demonstrates that the different sources of public demand for police services, as indicated by the six demand classifications, can have distinct temporal patterning.
-Community calls appear to concentrate on weekdays, around lunchtime.
+Community calls appear to concentrate on weekdays, around lunchtime, when other classes of demand are low.
 Crime and quality of life-related calls both appear to occur in the evening and early hours of the morning, although crime incidents are more impacted by weekend activity stretching into Saturday and Sunday mornings.
@@
 
-## Spatial patterning
-
 Individual incident location coordinates were aggregated to synthetic grid cells laid over the city to generate heatmaps of counts throughout the year.
@@
 2% of incidents were excluded from the spatial visualisations due to incomplete geographic information.] These are visualised for each demand classification in Figure 3.
+Calls about community issues concentrate overwhelmingly in and around the Wayne State University (WSU) and the Cultural Center.
+Crime calls tend to cluster in hotspots near hospitals, most notably, the Henry Ford Hospital.
+Proactive call activity originates from the WSU area and Midtown, as does quality of life calls, which are also spread around residential areas.
+Health calls concentrate in Midtown and to the east, in the immediate vicinity of major well-being and mental health centres.
+Traffic incidents have a distinct spatial patterning along arterial roads into downtown.
 
-- Community: Wayne State University buildings.
-- Crime: Henry Ford Hospital, including the emergency department.
-         The hotspot in the far east is also a hosptial, as is the north west one.
-- Proactive: Wayne State University buildings, Midtown neighbourhood, just outside
-  downtown.
-- Quality of life: university disticts, downtown detroit and financial district.
-- Health: downtown, universities, schools and detroit rescue mission ministries
+# Discussion
+
+In replication of recent findings in Philadelphia [@ratcliffe2021policing] this short contribution has sought to describe the scale and composition of calls for police services in Detroit, United States.
+The data and code to reproduce the results reported here are openly available.
+Findings provide robustness to the Philadelphia study: a considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police.
+Instead, the police fulfill a variety of other social service functions, including the resolution of public (mental) health incidents and resolving community issues.
+Findings further highlight the potential implications of radically reforming (defunding) police forces in the United States through [@lum2021can].
 
 \newpage
-```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and week, for each demand type.", out.width = "\\textwidth"}
+```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and hour, for each demand type.", out.width = "\\textwidth"}
 knitr::include_graphics("visuals/fig2_time_heat.png")

### 2021-09-16 scripts/calls_detroit.R

@@ -78,6 +78,7 @@ calldesc_df <- tibble(calldescription = unique(detroit19_deploy_df$calldescripti
   arrange(calldescription)
 
 # Recode raw calldescription into broader categories and rename to short characters as appropriate.
+# We also remove 'warrant' and 'transport prisoner' are considered not officer initiated.
 detroit19_deploy_df <- detroit19_deploy_df %>% 
   mutate(calldescription2 = if_else(str_detect(calldescription, "DV")                                      ,"DOMESTIC INCIDENT"          , calldescription),
          calldescription2 = if_else(str_detect(calldescription, "MENTAL")                                  ,"MENTAL HEALTH"              , calldescription2),
@@ -143,10 +144,11 @@ detroit19_deploy_df <- detroit19_deploy_df %>%
          calldescription2 = if_else(calldescription %in% c("DPD DEMS DFD N TRO","ASSIST PERSONNEL")              ,"ASSIST EMERGENCY PERSONNEL"            , calldescription2),
          calldescription2 = if_else(calldescription %in% c("CITIZEN RADIO PATROL IN TROUBL","ASSIST CITIZEN","SENIOR CITIZEN ASSIST")    ,"ASSIST CITIZEN", calldescription2),
          calldescription2 = if_else(calldescription %in% c("INVESTIGATE PERSON", "INVESTIGATE YOUTH(S)")    ,"INVESTIGATE PERSON(S)"     , calldescription2),
-         calldescription2 = if_else(calldescription %in% c("EMERGENCY STANDBY", "BE ON THE LOOK OUT")       ,"STAND-BY"                  , calldescription2))
+         calldescription2 = if_else(calldescription %in% c("EMERGENCY STANDBY", "BE ON THE LOOK OUT")       ,"STAND-BY"                  , calldescription2)) %>% 
+  filter(calldescription2 != "WARRANT" & calldescription2 != "TRANSPORT PRISONER") 
 
 # How many unique categories after recoding?
-length(unique(detroit19_deploy_df$calldescription2)) # 101
+length(unique(detroit19_deploy_df$calldescription2)) # 99
 
 # Create new df to consult recoding.
 calldesc2_df <- detroit19_deploy_df %>% 
@@ -188,9 +190,9 @@ proactive_vec <- c("BACKGROUND CHECK",
                    "ESCORT",
[…]

### 2021-09-17 scripts/calls_detroit.R
finalise

@@ -33,9 +33,6 @@ detroit_df <- detroit_df %>%
   filter(str_detect(call_timestamp, "2019"), # 2019 only.
          officerinitiated == "No")           # exclude officer initiated calls.
 
-# How many?
-
-
 # Date checks.
 detroit_df %>% 
   mutate(year = as.character(str_extract_all(call_timestamp, "^.{4}"))) %>% 

### 2021-09-18 scripts/calls_detroit.R
final edits

@@ -593,22 +593,21 @@ grid_maps_list[[5]] <- grid_maps_list[[5]] +
                          style = north_arrow_fancy_orienteering)
 
 # Annotate select maps with key locations.
+grid_maps_list[[1]] <- grid_maps_list[[1]] +
+  annotate(geom = "text"    , x = 13467040, y = 322652, label = "WSU campus & Midtown", size = 4) +
+  annotate(geom = "curve" , x = 13467040, y = 320672, xend = 13471640, yend = 316142, size = 0.7,
+           arrow = arrow(length = unit(0.01, "npc")), curvature = 0.3)
+
+
 grid_maps_list[[2]] <- grid_maps_list[[2]] +
   annotate(geom = "text"    , x = 13477040, y = 324802, label = "Henry Ford Hospital", size = 4) +
-  annotate(geom = "segment" , x = 13473240, y = 322952, xend = 13470640, yend = 318852, size = 0.7,
-           arrow = arrow(type = "closed", length = unit(0.01, "npc")))
-
-
-grid_maps_list[[4]] <- grid_maps_list[[4]] +
-  annotate(geom = "text"    , x = 13485040, y = 322652, label = "WSU campus, Midtown", size = 4) +
-  annotate(geom = "segment" , x = 13479240, y = 320652, xend = 13476040, yend = 316652, size = 0.7,
-           arrow = arrow(type = "closed", length = unit(0.01, "npc")))
+  annotate(geom = "curve" , x = 13473240, y = 322952, xend = 13470640, yend = 318852, size = 0.7,
+           arrow = arrow(length = unit(0.01, "npc")), curvature = -0.3)
 
 grid_maps_list[[6]] <- grid_maps_list[[6]] +
   annotate(geom = "text"    , x = 13490040, y = 301652, label = "Downtown", size = 4) +
-  annotate(geom = "segment" , x = 13484240, y = 301752, xend = 13481040, yend = 304752, size = 0.7,
-           arrow = arrow(type = "closed", length = unit(0.01, "npc")))
-
+  annotate(geom = "curve" , x = 13484240, y = 301752, xend = 13481040, yend = 304752, size = 0.7,
+           arrow = arrow(length = unit(0.01, "npc")), curvature = -0.3)
 
 # Arrange maps.
 maps_gg <-  plot_grid(plotlist = grid_maps_list,

### 2022-02-01 scripts/calls_detroit.R
shift counts add

@@ -306,6 +306,7 @@ detroit_19_times_agg_df <- detroit_19_times_df %>%
   group_by(calldescription2, type) %>% 
   summarise(freq            = sum(sum_counts),
             prop_time       = sum(prop_time),
+            sum_time        = sum(sum_time), # optional
             prop_count      = sum(prop_counts)) %>% 
   ungroup() %>% 
   arrange(type)
@@ -321,8 +322,8 @@ des_stats_df <- detroit_19_times_agg_df %>%
   group_by(type) %>% 
   summarise(`Count`      = sum(freq),
             `Count (%)`  = round(sum(prop_count), 2),
-            `Total deployed time (%)`   = round(sum(prop_time ), 2)) %>% 
-            # `Total deployed time (hours)` = sum_time_hours) %>%
+            `Total deployed time (%)`   = round(sum(prop_time ), 2),
+            `Total deployed time (shifts)` = round((sum(sum_time)/60)/7.5, 0) )  %>%
   rename(`Demand type` = type) %>% 
   ungroup()
 
@@ -448,7 +449,7 @@ time_heat_gg <- plot_grid(plotlist = dh_agg_hm_list,
            x = 0.5, y = 0, size = 2)
 
 # Save.
-ggsave(filename = "visuals/fig2_time_heat_total_time.png", height = 20, width = 14, unit = "cm", dpi = 300)
+ggsave(filename = "visuals/fig3_time_heat_total_time.png", height = 20, width = 14, unit = "cm", dpi = 300)
 
 # Investigate missings in coordinates.
 sum(is.na(detroit19_deploy_df$latitude))  # 0
@@ -535,7 +536,21 @@ detroit19_deploy_clip_sf <- detroit19_deploy_known_sf %>%
 detroit_grid_sf <- detroit_uni_sf %>% 
   # st_buffer(dist = 1000) %>% # To check if grids miss incidents near boundary.
   st_make_grid(cellsize = 1000) %>% 
-  st_as_sf()
+  st_as_sf() 
+
+# Plot.
+# ggplot() +
+#   geom_sf(data = detroit_grid_sf) +
+#   geom_sf(data = detroit_uni_sf, colour = "red", fill = "transparent")
+
+# Clip the grid to the Detroit boundary.
+detroit_grid_sf <- detroit_grid_sf %>% 
+  st_intersection(detroit_uni_sf)
+
+# Plot.
+# ggplot() +
+#   geom_sf(data = detroit_grid_sf) +
[…]

### 2022-02-01 scripts/calls_detroit_time_on_scene.R

@@ -9,12 +9,16 @@ library(purrr)
 library(forcats)
 library(stringr)
 library(lubridate)
+library(lorenzgini)
 library(treemapify)
 library(ggplot2)
 library(sf)
 
 # Useful function.
 `%nin%` <- Negate(`%in%`)
+
+# Load workspace.
+# load(file = "data/workspace_complete.RData")
 
 # 2016-2021 911 calls for service from Detroit open data portal.
 # https://data.detroitmi.gov/datasets/911-calls-for-service?geometry=-86.035%2C42.028%2C-80.808%2C42.738
@@ -457,7 +461,7 @@ time_heat_gg <- plot_grid(plotlist = dh_agg_hm_list,
            x = 0.5, y = 0, size = 2)
 
 # Save.
-ggsave(filename = "visuals/fig2_time_heat_tos.png", height = 20, width = 14, unit = "cm", dpi = 300)
+ggsave(filename = "visuals/fig3_time_heat_tos.png", height = 20, width = 14, unit = "cm", dpi = 300)
 
 # Investigate missings in coordinates.
 sum(is.na(detroit19_deploy_df$latitude))  # 0
@@ -555,6 +559,14 @@ detroit_grid_sf <- detroit_uni_sf %>%
   st_as_sf() %>% 
   mutate(grid_id = 1:nrow(.))
 
+# Clip the grid to the Detroit boundary.
+detroit_grid_sf <- detroit_grid_sf %>% 
+  st_intersection(detroit_uni_sf)
+
+# Plot.
+# ggplot(data = detroit_grid_sf) +
+#   geom_sf()
+
 # Check counts used in maps. Lower due to incomplete coordinates.
 sum(detroit19_deploy_clip_sf$n) # 246971
 
@@ -595,8 +607,7 @@ write_csv(x = means_table_df, file = "results/table2_des_stats_tos.csv")
 names(detroit19_deploy_clip_sf)
 nrow(detroit19_deploy_clip_sf)
 
-# Concentration statistics.
-library(lorenzgini)
+# Concentration statistics (tangent).
 
 # Split time on scene into simple tibble by demand classification.
 demands_time_list <-  detroit19_deploy_clip_sf %>% 
@@ -627,20 +638,55 @@ lorenz_results_list <- lapply(lorenz_plot_list, lorenz_pull)
 lorenz_results_df <- lorenz_results_list %>% 
   bind_rows(.id = "type")
 
[…]

### 2022-02-01 paper.Rmd
paper update
including back-up of manuscript

@@
 ---
-title: "Describing the scale and composition of calls for service in Detroit"
-author: Samuel Langton \dag
+title: 'Describing the scale and composition of calls for police service: a replication and
+  extension using open data'
+author: Samuel Langton\dag, Stijn Ruiter\dag, Tim Verlaan\dag
 bibliography: refs.bib
 output:
-  pdf_document: default
-header-includes: \usepackage{lscape}
+  pdf_document: 
+    number_sections: true
+header-includes:
+  - \usepackage{lscape}
+  - \usepackage{lineno}
+  - \linenumbers
 csl: apa.csl
-abstract: This short paper describes the scale and composition of emergency public
-  demand for police services in Detroit, United States. The contribution is made in
-  replication of analyses reported by Ratcliffe (2021), which used data from Philadelphia.
-  Findings confirm those of Ratcliffe, namely, that requests for police services often
-  involve incidents which do not involve crime. Instead, police spend a considerable
-  proportion of deployment time performing a social service function, such as resolving
-  emergency incidents involving public (mental) health. The spatial and temporal patterning
-  of public demand for police services is distinct. In advance of the Philadelphia
-  study, the data and code used here are openly available for reproduction and re-use.
+abstract: "This paper describes the scale and composition of emergency public demand for police services in Detroit, United States.
+The contribution is made in replication and extension of analyses reported by Ratcliffe (2021) and Lum et al.
+(2021).
+Findings indicate that police spend a considerable proportion of time performing a social service function.
+Just 47% of total deployed time responding to 911 calls is consumed by crime incidents.
+The remainder is spent on community issues (6%), health (7%), proactive (5%), quality of life (19%) and traffic duties (13%).
+A small number of calls consume a disproportionately large amount of police officer time.
+Emergency demand is concentrated in time and space, and can differ between types of demand.
+Findings further highlight the potential implications of radically reforming police forces in the United States.
+In advance of existing studies, the data and code used here are openly available for reproduction, re-use and scrutiny."
 ---
@@
 
-\dag Netherlands Institute for the Study of Crime and Law Enforcement, Amsterdam, Netherlands.
+\dag Netherlands Institute for the Study of Crime and Law Enforcement, Amsterdam, Netherlands. \linebreak \linebreak \linebreak \linebreak
+
+**ORCID** Samuel Langton (0000-0002-1322-1553). Stijn Ruiter (0000-0003-2872-2710). Tim Verlaan (0000-0002-3048-5817).
+
+**Acknowledgments** To be completed.
+
+**Disclosure statement** The author declares that they have no competing interests.
+
+**Funding details** To be completed.
+
+**Availability of data** The data and code used in this study are openly available (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
+
+\newpage
 
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message=F, comment=F, warning=F, fig.pos = 'H')
+knitr::opts_chunk$set(echo = F, message=F, comment=F, warning=F, fig.pos = 'H', eval = T)
 library(readr)
+library(readxl)
 library(dplyr)
@@
 
-# Introduction
+# Background
 
@@
 Without a grasp on the scale ('how much?') and composition ('what type?') of police demand, we are likely to observe sub-optimal and inequitable outcomes for the public, the inappropriate distribution of public funds, and unnecessary strain on officers [@ellison2021understanding; @lum2021can].
-Understanding the characteristics of public demand for the police has become particularly pertinent following recent calls to rethink, and some cases, radically reform, the role and reach of contemporary police forces [@lum2021can].
+Understanding the characteristics of public demand for the police has become particularly pertinent following recent calls to rethink, and in some cases, radically reform, the role and reach of contemporary police forces [@lum2021can].
 
-A study from @ratcliffe2021policing in _Crime Science_ sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States.
+A recent study from @ratcliffe2021policing sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States (US), during 2019.
 The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside of the traditional crime-fighting role of the police.
 Instead, the origins of public demand for the police can often be attributed to a lack of supply and/or accessibility failure in other organisations, such as those providing support for people requiring (mental) health assistance [@wood2021can; @van2017law].
-The study's findings provided data-driven evidence to suggest that the police indeed spend considerable amounts of time dealing with incidents that could be resolved by other -- often more appropriate -- agencies.
-Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can].
-In advance of recent contributions, this short paper describes the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
+Ratcliffe reported that just 40% of public emergency calls for service during the year were crime incidents.
+These calls consumed 55% of total officer shift time.
+The remaining public demand from calls involved community issues (10% of calls, 7% of time), medical/public health incidents (7% of calls, 9% of time), proactive policing (8% of calls, 5% of time), quality of life (24% of calls, 14% of time) and traffic duties (11% of calls and 11% of time).
+Comparable results were subsequently reported in the US across multiple anonymised police jurisdictions [@lum2021can; see also @lumwash in _The Washington Post_].
 
-<!-- # Ratcliffe (2021) -->
+These descriptive findings provided data-driven evidence to suggest that the police indeed spend considerable amounts of time dealing with incidents in a social rather than solely crime-fighting capacity.
+This has implications to and beyond policing practice.
+Findings directly contribute to the fierce and politically-charged debate around the role and responsibilities of police officers in contemporary society.
+Both @ratcliffe2021policing and @lum2021can were published amidst the backdrop of the killing of George Floyd in May, 2020 and subsequent 'defund the police' protests across the US.
+Evidence which demonstrates that the public rely on the police for social duties beyond law enforcement suggests that a radical reduction in their role and responsibilities (e.g., through defunding) would increase demand on other social services.
+Should such radical reform take place without first understanding the scale and composition of public demand for the police, and without preemptively providing appropriate substitutes, there will be no supply to meet this shift in demand.
+Thus, @ratcliffe2021policing and @lum2021can contributed important empirical data to the (ongoing) discussion on police reform in the US.
 
-<!-- Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
-CAD codes, discrete descriptions of incidents based on information available to the dispatcher, were categorised into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
-Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type.
-Findings indicated that around 55% of officer shift activity time was spent on calls relating to crime, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%).
-Focusing on medical/public health incidents, findings also indicated that calls concentrated in particular areas of the city and during particular hours of the day.
--->
+That said, a notable shortcoming of these recent contributions is the lack of openly available data and code to reproduce and replicate findings.
+This is not an uncommon issue in policing research.
+A recent scoping review of Canadian policing studies concluded that there has been "little effort...
+to replicate or reproduce findings", with just 8% of the literature sampled constituting replications or reproductions [@huey2017replication, p.
+113].
+More broadly, in a sample of papers from top criminology journals, @mcneeley2015replication found that just 2% of papers were replications -- either a direct replication of the same population, or empirical generalizations from a different population.
+The scarcity of replication has been identified as major obstacle to the goals of evidence-based policing, namely, "to develop a rigorous evidence base that can be used by police professionals to develop more efficient and effective police practices, policies, and/or programs" [@openedp2019, p.
+126].
+In the long-term, open science practices are thought to provide a more concrete base from which police practitioners can make real-life decisions [@openedp2019] and comprise a fundamental component of related fields such as crime science [@laycock2013defining].
+It therefore seems appropriate that the investigations of @ratcliffe2021policing and @lum2021can are replicated using open data and methods.
 
-# Public demand in Detroit
+Here, the necessity for a fully open investigation into the scale and composition of police demand is perhaps especially important.
+This is for two reasons.
+First, the implications of redefining the role of the police is considerable.
+It is reasonable for police practitioners and the public to expect a thorough, open and replicable empirical investigation into police demand across numerous contexts before academia concludes that the issue is resolved -- certainly before findings are used to change or introduce police practice or public policy.
+As noted, committing to radical reform without such investigations might result in shifts in public demand for services for which there is no suitable supply.
+Second, the data handling required to describe the public demand for the police is not straightforward.
+Reporting the scale and composition of emergency calls for service data necessitates the grouping of incident types through categorization (e.g., 'mental health', 'community').
+As both @ratcliffe2021policing and @lum2021can point out, categorizing incident types is challenging and subject to a degree of interpretation.
+This issue is exacerbated by the fact that different police jurisdictions will use different incident categories.
+Thus, an open replication, inclusive of the categorization decisions, facilitates scrutiny through transparency and provides robustness to existing findings.
 
-The data used for this study covers the city of Detroit in Michigan, United States.
-The City of Detroit authority publish calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with @ratcliffe2021policing the data used here is subset for the year 2019 and excludes those calls initiated by an officer.
-For each incident, the response time and time on the scene are reported.
+With these points in mind, this paper describes the emergency demand for police services in a large city, Detroit, US, using open data and open code.
+All materials are publicly available (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
+It does so in replication--and where appropriate, extension--of the main findings reported by @ratcliffe2021policing and later @lum2021can.
+In this way, the contribution serves to test the robustness of existing findings, but also provides the materials required to reproduce and replicate analyses in the interests of transparency and public scrutiny.
+
+# Describing demand
+
+Given the aims of this study, this section first provides a brief outline of the data and methods used by @ratcliffe2021policing and @lum2021can to describe the scale and composition of emergency reactive police demand.
+
+## Ratcliffe (2021)
+
+Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
+Calls were subset to exclude those which did not require officer dispatch or did not originate from the public.
+CAD codes, discrete descriptions of incidents based on information available to the dispatcher, were categorized into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
+The classification scheme was used to report on the frequency and proportional breakdown of calls relating to each type of demand throughout 2019.
+
+Recognising that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by public demand, raw counts were augmented with the time spent by officers on each of the six classifications of demand.
+This information permits the calculation of a proportional breakdown of police time consumed by the different demand classifications throughout the year.
+The raw numbers were reported in a table.
+A visual representation of call types nested within each demand classification was summarised in a tree map graphic.
+The time committed by officers was calculated from when the call was dispatched to an officer to when the officer indicated that the event has been closed.
+Here, this is referred to as the _total deployed time_.
+The aggregate time consumed by different demand classifications was reported using a single count of officer shifts (7.5 hours).
+
+The spatio-temporal patterning of this demand was summarised in two visuals, focusing specifically on those calls relating to medical/public health.
+A heatmap showed the mean call volume during hours of the day and days of the week, identifying peak times for demand.
+A geographic map visualized the spatial patterning of these call count concentrations using kernel density in order to identify hotspots of police demand involving medical/public health issues.
+To supplement these findings, Ratcliffe also reported on the differences between the initial CAD incident type classification and the final disposition using a sankey diagram -- again focusing specifically on medical and public health calls.
+
+## Lum et al (2021)
+
+To describe the scale and composition of emergency police demand across the US more broadly, @lum2021can used CAD data from nine different police agencies.
+The agencies themselves were anonymized.
+Individual incidents were categorized into fourteen different demand classifications, namely, alarms, disorders, vice, domestic, follow-ups and service requests, mental, medical, missing persons, violence, interpersonal-other, property, suspicions, traffic, and admin, agency and non-crime.
+The authors reported a proportional breakdown of call counts and police time spent across the nine agencies according to the fourteen demand classifications, including the average (mean).
+For each agency, the authors also reported the average time spent on each demand classification in minutes, and the average across the nine agencies.
+No descriptive statistics on the range or spread of these figures was reported.
+The 'time spent' measure used was the _time on scene_.
+The authors considered this to be preferable to _total deployed time_ when measuring police demands, since low priority calls may have a non-urgent (slow) response due to the nature of the call, and officers may be reassigned to other duties while on an assignment.
+A proportional breakdown of the disposition (assist only, report written or arrest citation) was provided for each of the fourteen call types.
+No spatial or temporal findings were reported.
+
+# Data
+
+## Detroit calls for service
+
+The data used for this study covers the city of Detroit in Michigan, US.
+The City of Detroit authority publish 911 emergency calls for service data through their Open Data Portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with @ratcliffe2021policing the data used here is subset for the year 2019 and excludes those calls initiated by an officer in order to capture public-initiated calls.
+Each individual incident has a corresponding timestamp and location coordinates to define the call in time and space.
+
+For each incident, the police response time and time on the scene are reported.
 In sum, these comprise the total deployed time officers spend on the incident.
-Incidents which had a total allocated time of zero minutes (or negative values) were excluded.
+Calls which had a _total deployed time_ (response time + time on scene) or _time on scene_ of zero minutes, negative values or missing were excluded.
+The data permits a distinction between the time spent resolving incidents with and without the response time.
+As noted, @lum2021can considered the _time on scene_ measure to be preferable.
+Here, for completeness, both measurements are reported.
 
@@
 In the interests of parsimony and ease of interpretation these categories were recoded into 99 broader call descriptions.^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
-Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] This included the removal of incidents involving transporting prisoners and executing warrants, which were deemed to have been erroneously flagged as 'officer initiated'.
-Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing.
+Readers unfamiliar with R code can review the reference table entitled 'categorization summary'.] This included the removal of incidents involving transporting prisoners and executing warrants, which were deemed to have been erroneously flagged as not being officer-initiated.
+Each call description was then categorized into the six demand type classifications used in @ratcliffe2021policing, namely, crime, community, medical/public health, proactive, quality of life and traffic.
 2.8% of incidents were deemed unclassifiable.
-For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of police time during the year.
-This left a total of 46 categories for the reported breakdowns, capturing ~265,000 individual public calls for service in Detroit during 2019.
+For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of total deployed time during the year.
+This left a total of 49 unique categories for the reported breakdowns, capturing ~260,000 individual public calls for service in Detroit during 2019.
+No information on the final disposition of calls is provided in the open data.
+
+# Methods
+
+The methods used largely replicate those used by @ratcliffe2021policing and @lum2021can, and where appropriate, extensions.
+First, descriptive statistics and a treemap graphic provide an account of the scale and composition of police demand during the year.
+Second, visualizations summarise the spatial and temporal patterning of this demand.
+
+## Descriptive statistics
+
+Summary statistics will describe the overall composition of demand according to the six demand type classifications.
+This includes call frequency counts, a proportional breakdown of call counts, and a proportional breakdown of police time spent on each demand type.
+Here, rather than reporting one measure in isolation, both _total deployed time_ and _time on scene_ measures of police resource are reported.
+The proportional breakdown of _time on scene_ is visualized using a treemap graphic for each incident type nested within the demand classifications.
+The equivalent graphic for _total deployed time_ is available in the corresponding GitHub repository.
+
+A series of descriptive statistics summarise the time consumed (in minutes) by each demand type.
+For comprehensiveness and transparency, this includes the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution.
+This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.
+Such a skew would suggest that averages of time [e.g., @lum2021can] or aggregate measures of time [e.g., @ratcliffe2021policing] in isolation are insufficient and potentially mask important underlying variation, namely, that a small number of high-demand calls drag up the mean and inflate aggregate measures of time.
+To investigate this distribution, we plot Lorenz curves for each demand classification.
+Lorenz curves plot the cumulative distribution of a variable (in this case, time on scene) against the cumulative distribution of a unit (in this case, calls) [@bernasco2017more].
+By way of example, we report on these figures according to arbitrary thresholds of 25% and 50%.
+This sheds light on the degree to which disproportionately large amount of officer time might be consumed by just a small number of calls.
+
+## Spatio-temporal patterning
+
+The temporal patterning of all six demand type classifications are visualized using a heatmap of mean call counts, aggregated by day of the week, and hour of the day.
+To summarize the spatial patterning, call location coordinates were aggregated to synthetic grid cells using spatial joins.
+Grid cells were defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localized variation and interpretable visualizations.
+2% of incidents were excluded from the spatial visualizations due to incomplete geographic information.
+To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year, for each demand type classification.
+The equivalent raw count maps are available in the Appendix.
+In this, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, but here, we provide a comparison across different demand classifications and two different measures of demand.
+
+## Data and methods summary
+
+To summarise the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
+Broadly, this demonstrates that, with the exception of the final disposition, this study provides a comprehensive replication of the analyses reported by @ratcliffe2021policing and @lum2021can.
+Either identical descriptive statistics are reported in replication of one or both existing studies, or additional analyses are provided for enhanced insight and transparency.
+In sum, the replications and extensions summarised here, coupled with the open code and open data to reproduce findings, constitute the primary contributions of this study.
+
+\newpage
+
+```{r table1}
+options(knitr.kable.NA = '')
+
+table1_df <- read_xlsx("results/table1_methods_comparison.xlsx")
+
+kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.",
+      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
+    kable_styling(latex_options = c("striped"),
+                  font_size = 10, full_width = F) %>% 
+  column_spec(column = c(2,3), width = c("2cm", "2cm") ) %>% 
+  row_spec(row = 14, hline_after = 14) %>%
+  row_spec(row = 15, italic = T)
+```
+
+# Results
 
@@
 
-In alignment with @ratcliffe2021policing the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1.
-This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
-In Detroit, just 47% of police time was spent dealing with calls for service relating to crime.
-The remainder of time was spent dealing with community issues (7%), public and mental health (7%), proactive policing (5%), quality of life incidents (19%) and traffic duties (13%).
-The proportional breakdown of time consumed for each incident type, grouped by the six demand classifications, are visualised in Figure 1.
+The frequency counts and proportional breakdowns for each demand type classification are reported in Table 2.
+This demonstrates that in Detroit, a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
+Just 44% of emergency calls for service related to crime incidents.
+These crime incidents consumed 47% of total deployed time throughout the year, although the figure increases to 51% when considering time on scene in isolation.
+The remainder of total time on scene is spent dealing with community issues (5%), public and mental health (7%), proactive policing (4%), quality of life incidents (15%) and traffic duties (16%).
 
-```{r table1}
-table1_df <- read_csv("results/table1_des_stats.csv")
+```{r table2}
+table2a_df <- read_csv("results/table2_des_stats_total_time.csv")
+table2b_df <- read_csv("results/table2_des_stats_tos.csv")
 
-kable(table1_df, caption = "Breakdown of frequency counts and time committed for each demand type.",
+table2_df <- table2a_df %>% 
+  select(-`Demand type`, -Count, -`Count (%)`) %>%  
+  bind_cols(table2b_df) %>% 
+  select(`Demand type`, Count, `Count (%)`,
+         `Time on scene (%)`, `Total deployed time (%)`,  `Total deployed time (shifts)`) %>%  # reorder for final table.
+  rename( `Total deployed \n time (shifts)` =  `Total deployed time (shifts)`)
+
+kable(table2_df, caption = "Breakdown of frequency counts and time consumed for each demand type. Total deployed time is the sum of the time on scene and response time.",
       booktabs = T, longtable = T, linesep = "", format = "latex") %>%
     kable_styling(latex_options = c("striped"),
-                  font_size = 10, full_width = F)
+                  font_size = 10, full_width = F) %>% 
+  column_spec(column = 1:6, width = c("2cm","2cm","2cm","2.5cm","2.5cm","2.5cm"))
+```
+
+The proportional breakdown of time spent on scene for each incident type, grouped by the six demand classifications, is visualized in Figure 1.^[The purpose of the graphic is to provide an broad summary, in replication of Ratcliffe's visual.
+The numbers underlying the graphic are reported in the corresponding GitHub repository.] This figure demonstrates the diversity and complexity of public demand for police services in Detroit.
+Time consumed on the scene of crime incidents is largely comprised of domestic incidents, burglary and assault.
+For traffic incidents, police time was mostly spent attending to accidents, while for quality of life, attending to disturbances consumed the most time on scene.
+Calls relating to public health involved mental health incidents, overdoses, (threats of) suicide  and well-being checks.
+Proactive incidents largely involved investigations, while time consumed on community issues consisted mostly of missing/found persons and animal incidents (e.g., aggressive dogs).
+
+## Time distributions
+
+To further quantify the resources consumed, Table 3 summarises the time spent on the scene for each demand classification in minutes.
+A histogram of the underlying distributions are reported in the Appendix.
+Table 3 demonstrates a number of things.
+First, there is a modest amount of variation in the typical time at scene of different types of emergency call.
+For instance, the mean time spent on the scene of traffic incidents (55 minutes) and crime (51 minutes) is considerably more that quality of life issues (31 minutes) or proactive activity (28 minutes).
+These differences will be a result of a combination of factors, including the complexity and severity of incidents and the number of officers required to attend the scene.
+Second, there is considerable variation in the time consumed by incidents _within_ each demand type.
+While many calls are resolved in minutes, a handful consume hours of cumulative officer time, resulting in a highly skewed distribution.
+This highlights how mean or aggregate summary statistics of police time as a measure of demand can mask (important) underlying variation.
+
+```{r table3}
+table3_df <- read_csv("results/table3_des_stats_tos.csv")
+
+kable(table3_df, caption = "Descriptive statistics of minutes spent on scene for each demand type classification.",
+      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
+    kable_styling(latex_options = c("striped"),
+                  font_size = 10, full_width = F) %>% 
+  column_spec(column = 1:6, width = c("2.5cm","1cm","1cm","1cm","1.2cm", "1cm"))
 ```
@@
 \newpage
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of deployed time spent on each incident category.", out.width = "\\textwidth"}
-knitr::include_graphics("visuals/fig1_time_port.png")
+
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of police time consumed over the year, defined as time spent on the scene, for each call category.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig1_time_tos.png")
+```
+
+\newpage
+
+Figure 2 visualises the disproportionately in time spent on calls explicitly using Lorenz curves for each demand classification.
+These can be compared against the reference line of perfect equality.
+Here, the line of perfect equality represents a situation in which officer time is distributed evenly across all calls.
+In reality, we can see that this is not the case: the Lorenz curves for each demand classification deviate considerably from perfect equality.
+The curves for each classification are almost indistinguishable, so we focus on the curve for all calls (in blue).
+This shows that 50% of the total time officers spent on the scene during 2019 was consumed by just 15% of calls.
+25% of time on scene was consumed by just 5% of calls.
+In other words, a small amount of emergency calls for service account for disproportionately large amount of police resources out on the street.
+This finding is consistent (and comparable) across all six classifications of demand.
+
+```{r fig2, fig.align="center", fig.cap = "Lorenz curve for time spent on scene across all calls (in blue). Lorenz curves in grey for each of the six demand classifications."}
+knitr::include_graphics("visuals/fig2_lorenz_demands.png")
+```
+
+## Spatio-temporal patterning
+
+Mean incident counts by day of the week and by hour, for each demand classification, are visualized in Figure 3.
+This demonstrates that the different sources of public demand for police services, as indicated by the six demand classifications, can have a distinct temporal patterning.
+Community calls appear to concentrate on weekdays, around lunchtime, when other classes of demand are low.
+Crime and quality of life-related calls both appear to occur in the evening and early hours of the morning.
+Traffic incidents concentrate in the evenings, particularly on a Friday.
+Proactive policing does not have a distinct hourly patterning, but this could be attributed to the low hourly counts.
+Calls relating to health issues, the primary interest of @ratcliffe2021policing, appear to largely take place in the evening and early morning.
+
+The spatial patterning of police demand in Detroit, defined as the aggregate time spent on scene throughout the year, is visualized for each demand classification in Figure 4.^[Noting that the aggregate time spent on scene may be influenced by outliers, the raw count maps and underlying distributions are reported in the Appendix.
+These indicate that 'high time' outliers do not misrepresent the aggregate figures in Figure 4.] This demonstrates that the time police spent attending to incidents is highly concentrated in space, and can vary between different demand types.
+The police spend a considerable amount of time around the Wayne State University (WSU) and Midtown attending to community, proactive and quality of life calls for service.
+Officers spent vast amounts of time on crime-related calls in or near hospitals, most notably, the Henry Ford Hospital.
+Police time spent attending to health-related calls concentrates in and around WSU and Midtown, but also among grid cells containing mental health and well-being facilities outside of the city center.
+Time on traffic incidents have a distinct spatial patterning along arterial roads into Downtown.
+
+\newpage
+```{r fig3, fig.align = "center", fig.cap = "Mean call counts by day and hour, for each demand type.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig3_time_heat_tos.png")
+```
+
+```{r fig4, fig.align = "center", fig.cap = "Spatial patterning of the aggregate time spent on scene (in hours) for each demand type.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig4_maps_tos.png")
 ```
@@
 
-## Spatio-temporal patterning
-
-The mean incident counts by day of the week and by hour, for each demand classification, are visualised in Figure 2.
-This demonstrates that the different sources of public demand for police services, as indicated by the six demand classifications, can have distinct temporal patterning.
-Community calls appear to concentrate on weekdays, around lunchtime, when other classes of demand are low.
-Crime and quality of life-related calls both appear to occur in the evening and early hours of the morning, although crime incidents are more impacted by weekend activity stretching into Saturday and Sunday mornings.
-Traffic incidents concentrate in the evenings, particularly on a Friday.
-Proactive policing does not have a distinct hourly patterning, but this could be attributed to the low hourly counts.
-Calls relating to health issues, the primary interest of @ratcliffe2021policing, appear to largely take place in the evening and early morning, aligning with findings from Philadelphia.
-
-Individual incident location coordinates were aggregated to synthetic grid cells laid over the city to generate heatmaps of counts throughout the year.
-^[Grid cells were defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localised variation and interpretable visualisations.
-2% of incidents were excluded from the spatial visualisations due to incomplete geographic information.] These are visualised for each demand classification in Figure 3.
-Calls about community issues concentrate overwhelmingly in and around the Wayne State University (WSU) and the Cultural Center.
-Crime calls tend to cluster in hotspots near hospitals, most notably, the Henry Ford Hospital.
-Proactive call activity originates from the WSU area and Midtown, as does quality of life calls, which are also spread around residential areas.
-Health calls concentrate in Midtown and to the east, in the immediate vicinity of major well-being and mental health centres.
-Traffic incidents have a distinct spatial patterning along arterial roads into downtown.
-
 # Discussion
 
-In replication of recent findings in Philadelphia [@ratcliffe2021policing] this short contribution has sought to describe the scale and composition of calls for police services in Detroit, United States.
-The data and code to reproduce the results reported here are openly available.
-Findings provide robustness to the Philadelphia study: a considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police.
+In replication and extension of the main findings reported by @ratcliffe2021policing and later @lum2021can this contribution has described the scale and composition of calls for police service in Detroit, US.
+Broadly speaking, the results presented here provide robustness to those reported by @ratcliffe2021policing and @lum2021can.
+There is now substantial, recent empirical evidence to demonstrate that the US public rely on the police for both social _and_ crime-fighting functions.
+A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
+This contribution provides the first such demonstration using open data and open code, opening up analyses for scrutiny, replication and extension by the public, academics and/or police practitioners.
+This particularly pertinent given the challenges associated with categorizing individual call descriptions into meaningful demand classifications.
+
+In this study, calls were categorized according the six classes used by @ratcliffe2021policing.
+While concrete comparisons are problematic due to the different array of call descriptions within each demand classifications, the overall composition of police demand in terms of call volume and proportional time consumed are remarkably similar.
+A considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police.
 Instead, the police fulfill a variety of other social service functions, including the resolution of public (mental) health incidents and resolving community issues.
-Findings further highlight the potential implications of radically reforming (defunding) police forces in the United States through [@lum2021can].
+That said, descriptive statistics which drill down into the underlying distribution of officer time consumed by each demand type provide additional insight which has been overlooked in existing studies.
+First, it is clear that there is considerable variation in the amount of time consumed both between and within demand classifications.
+On a methodological point, this highlights how average (e.g., mean times) or aggregate (e.g., sum times) might mask underlying variation.
+Second, and relatedly, these extensions demonstrate that a disproportionately large amount of police time is consumed on a small number of emergency calls.
+50% of total time on scene in consumed by just 15% of calls.
+High-demand calls will likely be complex and/or life-threatening incidents which require numerous officers to be present at the scene, or for extended periods of time.
+A targeted examination of these calls might shed light on the extent to which such high-intensity demands could be reduced or better resolved through or in concert with other public services.
+Targeted interventions on just a handful of calls could drastically reduce the public demand for police services.
 
-\newpage
-```{r fig2, fig.align = "center", fig.cap = "Mean incident counts by day and hour, for each demand type.", out.width = "\\textwidth"}
-knitr::include_graphics("visuals/fig2_time_heat.png")
+The spatial patterning of demand for police services in Detroit--as measured through the aggregate time spent on scene and raw counts--demonstrates that different demands often originate from different places.
+This suggests that the underlying environmental factors generating different forms of police demand are (or can be) distinct.
+Recent research has found that the explanatory frameworks underpinning environmental criminology (e.g., place attractors) are relevant for non-criminal forms of demand, such as those calls involving emotionally disturbed persons [@vaughan2016exploring].
+That said, we know little about the nuances of how different environmental characteristics might drive different forms of demand in the same study area.
+The study materials made available here could be extended using open data resources, such as Open Street Map, to provide a theoretically-driven investigation into the urban character (e.g., place attractors) of high-demand areas, according to the six demand classifications.
+Based on the distinct temporal patterning of different demand types, such analyses should be time-sensitive, capturing, for instance, the ambient population flows of different places throughout the day.
+To date, ambient population analyses have focused on crime-specific calls for service [@andresen2007hot] which as we have seen, constitutes only part of the emergency demands placed on police.
+
+The study is not without shortcomings.
+First, while the Detroit data permits an open and reproducible case site, the raw incident data does not make a distinction between the initial call classification and the final disposition.
+This would permit a comparison, as conducted by @ratcliffe2021policing, into the extent to which the call handler's initial incident classification changes following the attendance of officers at the scene.
+Here, we cannot conduct such an investigation.
+Second, findings presented here, and those reported by @ratcliffe2021policing and @lum2021can, rely on the reliability and validity of police incident classifications.
+These are imperfect.
+Evidence from the United Kingdom, for instance, indicates that police underrecord the proportion of emergency calls involving persons with mental ill-health, which in turn underestimates the amount of police time consumed by such incidents [@langton2021policing].
+Investigations into recording practices and call handler decision-making will be necessary to further understand this issue across US police jurisdictions [see @lum2020constrained; @simpson2021calling].
+
+The findings presented here make two principal contributions.
+First, augmented to those already presented by @ratcliffe2021policing and @lum2021can, results further highlight the potential implications of radically reforming (defunding) police forces in the US.
+There is now substantial, recent empirical evidence to demonstrate that the US public rely on the police for both social _and_ crime-fighting functions.
+A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
+Second, in advance of existing contributions, the materials required to replicate and reproduce these findings are openly available for transparency, scrutiny and extension.
+
+# Appendix
+
+## Raw count maps
+
+```{r fig5app, fig.align = "center", fig.cap = "Spatial patterning of the raw call counts for each demand classification", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig5_maps_counts.png")
 ```
 
-```{r fig3, fig.align = "center", fig.cap = "Spatial patterning of total incident counts for each demand type.", out.width = "\\textwidth"}
-knitr::include_graphics("visuals/fig3_maps.png")
+## Time on scene distributions
+
+```{r fig6app, fig.align = "center", fig.cap = "Distribution and descriptive statistics of time spent on scene by demand classification at the call level.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig6_histogram_mins_tos.png")
 ```
-\newpage
 

### 2022-02-01 scripts/calls_detroit.R

@@ -328,7 +328,7 @@ des_stats_df <- detroit_19_times_agg_df %>%
   ungroup()
 
 # Save.
-write_csv(x = des_stats_df, file = "results/table1_des_stats_total_time.csv")
+write_csv(x = des_stats_df, file = "results/table2_des_stats_total_time.csv")
 
 # Create categorical colour scheme for future use. Colourblind friendly.
 # col_vec <- c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")

### 2022-02-01 scripts/calls_detroit_time_on_scene.R

@@ -331,7 +331,7 @@ des_stats_df <- detroit_19_times_agg_df %>%
   ungroup()
 
 # Save.
-write_csv(x = des_stats_df, file = "results/table1_des_stats_tos.csv")
+write_csv(x = des_stats_df, file = "results/table2_des_stats_tos.csv")
 
 # Create categorical colour scheme for future use. Colourblind friendly.
 # col_vec <- c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")
@@ -601,7 +601,7 @@ means_table_df <- means_df %>%
 # sum(means_df$`Total (hours)`) # 176616
 
 # Save.
-write_csv(x = means_table_df, file = "results/table2_des_stats_tos.csv")
+write_csv(x = means_table_df, file = "results/table3_des_stats_tos.csv")
 
 # Calculate concentration of calls.
 names(detroit19_deploy_clip_sf)
@@ -675,7 +675,7 @@ lorenz_gg <- ggplot() +
   geom_point(data = lorenz_example_df,
              mapping = aes(x = data.cumpercUnit, y = data.cumpercEvents, group = thresh), size = 1.5) +
   geom_line(mapping = aes(x = 0:100, y = 0:100), linetype = "dotted", alpha = 0.3) +
-  annotate(geom = "text", x = 75, y = 79, label = "Line of perfect equality", angle = 43, size = 2, alpha = 0.5) +
+  annotate(geom = "text", x = 75, y = 79, label = "Line of perfect equality", angle = 42.5, size = 2, alpha = 0.5) +
   annotate(geom = "text", x = 15, y = 80, label = "50% of time on scene\n consumed by 15% of calls", size = 2, hjust = 0.5) +
   annotate(geom = "curve" , x = 9, y = 75, xend = 13, yend = 52, size = 0.3, arrow = arrow(length = unit(0.01, "npc")), curvature = 0.3) +
   annotate(geom = "text"  , x = 35 , y = 46, label = "25% of time on scene\n consumed by 5% of calls", size = 2, hjust = 0.5) +

### 2022-02-01 scripts/calls_detroit.R
run thru

@@ -561,7 +561,10 @@ detroit19_deploy_clip_list <- detroit19_deploy_clip_sf %>%
   group_split(type)
 
 # Create list of the duplicate grid sf objects to match. Not an ideal approach but it works.
-grids_list <- rep(list(detroit_grid_sf), 6)
+grids_list <- rep(list(detroit_grid_sf), 6) # run this instead if the below produced length of 18.
+# grids_list <- list(detroit_grid_sf, detroit_grid_sf, detroit_grid_sf,
+#                    detroit_grid_sf, detroit_grid_sf, detroit_grid_sf)
+
 
 # Create point (incidents) to polygon (grids) function.
 p2p_fun <- function(x, y){
@@ -569,9 +572,6 @@ p2p_fun <- function(x, y){
     st_as_sf() %>% 
     mutate(call_count = lengths(st_intersects(x, y)))
 }
-
-# Save workspace.
-# save.image(file = "workspce_total_time.RData")
 
 # Run p2p through lists of incidents and duplicate grids.
 detroit19_grid_list <- map2(grids_list, detroit19_deploy_clip_list, p2p_fun)
@@ -664,3 +664,6 @@ maps_gg <-  plot_grid(plotlist = grid_maps_list,
                       scale = 1.05)
 # Save.
 ggsave(filename = "visuals/fig5_maps_counts.png", height = 48, width = 40, unit = "cm", dpi = 300)
+
+# Save workspace.
+# save.image(file = "workspce_total_time.RData")

### 2022-02-01 scripts/calls_detroit_time_on_scene.R

@@ -749,9 +749,9 @@ histo_comp_gg <- histo_gg +
 ggsave(plot = histo_comp_gg, filename = "visuals/fig6_histogram_mins_tos.png", height = 20, width = 16)
 
 # Create list of the duplicate grid sf objects to match. Not an ideal approach but it works.
-#grids_list <- rep(list(detroit_grid_sf), 6) # run this instead if the below produced length of 18.
-grids_list <- list(detroit_grid_sf, detroit_grid_sf, detroit_grid_sf,
-                   detroit_grid_sf, detroit_grid_sf, detroit_grid_sf)
+grids_list <- rep(list(detroit_grid_sf), 6) # run this instead if the below produced length of 18.
+# grids_list <- list(detroit_grid_sf, detroit_grid_sf, detroit_grid_sf,
+#                    detroit_grid_sf, detroit_grid_sf, detroit_grid_sf)
 
 # Create point (incidents) to polygon (grids) function.
 p2p_fun <- function(x, y){

### 2022-02-02 scripts/calls_detroit_time_on_scene.R
create table for TV

@@ -162,7 +162,7 @@ calldesc2_df <- detroit19_deploy_df %>%
 calldesc2_df %>%
   rename(calldescription_original = calldescription,
          calldescription_new      = calldescription2) %>%
-write_csv(file = "results/categorization_summary.csv")
+  write_csv(file = "results/categorization_summary.csv")
 
 # Aggregate.
 detroit_19_times_df <- detroit19_deploy_df %>% 
@@ -386,6 +386,12 @@ time_ann_gg <- time_gg +
 
 # Save portrait version.
 ggsave(filename = "visuals/fig1_time_tos.png", height = 48, width = 40, unit = "cm", dpi = 300)
+
+# Save list of the 202 raw call descriptions for TV.
+temp <- detroit_19_times_agg_df %>% 
+  filter(type != "unclassified") %>% 
+  left_join(calldesc2_df)
+  
 
 # For further descriptive statistics, we join the new categories back with the raw data.
 detroit19_deploy_df <- detroit19_deploy_df %>% 

### 2022-02-08 paper.Rmd
paper backup

@@
 
+Here, we propose an extension of @ratcliffe2021policing and @lum2021can.
+This extension is motivated by three principal components.
+First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials.
+To date, investigations have been conducted on data which is not publicly available and without the materials (e.g., code) to reproduce the results.
+Yet, open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that real-life decisions are made based on a concrete pool of evidence [@openedp2019].
+In this case.
+the necessity for a fully open investigation into the scale and composition of police demand is perhaps especially important given the public nature of debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
+It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand across different contexts before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
+
+Second, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
+This includes a direct comparison between different measures of 'officer time', namely, total deployment time and time on scene.
+These measures have been reported separately in each existing analyses, so the precise impact of their usage on the same data is unclear.
+We also provide a comprehensive account of the variability in the time consumed by different demand classifications (e.g., crime, community, health).
+These statistics shed light on the degree to which a disproportionately large amount of officer time is consumed by just a small number of calls - insight which can be used to assess how a reduction in demand, or a shift away from calls for police service towards other social services, might free-up officer time.
+In addition, we report on the spatial and temporal patterning of each demand classification.
+Distinctions in these patterns across demand types would suggest similarly distinct demand-generating mechanisms (e.g,, environmental characteristics) and inform deployment decisions based on officer expertise and experience, and/or collaboration with other services (e.g., mental health professionals).
+
 That said, a notable shortcoming of these recent contributions is the lack of openly available data and code to reproduce and replicate findings.
@@
 More broadly, in a sample of papers from top criminology journals, @mcneeley2015replication found that just 2% of papers were replications -- either a direct replication of the same population, or empirical generalizations from a different population.
-The scarcity of replication has been identified as major obstacle to the goals of evidence-based policing, namely, "to develop a rigorous evidence base that can be used by police professionals to develop more efficient and effective police practices, policies, and/or programs" [@openedp2019, p.
+Yet, the practice of open science is fundamental in achieving the goals of evidence-based policing, namely, "to develop a rigorous evidence base that can be used by police professionals to develop more efficient and effective police practices, policies, and/or programs" [@openedp2019, p.
 126].
-In the long-term, open science practices are thought to provide a more concrete base from which police practitioners can make real-life decisions [@openedp2019] and comprise a fundamental component of related fields such as crime science [@laycock2013defining].
+Comparable statements have been made in related fields, such as science [@laycock2013defining].
 It therefore seems appropriate that the investigations of @ratcliffe2021policing and @lum2021can are replicated using open data and methods.
 
-Here, the necessity for a fully open investigation into the scale and composition of police demand is perhaps especially important.
-This is for two reasons.
-First, the implications of redefining the role of the police is considerable.
-It is reasonable for police practitioners and the public to expect a thorough, open and replicable empirical investigation into police demand across numerous contexts before academia concludes that the issue is resolved -- certainly before findings are used to change or introduce police practice or public policy.
+One reason to hesitate over conducting such a replication is that these two existing studies were descriptive accounts of police demand in specific study sites.
+In scientific research more broadly, replications are often motivated by the need to check for Type I or Type II errors using a different sample from the same population and/or adjust for control variables which may have influenced the dependent variable of interest [@gross1997need; see also @mackey2012or].
+We might therefore conclude that a replication of @ratcliffe2021policing and @lum2021can is unwarranted: instead, replications should be reserved only for those studies deploying experimental designs or inferential statistics from a random sample, the conclusions from which will be generalized to a population.
+We disagree.
+There are plentiful reasons to conduct an open descriptive investigation into the scale and composition of police demand.
+This is for two principal reasons.
+
+the descriptive findings from @lum2021can, without the use of random sampling and inferential statistics, are already being generalized to the United States as a whole.
+This is most evident in _The Washington Post_ article [@lumwash].
+
+the implications of redefining the role of the police is considerable.
+
+It is reasonable for police practitioners and the public to expect a thorough, open and replicable empirical investigation into police demand across numerous contexts before academia concludes that the issue is resolved.
+B
+
+certainly before findings are used to change or introduce police practice or public policy.
 As noted, committing to radical reform without such investigations might result in shifts in public demand for services for which there is no suitable supply.
+
 Second, the data handling required to describe the public demand for the police is not straightforward.
@@
 All materials are publicly available (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
-It does so in replication--and where appropriate, extension--of the main findings reported by @ratcliffe2021policing and later @lum2021can.
+It does so in replication, and where appropriate extension, of the main findings reported by @ratcliffe2021policing and later @lum2021can.
 In this way, the contribution serves to test the robustness of existing findings, but also provides the materials required to reproduce and replicate analyses in the interests of transparency and public scrutiny.
@@
 Either identical descriptive statistics are reported in replication of one or both existing studies, or additional analyses are provided for enhanced insight and transparency.
-In sum, the replications and extensions summarised here, coupled with the open code and open data to reproduce findings, constitute the primary contributions of this study.
 

### 2022-02-08 scripts/calls_detroit_time_on_scene.R

@@ -158,7 +158,7 @@ calldesc2_df <- detroit19_deploy_df %>%
   distinct() %>% 
   arrange(calldescription)
 
-# Save for reader reference. No need to save again.
+# Save for reader reference. This is what TV reviewed.
 calldesc2_df %>%
   rename(calldescription_original = calldescription,
          calldescription_new      = calldescription2) %>%
@@ -386,12 +386,6 @@ time_ann_gg <- time_gg +
 
 # Save portrait version.
 ggsave(filename = "visuals/fig1_time_tos.png", height = 48, width = 40, unit = "cm", dpi = 300)
-
-# Save list of the 202 raw call descriptions for TV.
-temp <- detroit_19_times_agg_df %>% 
-  filter(type != "unclassified") %>% 
-  left_join(calldesc2_df)
-  
 
 # For further descriptive statistics, we join the new categories back with the raw data.
 detroit19_deploy_df <- detroit19_deploy_df %>% 

### 2022-02-08 paper.Rmd
paper backup

@@
 To date, investigations have been conducted on data which is not publicly available and without the materials (e.g., code) to reproduce the results.
-Yet, open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that real-life decisions are made based on a concrete pool of evidence [@openedp2019].
+Yet, open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete pool of evidence [@openedp2019].
 In this case.
@@
 
-That said, a notable shortcoming of these recent contributions is the lack of openly available data and code to reproduce and replicate findings.
-This is not an uncommon issue in policing research.
-A recent scoping review of Canadian policing studies concluded that there has been "little effort...
-to replicate or reproduce findings", with just 8% of the literature sampled constituting replications or reproductions [@huey2017replication, p.
-113].
-More broadly, in a sample of papers from top criminology journals, @mcneeley2015replication found that just 2% of papers were replications -- either a direct replication of the same population, or empirical generalizations from a different population.
-Yet, the practice of open science is fundamental in achieving the goals of evidence-based policing, namely, "to develop a rigorous evidence base that can be used by police professionals to develop more efficient and effective police practices, policies, and/or programs" [@openedp2019, p.
-126].
-Comparable statements have been made in related fields, such as science [@laycock2013defining].
-It therefore seems appropriate that the investigations of @ratcliffe2021policing and @lum2021can are replicated using open data and methods.
-
-One reason to hesitate over conducting such a replication is that these two existing studies were descriptive accounts of police demand in specific study sites.
-In scientific research more broadly, replications are often motivated by the need to check for Type I or Type II errors using a different sample from the same population and/or adjust for control variables which may have influenced the dependent variable of interest [@gross1997need; see also @mackey2012or].
-We might therefore conclude that a replication of @ratcliffe2021policing and @lum2021can is unwarranted: instead, replications should be reserved only for those studies deploying experimental designs or inferential statistics from a random sample, the conclusions from which will be generalized to a population.
-We disagree.
-There are plentiful reasons to conduct an open descriptive investigation into the scale and composition of police demand.
-This is for two principal reasons.
-
-the descriptive findings from @lum2021can, without the use of random sampling and inferential statistics, are already being generalized to the United States as a whole.
-This is most evident in _The Washington Post_ article [@lumwash].
-
-the implications of redefining the role of the police is considerable.
-
-It is reasonable for police practitioners and the public to expect a thorough, open and replicable empirical investigation into police demand across numerous contexts before academia concludes that the issue is resolved.
-B
-
-certainly before findings are used to change or introduce police practice or public policy.
-As noted, committing to radical reform without such investigations might result in shifts in public demand for services for which there is no suitable supply.
-
-Second, the data handling required to describe the public demand for the police is not straightforward.
-Reporting the scale and composition of emergency calls for service data necessitates the grouping of incident types through categorization (e.g., 'mental health', 'community').
-As both @ratcliffe2021policing and @lum2021can point out, categorizing incident types is challenging and subject to a degree of interpretation.
-This issue is exacerbated by the fact that different police jurisdictions will use different incident categories.
-Thus, an open replication, inclusive of the categorization decisions, facilitates scrutiny through transparency and provides robustness to existing findings.
-
 With these points in mind, this paper describes the emergency demand for police services in a large city, Detroit, US, using open data and open code.
@@
 It does so in replication, and where appropriate extension, of the main findings reported by @ratcliffe2021policing and later @lum2021can.
-In this way, the contribution serves to test the robustness of existing findings, but also provides the materials required to reproduce and replicate analyses in the interests of transparency and public scrutiny.
+In this way, the contribution serves to test the robustness of existing findings, but also provides additional insight through new analyses, and provides the materials required for reproduction and replication in the interests of transparency and public scrutiny.
 

### 2022-02-09 paper.Rmd
paper update

@@
 ---
-title: 'Describing the scale and composition of calls for police service: a replication and
-  extension using open data'
-author: Samuel Langton\dag, Stijn Ruiter\dag, Tim Verlaan\dag
+title: 'Describing the scale and composition of calls for police service: a replication
+  and extension using open data'
+author: "Samuel Langton\dag, Stijn Ruiter\dag, Tim Verlaan\dag"
 bibliography: refs.bib
@@
   pdf_document:
-    number_sections: true
+    number_sections: yes
 header-includes:
@@
 The remaining public demand from calls involved community issues (10% of calls, 7% of time), medical/public health incidents (7% of calls, 9% of time), proactive policing (8% of calls, 5% of time), quality of life (24% of calls, 14% of time) and traffic duties (11% of calls and 11% of time).
-Comparable results were subsequently reported in the US across multiple anonymised police jurisdictions [@lum2021can; see also @lumwash in _The Washington Post_].
+Comparable results were subsequently reported in the US across multiple anonymized police jurisdictions [@lum2021can; see also @lumwash in _The Washington Post_].
 
@@
 Here, we propose an extension of @ratcliffe2021policing and @lum2021can.
-This extension is motivated by three principal components.
+This extension has three principal components.
 First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials.
@@
 the necessity for a fully open investigation into the scale and composition of police demand is perhaps especially important given the public nature of debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
-It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand across different contexts before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
+It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
 
-Second, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
+Second, and relatedly, the materials provided here can be re-used as a tool to conduct further replications for the purposes of comparison and generalization.
+One way in which our materials facilitate this is through an open classification scheme, namely, the recoding of raw call descriptions into meaningful groups (e.g., 'animal incident'), and the subsequent classification into demand types (e.g., 'quality of life').
+As both @ratcliffe2021policing and @lum2021can point out, categorizing call types is challenging and subject to a degree of interpretation.
+But, without a systematic application of such a classification, we cannot compare jurisdictions, make generalizations, or make iterative improvements to previous studies.
+Thus, we encourage practitioners and academics to use these materials to help build a reliable and valid evidence-base around the scale and composition of demand for police services.
+
+Third, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
 This includes a direct comparison between different measures of 'officer time', namely, total deployment time and time on scene.
@@
 It does so in replication, and where appropriate extension, of the main findings reported by @ratcliffe2021policing and later @lum2021can.
-In this way, the contribution serves to test the robustness of existing findings, but also provides additional insight through new analyses, and provides the materials required for reproduction and replication in the interests of transparency and public scrutiny.
+In this way, the contribution serves to test the robustness of existing findings, provide the materials required for re-use, reproduction and replication, and provides additional insight through new analyses.
 
@@
 
-Recognising that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by public demand, raw counts were augmented with the time spent by officers on each of the six classifications of demand.
+Recognizing that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by public demand, raw counts were augmented with the time spent by officers on each of the six classifications of demand.
 This information permits the calculation of a proportional breakdown of police time consumed by the different demand classifications throughout the year.
 The raw numbers were reported in a table.
-A visual representation of call types nested within each demand classification was summarised in a tree map graphic.
+A visual representation of call types nested within each demand classification was summarized in a tree map graphic.
 The time committed by officers was calculated from when the call was dispatched to an officer to when the officer indicated that the event has been closed.
@@
 
-The spatio-temporal patterning of this demand was summarised in two visuals, focusing specifically on those calls relating to medical/public health.
+The spatio-temporal patterning of this demand was summarized in two visuals, focusing specifically on those calls relating to medical/public health.
 A heatmap showed the mean call volume during hours of the day and days of the week, identifying peak times for demand.
@@
 First, descriptive statistics and a treemap graphic provide an account of the scale and composition of police demand during the year.
-Second, visualizations summarise the spatial and temporal patterning of this demand.
+Second, visualizations summarize the spatial and temporal patterning of this demand.
 
@@
 
-A series of descriptive statistics summarise the time consumed (in minutes) by each demand type.
+A series of descriptive statistics summarize the time consumed (in minutes) by each demand type.
 For comprehensiveness and transparency, this includes the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution.
@@
 
-## Data and methods summary
+<!-- ## Data and methods summary -->
 
-To summarise the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
+<!-- To summarize the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
 Broadly, this demonstrates that, with the exception of the final disposition, this study provides a comprehensive replication of the analyses reported by @ratcliffe2021policing and @lum2021can.
 Either identical descriptive statistics are reported in replication of one or both existing studies, or additional analyses are provided for enhanced insight and transparency.
+-->
 
-\newpage
+<!-- \newpage -->
 
-```{r table1}
-options(knitr.kable.NA = '')
+<!-- ```{r table1} -->
+<!-- options(knitr.kable.NA = '') -->
 
-table1_df <- read_xlsx("results/table1_methods_comparison.xlsx")
+<!-- table1_df <- read_xlsx("results/table1_methods_comparison.xlsx") -->
 
-kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.",
-      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
-    kable_styling(latex_options = c("striped"),
-                  font_size = 10, full_width = F) %>% 
-  column_spec(column = c(2,3), width = c("2cm", "2cm") ) %>% 
-  row_spec(row = 14, hline_after = 14) %>%
-  row_spec(row = 15, italic = T)
-```
+<!-- kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.", -->
+<!--       booktabs = T, longtable = T, linesep = "", format = "latex") %>% -->
+<!--     kable_styling(latex_options = c("striped"), -->
+<!--                   font_size = 10, full_width = F) %>%  -->
+<!--   column_spec(column = c(2,3), width = c("2cm", "2cm") ) %>%  -->
+<!--   row_spec(row = 14, hline_after = 14) %>% -->
+<!--   row_spec(row = 15, italic = T) -->
+<!-- ``` -->
 
@@
 These crime incidents consumed 47% of total deployed time throughout the year, although the figure increases to 51% when considering time on scene in isolation.
+This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.
 The remainder of total time on scene is spent dealing with community issues (5%), public and mental health (7%), proactive policing (4%), quality of life incidents (15%) and traffic duties (16%).
@@
 
-To further quantify the resources consumed, Table 3 summarises the time spent on the scene for each demand classification in minutes.
+To further quantify the resources consumed, Table 3 summarizes the time spent on the scene for each demand classification in minutes.
 A histogram of the underlying distributions are reported in the Appendix.
@@
 
-Figure 2 visualises the disproportionately in time spent on calls explicitly using Lorenz curves for each demand classification.
+Figure 2 visualizes the disproportionately in time spent on calls explicitly using Lorenz curves for each demand classification.
 These can be compared against the reference line of perfect equality.
@@
 In this study, calls were categorized according the six classes used by @ratcliffe2021policing.
-While concrete comparisons are problematic due to the different array of call descriptions within each demand classifications, the overall composition of police demand in terms of call volume and proportional time consumed are remarkably similar.
+The overall composition of police demand in terms of call volume and proportional time consumed are remarkably similar.
 A considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police.
@@
 These are imperfect.
-Evidence from the United Kingdom, for instance, indicates that police underrecord the proportion of emergency calls involving persons with mental ill-health, which in turn underestimates the amount of police time consumed by such incidents [@langton2021policing].
+Evidence from the United Kingdom, for instance, indicates that police under-record the proportion of emergency calls involving persons with mental ill-health, which in turn underestimates the amount of police time consumed by such incidents [@langton2021policing].
 Investigations into recording practices and call handler decision-making will be necessary to further understand this issue across US police jurisdictions [see @lum2020constrained; @simpson2021calling].
 
-The findings presented here make two principal contributions.
-First, augmented to those already presented by @ratcliffe2021policing and @lum2021can, results further highlight the potential implications of radically reforming (defunding) police forces in the US.
-There is now substantial, recent empirical evidence to demonstrate that the US public rely on the police for both social _and_ crime-fighting functions.
+The findings presented here further highlight the potential implications of radically reforming (defunding) police forces in the US.
+In Detroit, there is substantial, recent and reproducible evidence to demonstrate that the US public rely on the police for both social _and_ crime-fighting functions.
 A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
-Second, in advance of existing contributions, the materials required to replicate and reproduce these findings are openly available for transparency, scrutiny and extension.
+The materials required to replicate and reproduce these findings are openly available for transparency, re-use, scrutiny and extension.
 

### 2022-02-09 paper.Rmd
paper update

@@
 ---
-title: 'Describing the scale and composition of calls for police service: a replication
-  and extension using open data'
-author: "Samuel Langton\dag, Stijn Ruiter\dag, Tim Verlaan\dag"
+title: 'Describing the scale and composition of calls for police service: a replication and extension using open data'
+author: Samuel Langton\dag, Stijn Ruiter\dag, Tim Verlaan\dag
 bibliography: refs.bib
@@
   pdf_document: 
-    number_sections: yes
+    number_sections: true
 header-includes:
@@
 csl: apa.csl
-abstract: "This paper describes the scale and composition of emergency public demand for police services in Detroit, United States.
+abstract: "This paper describes the scale and composition of emergency demand for police services in Detroit, United States.
 The contribution is made in replication and extension of analyses reported by Ratcliffe (2021) and Lum et al.
@@
 Findings further highlight the potential implications of radically reforming police forces in the United States.
-In advance of existing studies, the data and code used here are openly available for reproduction, re-use and scrutiny."
+The data and code used here are openly available for reproduction, re-use and scrutiny."
 ---
@@
 Here, we propose an extension of @ratcliffe2021policing and @lum2021can.
-This extension has three principal components.
+This extension has two principal components.
 First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials.
 To date, investigations have been conducted on data which is not publicly available and without the materials (e.g., code) to reproduce the results.
-Yet, open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete pool of evidence [@openedp2019].
-In this case.
-the necessity for a fully open investigation into the scale and composition of police demand is perhaps especially important given the public nature of debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
+Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete pool of evidence [@openedp2019].
+In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
 It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
+Open materials have the additional benefit of being available for re-use in other jurisdictions, helping build a reliable and comparable evidence-base around the scale and composition of police demand across different contexts.
 
-Second, and relatedly, the materials provided here can be re-used as a tool to conduct further replications for the purposes of comparison and generalization.
+<!-- Second, and relatedly, the materials provided here can be re-used as a tool to conduct further replications for the purposes of comparison and generalization.
 One way in which our materials facilitate this is through an open classification scheme, namely, the recoding of raw call descriptions into meaningful groups (e.g., 'animal incident'), and the subsequent classification into demand types (e.g., 'quality of life').
@@
 Thus, we encourage practitioners and academics to use these materials to help build a reliable and valid evidence-base around the scale and composition of demand for police services.
+-->
 
-Third, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
+Second, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
 This includes a direct comparison between different measures of 'officer time', namely, total deployment time and time on scene.
@@
 In addition, we report on the spatial and temporal patterning of each demand classification.
-Distinctions in these patterns across demand types would suggest similarly distinct demand-generating mechanisms (e.g,, environmental characteristics) and inform deployment decisions based on officer expertise and experience, and/or collaboration with other services (e.g., mental health professionals).
+Distinctions in these patterns across demand types would be indicative of similarly distinct demand-generating mechanisms (e.g,, environmental characteristics).
+Such concentrations in time and space might also inform deployment decisions based on officer expertise and experience, and/or collaboration with other services (e.g., mental health professionals).
 
@@
 It does so in replication, and where appropriate extension, of the main findings reported by @ratcliffe2021policing and later @lum2021can.
-In this way, the contribution serves to test the robustness of existing findings, provide the materials required for re-use, reproduction and replication, and provides additional insight through new analyses.
+In this way, the contribution serves to test the robustness of existing findings, provides the materials required for re-use, reproduction and replication, and provides additional insight through new analyses.
 
@@
 Summary statistics will describe the overall composition of demand according to the six demand type classifications.
-This includes call frequency counts, a proportional breakdown of call counts, and a proportional breakdown of police time spent on each demand type.
+This includes call frequency counts, a proportional breakdown of call counts, and a proportional breakdown of police time.
 Here, rather than reporting one measure in isolation, both _total deployed time_ and _time on scene_ measures of police resource are reported.
@@
 
-A series of descriptive statistics summarize the time consumed (in minutes) by each demand type.
-For comprehensiveness and transparency, this includes the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution.
+A series of descriptive statistics summarize the time consumed by each demand type.
+As an aggregate measure, we report the sum time spent on each demand classification by the number of officer shifts.^[Consistent with Ratcliffe (2021) a shift is defined as 7.5 hours.
+The sum total of hours for each demand classification is divided by 7.5.] For comprehensiveness and transparency, we also report the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution in minutes.
 This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.

### 2022-02-10 paper.Rmd
update paper

@@
 To date, investigations have been conducted on data which is not publicly available and without the materials (e.g., code) to reproduce the results.
-Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete pool of evidence [@openedp2019].
+Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete and replicable pool of evidence [@openedp2019].
 In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
@@
 A visual representation of call types nested within each demand classification was summarized in a tree map graphic.
-The time committed by officers was calculated from when the call was dispatched to an officer to when the officer indicated that the event has been closed.
+The time committed by officers was calculated from when the call was dispatched to an officer to when the officer indicated that the event had been closed.
 Here, this is referred to as the _total deployed time_.
-The aggregate time consumed by different demand classifications was reported using a single count of officer shifts (7.5 hours).
+The aggregate time consumed by different demand classifications was reported using a single count of officer shifts (7.5 hours) over the year.
 
@@
 
-Summary statistics will describe the overall composition of demand according to the six demand type classifications.
+Summary statistics describe the overall composition of demand according to the six demand type classifications.
 This includes call frequency counts, a proportional breakdown of call counts, and a proportional breakdown of police time.
-Here, rather than reporting one measure in isolation, both _total deployed time_ and _time on scene_ measures of police resource are reported.
+Both _total deployed time_ and _time on scene_ measures of police resource are reported.
 The proportional breakdown of _time on scene_ is visualized using a treemap graphic for each incident type nested within the demand classifications.
@@
 As an aggregate measure, we report the sum time spent on each demand classification by the number of officer shifts.^[Consistent with Ratcliffe (2021) a shift is defined as 7.5 hours.
-The sum total of hours for each demand classification is divided by 7.5.] For comprehensiveness and transparency, we also report the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution in minutes.
+The sum total of hours for each demand classification over the year is divided by 7.5.] For comprehensiveness and transparency, we also report the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution in minutes.
 This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.
@@
 By way of example, we report on these figures according to arbitrary thresholds of 25% and 50%.
-This sheds light on the degree to which disproportionately large amount of officer time might be consumed by just a small number of calls.
+This sheds light on the degree to which a disproportionately large amount of officer time might be consumed by just a small number of calls.
 
@@
 
-The temporal patterning of all six demand type classifications are visualized using a heatmap of mean call counts, aggregated by day of the week, and hour of the day.
-To summarize the spatial patterning, call location coordinates were aggregated to synthetic grid cells using spatial joins.
-Grid cells were defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localized variation and interpretable visualizations.
-2% of incidents were excluded from the spatial visualizations due to incomplete geographic information.
+The temporal patterning of the six demand type classifications are visualized using a heatmap of mean call counts, aggregated by day of the week, and hour of the day.
+To summarize the spatial patterning, call location coordinates are aggregated to synthetic grid cells using spatial joins.
+Grid cells are defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localized variation and interpretable visualizations.
+2% of incidents are excluded from the spatial visualizations due to incomplete geographic information.
 To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year, for each demand type classification.
 The equivalent raw count maps are available in the Appendix.
-In this, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, but here, we provide a comparison across different demand classifications and two different measures of demand.
+In this way, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, but here, we provide a comparison across different demand classifications and two different measures of demand.
 
-<!-- ## Data and methods summary -->
+## Data and methods summary
 
-<!-- To summarize the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
+To summarize the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
 Broadly, this demonstrates that, with the exception of the final disposition, this study provides a comprehensive replication of the analyses reported by @ratcliffe2021policing and @lum2021can.
 Either identical descriptive statistics are reported in replication of one or both existing studies, or additional analyses are provided for enhanced insight and transparency.
--->
 
-<!-- \newpage -->
+\newpage
 
-<!-- ```{r table1} -->
-<!-- options(knitr.kable.NA = '') -->
+```{r table1}
+options(knitr.kable.NA = '')
 
-<!-- table1_df <- read_xlsx("results/table1_methods_comparison.xlsx") -->
+table1_df <- read_xlsx("results/table1_methods_comparison.xlsx")
 
-<!-- kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.", -->
-<!--       booktabs = T, longtable = T, linesep = "", format = "latex") %>% -->
-<!--     kable_styling(latex_options = c("striped"), -->
-<!--                   font_size = 10, full_width = F) %>%  -->
-<!--   column_spec(column = c(2,3), width = c("2cm", "2cm") ) %>%  -->
-<!--   row_spec(row = 14, hline_after = 14) %>% -->
-<!--   row_spec(row = 15, italic = T) -->
-<!-- ``` -->
+kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.",
+      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
+    kable_styling(latex_options = c("striped"),
+                  font_size = 10, full_width = F) %>%
+  column_spec(column = c(2,3,4), width = c("2cm", "2cm", "2cm") ) %>%
+  row_spec(row = 14, hline_after = 14) %>%
+  row_spec(row = 15, italic = T)
+```
 
@@
 The spatial patterning of police demand in Detroit, defined as the aggregate time spent on scene throughout the year, is visualized for each demand classification in Figure 4.^[Noting that the aggregate time spent on scene may be influenced by outliers, the raw count maps and underlying distributions are reported in the Appendix.
-These indicate that 'high time' outliers do not misrepresent the aggregate figures in Figure 4.] This demonstrates that the time police spent attending to incidents is highly concentrated in space, and can vary between different demand types.
+These indicate that 'high time' outliers do not seriously misrepresent the aggregate figures in Figure 4.] This demonstrates that the time police spent attending to incidents is highly concentrated in space, and can vary between different demand types.
 The police spend a considerable amount of time around the Wayne State University (WSU) and Midtown attending to community, proactive and quality of life calls for service.
@@
 Broadly speaking, the results presented here provide robustness to those reported by @ratcliffe2021policing and @lum2021can.
-There is now substantial, recent empirical evidence to demonstrate that the US public rely on the police for both social _and_ crime-fighting functions.
+There is now substantial empirical evidence to demonstrate that, in multiple jurisdictions, the public rely on the police for both social _and_ crime-fighting functions.
 A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
-This contribution provides the first such demonstration using open data and open code, opening up analyses for scrutiny, replication and extension by the public, academics and/or police practitioners.
-This particularly pertinent given the challenges associated with categorizing individual call descriptions into meaningful demand classifications.
+This contribution provides the first such demonstration using open data and open code, opening up analyses for scrutiny, re-use, and extension by the public, academics and/or police practitioners.
+We encourage usage of the materials provided with this study to replicate analyses across other jurisdictions.
+In doing so, we hope to see a consistent and reliable evidence-base across different contexts for the purposes of comparison and generalization.
 
@@
 Instead, the police fulfill a variety of other social service functions, including the resolution of public (mental) health incidents and resolving community issues.
-That said, descriptive statistics which drill down into the underlying distribution of officer time consumed by each demand type provide additional insight which has been overlooked in existing studies.
+That said, descriptive statistics which drill down into the underlying distribution of officer time consumed by each demand type provide additional insight which would otherwise be overlooked.
 First, it is clear that there is considerable variation in the amount of time consumed both between and within demand classifications.
@@
 The spatial patterning of demand for police services in Detroit--as measured through the aggregate time spent on scene and raw counts--demonstrates that different demands often originate from different places.
-This suggests that the underlying environmental factors generating different forms of police demand are (or can be) distinct.
+Given the specialized training and experience required to deal with certain forms of police demand (e.g., well being checks, suicide threats), forces might consider accounting for this variation when assigning officers to shifts and patrol areas, or organizing co-patrols with other social service professionals.
+This might lead to more favourable outcomes for the public and less strain on officers.
+The spatio-temporal findings also suggests that the underlying environmental factors generating different forms of police demand are (or can be) distinct.
 Recent research has found that the explanatory frameworks underpinning environmental criminology (e.g., place attractors) are relevant for non-criminal forms of demand, such as those calls involving emotionally disturbed persons [@vaughan2016exploring].
 That said, we know little about the nuances of how different environmental characteristics might drive different forms of demand in the same study area.
-The study materials made available here could be extended using open data resources, such as Open Street Map, to provide a theoretically-driven investigation into the urban character (e.g., place attractors) of high-demand areas, according to the six demand classifications.
+The study materials made available here could be extended using open data resources, such as Open Street Map, to provide a theoretically-driven investigation into the urban character (e.g., place attractors) of high-demand areas according to the six demand classifications.
 Based on the distinct temporal patterning of different demand types, such analyses should be time-sensitive, capturing, for instance, the ambient population flows of different places throughout the day.
@@
 Investigations into recording practices and call handler decision-making will be necessary to further understand this issue across US police jurisdictions [see @lum2020constrained; @simpson2021calling].
-
-The findings presented here further highlight the potential implications of radically reforming (defunding) police forces in the US.
-In Detroit, there is substantial, recent and reproducible evidence to demonstrate that the US public rely on the police for both social _and_ crime-fighting functions.
-A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
-The materials required to replicate and reproduce these findings are openly available for transparency, re-use, scrutiny and extension.
 

### 2022-02-10 paper.Rmd
Edited until Detroit data

@@
 title: 'Describing the scale and composition of calls for police service: a replication and extension using open data'
-author: Samuel Langton\dag, Stijn Ruiter\dag, Tim Verlaan\dag
+author: Samuel Langton\dag, Stijn Ruiter\dag\*, Tim Verlaan\dag
 bibliography: refs.bib
@@
 Just 47% of total deployed time responding to 911 calls is consumed by crime incidents.
-The remainder is spent on community issues (6%), health (7%), proactive (5%), quality of life (19%) and traffic duties (13%).
+The remainder is spent on community (6%), health (7%), proactive (5%), quality of life (19%), and traffic (13%) duties.
 A small number of calls consume a disproportionately large amount of police officer time.
@@
 
-Keywords: police, calls for service, 911, demand, reactive.
+Keywords: police, calls for service, 911, demand
 
@@
 
-\dag Netherlands Institute for the Study of Crime and Law Enforcement, Amsterdam, Netherlands. \linebreak \linebreak \linebreak \linebreak
+\dag Netherlands Institute for the Study of Crime and Law Enforcement (NSCR), Amsterdam, Netherlands. \linebreak
+\* Department of Sociology, Utrecht University, Utrecht, Netherlands \linebreak \linebreak \linebreak \linebreak
 
@@
 
-**Disclosure statement** The author declares that they have no competing interests.
+**Disclosure statement** The authors declare that they have no competing interests.
 
-**Funding details** To be completed.
+**Funding details** This research is part of the NSCR research program _What Works in Policing. Towards	Evidence-Based Policing in the Netherlands_, which has been funded by Netherlands Police.
 
@@
 The remaining public demand from calls involved community issues (10% of calls, 7% of time), medical/public health incidents (7% of calls, 9% of time), proactive policing (8% of calls, 5% of time), quality of life (24% of calls, 14% of time) and traffic duties (11% of calls and 11% of time).
-Comparable results were subsequently reported in the US across multiple anonymized police jurisdictions [@lum2021can; see also @lumwash in _The Washington Post_].
+Comparable results were subsequently reported across multiple anonymized police jurisdictions in the US [@lum2021can; see also @lumwash in _The Washington Post_].
 
@@
 First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials.
-To date, investigations have been conducted on data which is not publicly available and without the materials (e.g., code) to reproduce the results.
+To date, investigations have been conducted on data which are not publicly available and without the materials (e.g., code) to reproduce the results.
 Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete and replicable pool of evidence [@openedp2019].
-In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
+In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of the debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
 It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
@@
 This includes a direct comparison between different measures of 'officer time', namely, total deployment time and time on scene.
-These measures have been reported separately in each existing analyses, so the precise impact of their usage on the same data is unclear.
+These measures have been reported separately in the existing analyses, so the precise impact of their usage on the same data is unclear.
 We also provide a comprehensive account of the variability in the time consumed by different demand classifications (e.g., crime, community, health).
@@
 In addition, we report on the spatial and temporal patterning of each demand classification.
-Distinctions in these patterns across demand types would be indicative of similarly distinct demand-generating mechanisms (e.g,, environmental characteristics).
+Distinctions in these patterns across demand types would be indicative of similarly distinct demand-generating mechanisms (e.g., built environment and sociodemographic backdrop).
 Such concentrations in time and space might also inform deployment decisions based on officer expertise and experience, and/or collaboration with other services (e.g., mental health professionals).
@@
 A heatmap showed the mean call volume during hours of the day and days of the week, identifying peak times for demand.
-A geographic map visualized the spatial patterning of these call count concentrations using kernel density in order to identify hotspots of police demand involving medical/public health issues.
-To supplement these findings, Ratcliffe also reported on the differences between the initial CAD incident type classification and the final disposition using a sankey diagram -- again focusing specifically on medical and public health calls.
+A geographic map visualized the spatial patterning of these call count concentrations using kernel density estimation in order to identify hotspots of police demand involving medical/public health issues.
+To supplement these findings, Ratcliffe also reported on the differences between the initial CAD incident type classification and the final disposition using a Sankey diagram -- again focusing specifically on medical and public health calls.
 

### 2022-02-10 paper.Rmd
Edited until Results section

@@
 
-The data used for this study covers the city of Detroit in Michigan, US.
-The City of Detroit authority publish 911 emergency calls for service data through their Open Data Portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with @ratcliffe2021policing the data used here is subset for the year 2019 and excludes those calls initiated by an officer in order to capture public-initiated calls.
+The data used for this study cover the city of Detroit in Michigan, US.
+The City of Detroit authority publish 911 emergency calls for service data through their Open Data Portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with @ratcliffe2021policing, the data used here are subset for the year 2019 and exclude those calls initiated by an officer in order to capture public-initiated calls.
 Each individual incident has a corresponding timestamp and location coordinates to define the call in time and space.
@@
 Calls which had a _total deployed time_ (response time + time on scene) or _time on scene_ of zero minutes, negative values or missing were excluded.
-The data permits a distinction between the time spent resolving incidents with and without the response time.
+The data permit a distinction between the time spent resolving incidents with and without the response time.
 As noted, @lum2021can considered the _time on scene_ measure to be preferable.
@@
 As an aggregate measure, we report the sum time spent on each demand classification by the number of officer shifts.^[Consistent with Ratcliffe (2021) a shift is defined as 7.5 hours.
-The sum total of hours for each demand classification over the year is divided by 7.5.] For comprehensiveness and transparency, we also report the mean, median, standard deviation, range, and in the Appendix, a histogram of the underlying distribution in minutes.
+The sum total of hours for each demand classification over the year is divided by 7.5.] For comprehensiveness and transparency, we also report the mean, median, standard deviation, and range in the Appendix, next to a histogram of the underlying distribution in minutes.
 This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.
@@
 Lorenz curves plot the cumulative distribution of a variable (in this case, time on scene) against the cumulative distribution of a unit (in this case, calls) [@bernasco2017more].
-By way of example, we report on these figures according to arbitrary thresholds of 25% and 50%.
+By way of example, we report on these figures according to arbitrary thresholds of 25% and 50% of time spent.
 This sheds light on the degree to which a disproportionately large amount of officer time might be consumed by just a small number of calls.

### 2022-02-10 paper.Rmd
Completely edited

@@
 This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.
-The remainder of total time on scene is spent dealing with community issues (5%), public and mental health (7%), proactive policing (4%), quality of life incidents (15%) and traffic duties (16%).
+The remainder of total time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
 
@@
 
-kable(table2_df, caption = "Breakdown of frequency counts and time consumed for each demand type. Total deployed time is the sum of the time on scene and response time.",
+kable(table2_df, caption = "Breakdown of frequency counts and time spent on each demand type. Total deployed time is the sum of the time on scene and response time.",
       booktabs = T, longtable = T, linesep = "", format = "latex") %>%
@@
 
-The proportional breakdown of time spent on scene for each incident type, grouped by the six demand classifications, is visualized in Figure 1.^[The purpose of the graphic is to provide an broad summary, in replication of Ratcliffe's visual.
+The proportional breakdown of time spent on scene for each incident type, grouped by the six demand classifications, is visualized in Figure 1.^[The purpose of the graphic is to provide a broad summary, in replication of Ratcliffe's visual.
 The numbers underlying the graphic are reported in the corresponding GitHub repository.] This figure demonstrates the diversity and complexity of public demand for police services in Detroit.
@@
 To further quantify the resources consumed, Table 3 summarizes the time spent on the scene for each demand classification in minutes.
-A histogram of the underlying distributions are reported in the Appendix.
+Histograms of the underlying distributions are reported in the Appendix.
 Table 3 demonstrates a number of things.
@@
 
-Figure 2 visualizes the disproportionately in time spent on calls explicitly using Lorenz curves for each demand classification.
+Figure 2 visualizes the disproportionate share of time spent on a small number of calls explicitly using Lorenz curves for each demand classification.
 These can be compared against the reference line of perfect equality.
@@
 The curves for each classification are almost indistinguishable, so we focus on the curve for all calls (in blue).
-This shows that 50% of the total time officers spent on the scene during 2019 was consumed by just 15% of calls.
-25% of time on scene was consumed by just 5% of calls.
-In other words, a small amount of emergency calls for service account for disproportionately large amount of police resources out on the street.
+This shows that 50% of the total time officers spent on the scene during 2019 was consumed by just 15% of calls, and 25% of time on scene was consumed by just 5% of all calls.
+In other words, a small amount of emergency calls for service account for a disproportionately large amount of police resources out on the street.
 This finding is consistent (and comparable) across all six classifications of demand.
@@
 
-Mean incident counts by day of the week and by hour, for each demand classification, are visualized in Figure 3.
-This demonstrates that the different sources of public demand for police services, as indicated by the six demand classifications, can have a distinct temporal patterning.
+Mean incident counts by day of the week and by hour of day, for each demand classification, are visualized in Figure 3.
+This demonstrates that the different types of public demand for police services, as indicated by the six demand classifications, can have a distinct temporal patterning.
 Community calls appear to concentrate on weekdays, around lunchtime, when other classes of demand are low.
@@
 The police spend a considerable amount of time around the Wayne State University (WSU) and Midtown attending to community, proactive and quality of life calls for service.
-Officers spent vast amounts of time on crime-related calls in or near hospitals, most notably, the Henry Ford Hospital.
+Officers spend vast amounts of time on crime-related calls in or near hospitals, most notably, the Henry Ford Hospital.
 Police time spent attending to health-related calls concentrates in and around WSU and Midtown, but also among grid cells containing mental health and well-being facilities outside of the city center.
@@
 
-In replication and extension of the main findings reported by @ratcliffe2021policing and later @lum2021can this contribution has described the scale and composition of calls for police service in Detroit, US.
+In replication and extension of the main findings reported by @ratcliffe2021policing and @lum2021can this contribution has described the scale and composition of calls for police service in Detroit, US.
 Broadly speaking, the results presented here provide robustness to those reported by @ratcliffe2021policing and @lum2021can.
@@
 In this study, calls were categorized according the six classes used by @ratcliffe2021policing.
-The overall composition of police demand in terms of call volume and proportional time consumed are remarkably similar.
+The overall composition of police demand in terms of call volume and proportional time spent are remarkably similar.
 A considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police.
@@
 On a methodological point, this highlights how average (e.g., mean times) or aggregate (e.g., sum times) might mask underlying variation.
-Second, and relatedly, these extensions demonstrate that a disproportionately large amount of police time is consumed on a small number of emergency calls.
+Second, and relatedly, these extensions demonstrate that a disproportionately large amount of police time is spent on a small number of emergency calls.
 50% of total time on scene in consumed by just 15% of calls.
@@
 Given the specialized training and experience required to deal with certain forms of police demand (e.g., well being checks, suicide threats), forces might consider accounting for this variation when assigning officers to shifts and patrol areas, or organizing co-patrols with other social service professionals.
-This might lead to more favourable outcomes for the public and less strain on officers.
+This might lead to more favorable outcomes for the public and less strain on officers.
 The spatio-temporal findings also suggests that the underlying environmental factors generating different forms of police demand are (or can be) distinct.
@@
 Based on the distinct temporal patterning of different demand types, such analyses should be time-sensitive, capturing, for instance, the ambient population flows of different places throughout the day.
-To date, ambient population analyses have focused on crime-specific calls for service [@andresen2007hot] which as we have seen, constitutes only part of the emergency demands placed on police.
+To date, ambient population analyses have focused on crime-specific calls for service [@andresen2007hot] which as we have seen, constitute only part of the emergency demands placed on police.
 
 The study is not without shortcomings.
-First, while the Detroit data permits an open and reproducible case site, the raw incident data does not make a distinction between the initial call classification and the final disposition.
+First, while the Detroit data permits an open and reproducible case site, the raw incident data do not make a distinction between the initial call classification and the final disposition.
 This would permit a comparison, as conducted by @ratcliffe2021policing, into the extent to which the call handler's initial incident classification changes following the attendance of officers at the scene.

### 2022-02-11 paper.Rmd
update discussion
few changes + seperate feekback
feedback

@@
 It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
+Especially since classifying police demand into these different categories is no straightforward or facile task.
+The difference in categories in Ratcliffe's and Lum's paper attest to this and showcase why they should not have to be taken at face-value.
 Open materials have the additional benefit of being available for re-use in other jurisdictions, helping build a reliable and comparable evidence-base around the scale and composition of police demand across different contexts.
@@
 The equivalent raw count maps are available in the Appendix.
-In this way, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, but here, we provide a comparison across different demand classifications and two different measures of demand.
+In this way, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, while adding a comparison across different demand classifications and two different measures of demand.
 
@@
 A targeted examination of these calls might shed light on the extent to which such high-intensity demands could be reduced or better resolved through or in concert with other public services.
-Targeted interventions on just a handful of calls could drastically reduce the public demand for police services.
 

### 2022-02-14 paper.Rmd
time updates

--- 2022-02-10 paper.Rmd
@@
 
+\begin{flushleft}
+
 \dag Netherlands Institute for the Study of Crime and Law Enforcement (NSCR), Amsterdam, Netherlands. \linebreak
-\* Department of Sociology, Utrecht University, Utrecht, Netherlands \linebreak \linebreak \linebreak \linebreak
+*Department of Sociology, Utrecht University, Utrecht, Netherlands \linebreak \linebreak \linebreak \linebreak
+
+\end{flushleft}
 
@@
 Second, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
-This includes a direct comparison between different measures of 'officer time', namely, total deployment time and time on scene.
-These measures have been reported separately in the existing analyses, so the precise impact of their usage on the same data is unclear.
+This includes a comparison between measures of 'officer time' with and without the response time.
+While concerns have been raised about the inclusion of the response time in measures of officer time, largely due to low-priority calls not necessitating a rapid response, the precise impact of their usage on the same data is unclear.
 We also provide a comprehensive account of the variability in the time consumed by different demand classifications (e.g., crime, community, health).
@@
 In addition, we report on the spatial and temporal patterning of each demand classification.
-Distinctions in these patterns across demand types would be indicative of similarly distinct demand-generating mechanisms (e.g., built environment and sociodemographic backdrop).
+Distinctions in these patterns across demand types would be indicative of similarly distinct demand-generating mechanisms (e.g., built environment).
 Such concentrations in time and space might also inform deployment decisions based on officer expertise and experience, and/or collaboration with other services (e.g., mental health professionals).
@@
 The time committed by officers was calculated from when the call was dispatched to an officer to when the officer indicated that the event had been closed.
-Here, this is referred to as the _total deployed time_.
-The aggregate time consumed by different demand classifications was reported using a single count of officer shifts (7.5 hours) over the year.
+Ratcliffe accounted for the number of officers attending each incident when calculating the amount of time consumed.
+The aggregate time consumed by different demand classifications was reported using a single count of officer shifts over the year.
 
@@
 No descriptive statistics on the range or spread of these figures was reported.
-The 'time spent' measure used was the _time on scene_.
-The authors considered this to be preferable to _total deployed time_ when measuring police demands, since low priority calls may have a non-urgent (slow) response due to the nature of the call, and officers may be reassigned to other duties while on an assignment.
+The 'time spent' measure used was the _time on scene_ (i.e., excluding response time).
+The authors considered this to be preferable, since low priority calls may have a non-urgent (slow) response due to the nature of the call, and officers may be reassigned to other duties while on an assignment.
+Lum and colleagues calculated _time on scene_ as the time from the first officer's arrival to when the call was reported complete.
+In this way, the calculation did not account for multi-officer responses (e.g., back-up).
+Instead, the authors calculated _time on scene_ as the time from the first officer's arrival to when the call was reported complete.
+This was justified for a number of reasons, among them, avoiding overinflation of time due to officers 'flocking' to non-serious incidents, and limitations with the CAD data.
 A proportional breakdown of the disposition (assist only, report written or arrest citation) was provided for each of the fourteen call types.
@@
 The data used for this study cover the city of Detroit in Michigan, US.
-The City of Detroit authority publish 911 emergency calls for service data through their Open Data Portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with @ratcliffe2021policing, the data used here are subset for the year 2019 and exclude those calls initiated by an officer in order to capture public-initiated calls.
+The City of Detroit authority publish 911 emergency calls for service data through their Open Data Portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with the two existing studies, the data used here exclude those calls initiated by an officer in order to capture public-initiated calls.
 Each individual incident has a corresponding timestamp and location coordinates to define the call in time and space.
+Data was subset for the year 2019.
 
 For each incident, the police response time and time on the scene are reported.
-In sum, these comprise the total deployed time officers spend on the incident.
-Calls which had a _total deployed time_ (response time + time on scene) or _time on scene_ of zero minutes, negative values or missing were excluded.
-The data permit a distinction between the time spent resolving incidents with and without the response time.
+The former measures the response time of the first officer on-scene.
+The latter mimics the measure used by @lum2021can, namely, the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
+Here, we refer to the sum of the response time and time on scene as the _total deployed time_.
+Calls which had a _total deployed time_ or _time on scene_ of zero minutes, negative values or missing were excluded.
+The data permit a distinction between the time spent with and without the response time.
 As noted, @lum2021can considered the _time on scene_ measure to be preferable.
-Here, for completeness, both measurements are reported.
+Here, for completeness, we report times with and without the response time.
 
@@
 
-A series of descriptive statistics summarize the time consumed by each demand type.
-As an aggregate measure, we report the sum time spent on each demand classification by the number of officer shifts.^[Consistent with Ratcliffe (2021) a shift is defined as 7.5 hours.
-The sum total of hours for each demand classification over the year is divided by 7.5.] For comprehensiveness and transparency, we also report the mean, median, standard deviation, and range in the Appendix, next to a histogram of the underlying distribution in minutes.
+A series of descriptive statistics summarize the time on scene consumed by each demand type.
+For comprehensiveness and transparency, we also report the mean, median, standard deviation, and range in the Appendix, next to a histogram of the underlying distribution in minutes.
 This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.
@@
 This sheds light on the degree to which a disproportionately large amount of officer time might be consumed by just a small number of calls.
+
+<!-- As an aggregate measure, we report the sum time spent on each demand classification by the number of officer shifts.^[Consistent with Ratcliffe (2021) a shift is defined as 7.5 hours.
+The sum total of hours for each demand classification over the year is divided by 7.5.]  -->
 
@@
 
-## Data and methods summary
+<!-- ## Data and methods summary -->
 
-To summarize the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
-Broadly, this demonstrates that, with the exception of the final disposition, this study provides a comprehensive replication of the analyses reported by @ratcliffe2021policing and @lum2021can.
+<!-- To summarize the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
+Broadly, this demonstrates that this study provides a comprehensive replication of the analyses reported by @ratcliffe2021policing and @lum2021can.
 Either identical descriptive statistics are reported in replication of one or both existing studies, or additional analyses are provided for enhanced insight and transparency.
+Here, we note two exceptions: we do not report an aggregate summary (total shifts) because our time measure does not  -->
 
-\newpage
+<!-- \newpage -->
 
-```{r table1}
-options(knitr.kable.NA = '')
+<!-- ```{r table1} -->
+<!-- options(knitr.kable.NA = '') -->
 
-table1_df <- read_xlsx("results/table1_methods_comparison.xlsx")
+<!-- table1_df <- read_xlsx("results/table1_methods_comparison.xlsx") -->
 
-kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.",
-      booktabs = T, longtable = T, linesep = "", format = "latex") %>%
-    kable_styling(latex_options = c("striped"),
-                  font_size = 10, full_width = F) %>%
-  column_spec(column = c(2,3,4), width = c("2cm", "2cm", "2cm") ) %>%
-  row_spec(row = 14, hline_after = 14) %>%
-  row_spec(row = 15, italic = T)
-```
+<!-- kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.", -->
+<!--       booktabs = T, longtable = T, linesep = "", format = "latex") %>% -->
+<!--     kable_styling(latex_options = c("striped"), -->
+<!--                   font_size = 10, full_width = F) %>% -->
+<!--   column_spec(column = c(2,3,4), width = c("2cm", "2cm", "2cm") ) %>% -->
+<!--   row_spec(row = 14, hline_after = 14) %>% -->
+<!--   row_spec(row = 15, italic = T) -->
+<!-- ``` -->
 
@@
 
-The frequency counts and proportional breakdowns for each demand type classification are reported in Table 2.
+The frequency counts and proportional breakdowns for each demand type classification are reported in Table 1.
 This demonstrates that in Detroit, a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
@@
 These crime incidents consumed 47% of total deployed time throughout the year, although the figure increases to 51% when considering time on scene in isolation.
-This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.
-The remainder of total time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
+This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.^[Crime calls for service had the highest proportion of Priority 1 (highest priority) responses at 52%, followed by traffic (24%), quality of life (17%), community (16%), health (14%), and proactive (3%).] The remainder of total time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
 
-```{r table2}
-table2a_df <- read_csv("results/table2_des_stats_total_time.csv")
-table2b_df <- read_csv("results/table2_des_stats_tos.csv")
+```{r table1}
+table1a_df <- read_csv("results/table1_des_stats_total_time.csv")
+table1b_df <- read_csv("results/table1_des_stats_tos.csv")
 
-table2_df <- table2a_df %>% 
+table1_df <- table1a_df %>% 
   select(-`Demand type`, -Count, -`Count (%)`) %>%  
-  bind_cols(table2b_df) %>% 
-  select(`Demand type`, Count, `Count (%)`,
-         `Time on scene (%)`, `Total deployed time (%)`,  `Total deployed time (shifts)`) %>%  # reorder for final table.
-  rename( `Total deployed \n time (shifts)` =  `Total deployed time (shifts)`)
+  bind_cols(table1b_df) %>% 
+  select(`Demand type`, Count, `Count (%)`, # ,  `Total deployed time (shifts)`
+         `Time on scene (%)`, `Total deployed time (%)`) #%>%  reorder for final table.
+  # rename( `Total deployed \n time (shifts)` =  `Total deployed time (shifts)`)
 
-kable(table2_df, caption = "Breakdown of frequency counts and time spent on each demand type. Total deployed time is the sum of the time on scene and response time.",
+kable(table1_df, caption = "Breakdown of frequency counts and time spent on each demand type. Total deployed time is the sum of the time on scene and response time.",
       booktabs = T, longtable = T, linesep = "", format = "latex") %>%
@@
 
-To further quantify the resources consumed, Table 3 summarizes the time spent on the scene for each demand classification in minutes.
+To further quantify the resources consumed, Table 2 summarizes the time spent on the scene for each demand classification in minutes.
 Histograms of the underlying distributions are reported in the Appendix.
-Table 3 demonstrates a number of things.
+Table 2 demonstrates a number of things.
 First, there is a modest amount of variation in the typical time at scene of different types of emergency call.
 For instance, the mean time spent on the scene of traffic incidents (55 minutes) and crime (51 minutes) is considerably more that quality of life issues (31 minutes) or proactive activity (28 minutes).
-These differences will be a result of a combination of factors, including the complexity and severity of incidents and the number of officers required to attend the scene.
+These differences will be a result of a combination of factors, including the complexity and severity of incidents.
 Second, there is considerable variation in the time consumed by incidents _within_ each demand type.
-While many calls are resolved in minutes, a handful consume hours of cumulative officer time, resulting in a highly skewed distribution.
+While many calls are resolved in minutes, a handful consume hours of officer time on the scene, resulting in a highly skewed distribution.
 This highlights how mean or aggregate summary statistics of police time as a measure of demand can mask (important) underlying variation.
 
-```{r table3}
-table3_df <- read_csv("results/table3_des_stats_tos.csv")
+```{r table2}
+table2_df <- read_csv("results/table2_des_stats_tos.csv")
 
-kable(table3_df, caption = "Descriptive statistics of minutes spent on scene for each demand type classification.",
+kable(table2_df, caption = "Descriptive statistics of minutes spent on scene for each demand type classification.",
       booktabs = T, longtable = T, linesep = "", format = "latex") %>%
@@
 The curves for each classification are almost indistinguishable, so we focus on the curve for all calls (in blue).
-This shows that 50% of the total time officers spent on the scene during 2019 was consumed by just 15% of calls, and 25% of time on scene was consumed by just 5% of all calls.
+This shows that 50% of the time officers spent on the scene during 2019 was consumed by just 15% of calls, and 25% of time on scene was consumed by just 5% of all calls.
 In other words, a small amount of emergency calls for service account for a disproportionately large amount of police resources out on the street.
@@
 50% of total time on scene in consumed by just 15% of calls.
-High-demand calls will likely be complex and/or life-threatening incidents which require numerous officers to be present at the scene, or for extended periods of time.
+High-demand calls will likely be complex and/or life-threatening incidents or calls which require officers to secure a scene for extended periods of time.
 A targeted examination of these calls might shed light on the extent to which such high-intensity demands could be reduced or better resolved through or in concert with other public services.
-Targeted interventions on just a handful of calls could drastically reduce the public demand for police services.
+
+On this point, this study raises useful discussion points around the different measures of officer time.
+First, it is clear that findings on the scale and composition of police demand are sensitive, albeit only slightly, to whether measures of officer time include the response time.
+The degree to which the two estimates differ varies by demand classification, likely as a result of the different call priorities, and in turn, the rapidness of the response times.
+That said, this claim would need verifying through detailed examination.
+On this point, second, we would reiterate that the 'time on scene' measure reported here, in alignment with @lum2021can, refers to the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
+This differs to that of @ratcliffe2021policing which accounted the number of officers attending the scene.
+@lum2021can note that including back-up officers in the time calculation might introduces complications.
+For instance, non-serious incidents might receive back-up unnecessarily, particularly during periods of low-demand, which would inflate the time consumed and overstate the resources required to resolve such an incident.
+Although the overall composition of police demand is comparable between Philadelphia [@ratcliffe2021policing] and Detroit, the true impact of these measures remains unclear.
+We propose that future research investigates these distinctions.
+This would require CAD (or equivalent) data, timestamped at each stage the response (call-handler dispatch, arrival on scene, scene departure), for individual police units nested within calls.
+This level of granularity would permit the calculation and comparison of multiple measures of officer time, as well as an investigation into the extent to which back-ups and/or 'flocking' might influence such measures.
 

### 2022-02-14 scripts/calls_detroit.R

@@ -322,13 +322,13 @@ des_stats_df <- detroit_19_times_agg_df %>%
   group_by(type) %>% 
   summarise(`Count`      = sum(freq),
             `Count (%)`  = round(sum(prop_count), 2),
-            `Total deployed time (%)`   = round(sum(prop_time ), 2),
-            `Total deployed time (shifts)` = round((sum(sum_time)/60)/7.5, 0) )  %>%
+            `Total deployed time (%)`   = round(sum(prop_time ), 2)) %>% 
+            # `Total deployed time (shifts)` = round((sum(sum_time)/60)/7.5, 0) )  %>%
   rename(`Demand type` = type) %>% 
   ungroup()
 
 # Save.
-write_csv(x = des_stats_df, file = "results/table2_des_stats_total_time.csv")
+write_csv(x = des_stats_df, file = "results/table1_des_stats_total_time.csv")
 
 # Create categorical colour scheme for future use. Colourblind friendly.
 # col_vec <- c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")

### 2022-02-14 scripts/calls_detroit_time_on_scene.R

@@ -320,7 +320,6 @@ sum(is.na(detroit_19_times_agg_df)) # 0
 # How many incidents in total?
 sum(detroit_19_times_agg_df$freq) # 258773
 
-
 # Descriptive stats. 
 des_stats_df <- detroit_19_times_agg_df %>% 
   group_by(type) %>% 
@@ -331,7 +330,26 @@ des_stats_df <- detroit_19_times_agg_df %>%
   ungroup()
 
 # Save.
-write_csv(x = des_stats_df, file = "results/table2_des_stats_tos.csv")
+write_csv(x = des_stats_df, file = "results/table1_des_stats_tos.csv")
+
+# Investigate priorities.
+sum(is.na(detroit19_deploy_df$priority)) # 142 missings.
+
+# Join to get demand types.
+priorities_df <- detroit19_deploy_df %>% 
+  select(incident_id, calldescription2, priority) %>%
+  left_join(detroit_19_times_df)
+
+# Frequenc and prop statistics.
+prior_freq_df <- priorities_df %>% 
+  filter(type != "unclassified") %>% 
+  drop_na(priority) %>% 
+  group_by(type) %>% 
+  count(priority) %>% 
+  mutate(totals = sum(n)) %>%  
+  ungroup() %>% 
+  mutate(props = round(100*n/totals, 2)) %>% 
+  filter(priority == 1)
 
 # Create categorical colour scheme for future use. Colourblind friendly.
 # col_vec <- c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")
@@ -601,7 +619,7 @@ means_table_df <- means_df %>%
 # sum(means_df$`Total (hours)`) # 176616
 
 # Save.
-write_csv(x = means_table_df, file = "results/table3_des_stats_tos.csv")
+write_csv(x = means_table_df, file = "results/table2_des_stats_tos.csv")
 
 # Calculate concentration of calls.
 names(detroit19_deploy_clip_sf)

### 2022-02-15 paper.Rmd
response time measure update

@@
 Second, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
-This includes a comparison between measures of 'officer time' with and without the response time.
+This includes a comparison between measures of officer time with and without the response time.
 While concerns have been raised about the inclusion of the response time in measures of officer time, largely due to low-priority calls not necessitating a rapid response, the precise impact of their usage on the same data is unclear.
@@
 Ratcliffe accounted for the number of officers attending each incident when calculating the amount of time consumed.
-The aggregate time consumed by different demand classifications was reported using a single count of officer shifts over the year.
+The aggregate time consumed by different demand classifications was reported using a single count of officer shifts (7.5 hours) over the year.
 
@@
 No descriptive statistics on the range or spread of these figures was reported.
-The 'time spent' measure used was the _time on scene_ (i.e., excluding response time).
+
+The 'time spent' measure used was the _time on scene_ (i.e., excluding the response time).
 The authors considered this to be preferable, since low priority calls may have a non-urgent (slow) response due to the nature of the call, and officers may be reassigned to other duties while on an assignment.
-Lum and colleagues calculated _time on scene_ as the time from the first officer's arrival to when the call was reported complete.
-In this way, the calculation did not account for multi-officer responses (e.g., back-up).
-Instead, the authors calculated _time on scene_ as the time from the first officer's arrival to when the call was reported complete.
+Lum and colleagues calculated _time on scene_ as the time from the first officer's arrival to when the call was reported as resolved.
+In this way, the calculation did not account for the additional time consumed by multi-officer responses (e.g., back-up).
 This was justified for a number of reasons, among them, avoiding overinflation of time due to officers 'flocking' to non-serious incidents, and limitations with the CAD data.
@@
 For each incident, the police response time and time on the scene are reported.
-The former measures the response time of the first officer on-scene.
-The latter mimics the measure used by @lum2021can, namely, the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
+The former measures the response time from dispatch to when the first officer arrives on-scene.^[In the raw data, response time is referred to as _traveltime_.
+This figure does not include the intake or dispatch time which we did not deem to be deployment.] The latter mimics the measure used by @lum2021can, namely, the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
 Here, we refer to the sum of the response time and time on scene as the _total deployed time_.
-Calls which had a _total deployed time_ or _time on scene_ of zero minutes, negative values or missing were excluded.
+Calls which had a response time of negative values or missing were dropped from analyses.
+Calls which had a _time on scene_ of zero minutes, negative values or missing were also excluded.
 The data permit a distinction between the time spent with and without the response time.
@@
 Each call description was then categorized into the six demand type classifications used in @ratcliffe2021policing, namely, crime, community, medical/public health, proactive, quality of life and traffic.
-2.8% of incidents were deemed unclassifiable.
+2.8% of calls were deemed unclassifiable.
 For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of total deployed time during the year.
@@
 Just 44% of emergency calls for service related to crime incidents.
-These crime incidents consumed 47% of total deployed time throughout the year, although the figure increases to 51% when considering time on scene in isolation.
-This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.^[Crime calls for service had the highest proportion of Priority 1 (highest priority) responses at 52%, followed by traffic (24%), quality of life (17%), community (16%), health (14%), and proactive (3%).] The remainder of total time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
+These crime incidents consume 51% of time on scene throughout the year.
+The remainder of total time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
+The differences in the proportional breakdown between time on scene and total deployment time are marginal.
+
+<!-- These crime incidents consumed 51% of total deployed time throughout the year, although the figure increases to 51% when considering time on scene in isolation.
+This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.^[Crime calls for service had the highest proportion of Priority 1 (highest priority) responses at 52%, followed by traffic (24%), quality of life (17%), community (16%), health (14%), and proactive (3%).] -->
 

### 2022-02-15 scripts/calls_detroit.R

@@ -55,8 +55,13 @@ length(unique(detroit_df$incident_id)) # 419193. Same as the number of rows.
 
 # Remove missings, zeros and negatives.
 detroit19_sub_df <- detroit_df %>% 
-  drop_na(totaltime, time_on_scene) %>%  # note the overlap.
-  filter(totaltime >0, time_on_scene >0)
+  # drop_na(totaltime, time_on_scene) %>%        # note the overlap.
+  drop_na(traveltime, time_on_scene) %>%         # note the overlap.
+  filter(traveltime >=0, time_on_scene >0) %>%   # has to be time on scene, but no travel time is ok.
+  mutate(totaltime = traveltime + time_on_scene) # this overwrite the original measure.
+
+# Check.
+nrow(detroit19_sub_df) # 293025
 
 # First, filter out calldescriptions which are admin, completely unknown or do not appear to
 # involve deployment.
@@ -315,7 +320,7 @@ detroit_19_times_agg_df <- detroit_19_times_df %>%
 sum(is.na(detroit_19_times_agg_df)) # 0
 
 # How many incidents in total?
-sum(detroit_19_times_agg_df$freq) # 258773
+sum(detroit_19_times_agg_df$freq) # 258786
 
 # Descriptive stats. 
 des_stats_df <- detroit_19_times_agg_df %>% 
@@ -340,7 +345,7 @@ write_csv(x = detroit_19_times_agg_df, file = "results/prop_breakdown_total_time
 detroit_19_times_agg_df <- detroit_19_times_agg_df %>% 
   mutate(calldescription2_labs = str_replace_all(string = calldescription2, pattern = " ", replacement = "\n"))
 
-length(unique(detroit_19_times_agg_df$calldescription2)) # 46
+length(unique(detroit_19_times_agg_df$calldescription2)) 
 
 # Create main colour.
 viridis_1 <- viridis::viridis(5)[2]
@@ -406,7 +411,7 @@ mini_df <- detroit19_deploy_df %>%
   slice(1:100)
 
 # Check categories again.
-length(unique(detroit19_deploy_df$calldescription2)) # 46
+length(unique(detroit19_deploy_df$calldescription2)) # 52
 
 # Aggregate by date and hour.
 dh_agg_df <- detroit19_deploy_df %>% 
@@ -484,10 +489,10 @@ detroit_sample_sf <- detroit19_deploy_df %>%
 #   write_csv(file = "data/detroit_sample.csv")
 
[…]

### 2022-02-15 scripts/calls_detroit_time_on_scene.R

@@ -59,8 +59,11 @@ length(unique(detroit_df$incident_id)) # 419193. Same as the number of rows.
 
 # Remove missings, zeros and negatives.
 detroit19_sub_df <- detroit_df %>% 
-  drop_na(totaltime, time_on_scene) %>%  # note the overlap.
-  filter(totaltime >0, time_on_scene >0)
+  # drop_na(totaltime, time_on_scene) %>%        # note the overlap.
+  drop_na(traveltime, time_on_scene) %>%         # note the overlap.
+  filter(traveltime >=0, time_on_scene >0) %>%   # has to be time on scene, but no travel time is ok.
+  mutate(totaltime = traveltime + time_on_scene) # this overwrite the original measure.
+
 
 # First, filter out calldescriptions which are admin, completely unknown or do not appear to
 # involve deployment.
@@ -318,7 +321,7 @@ detroit_19_times_agg_df <- detroit_19_times_df %>%
 sum(is.na(detroit_19_times_agg_df)) # 0
 
 # How many incidents in total?
-sum(detroit_19_times_agg_df$freq) # 258773
+sum(detroit_19_times_agg_df$freq) # 258786
 
 # Descriptive stats. 
 des_stats_df <- detroit_19_times_agg_df %>% 
@@ -517,7 +520,7 @@ detroit_sample_sf <- detroit19_deploy_df %>%
 length(unique(detroit19_deploy_df$calldescription2)) # 50
 
 # Check counts total.
-nrow(detroit19_deploy_df) # 258773
+nrow(detroit19_deploy_df) # 258786
 
 # This makes it clear that many incidents for which the location is not known are geocoded
 # to a specific arbitrary location (-83.111560213, 42.3003668800001).
@@ -586,7 +589,7 @@ detroit_grid_sf <- detroit_grid_sf %>%
 #   geom_sf()
 
 # Check counts used in maps. Lower due to incomplete coordinates.
-sum(detroit19_deploy_clip_sf$n) # 246971
+sum(detroit19_deploy_clip_sf$n) # 246913
 
 # Split incident sf object into list.
 detroit19_deploy_clip_list <- detroit19_deploy_clip_sf %>% 
@@ -905,4 +908,4 @@ maps_gg <-  plot_grid(plotlist = grid_maps_list,
 ggsave(filename = "visuals/fig4_maps_tos.png", height = 48, width = 40, unit = "cm", dpi = 300)
 
 # Save workspace.
[…]

### 2022-02-15 paper.Rmd
update paper

@@
 Findings indicate that police spend a considerable proportion of time performing a social service function.
-Just 47% of total deployed time responding to 911 calls is consumed by crime incidents.
-The remainder is spent on community (6%), health (7%), proactive (5%), quality of life (19%), and traffic (13%) duties.
+Just 51% of total deployed time responding to 911 calls is consumed by crime incidents.
+The remainder is spent on community (5%), health (7%), proactive (4%), quality of life (16%), and traffic (15%) duties.
 A small number of calls consume a disproportionately large amount of police officer time.
@@
 In this way, the calculation did not account for the additional time consumed by multi-officer responses (e.g., back-up).
-This was justified for a number of reasons, among them, avoiding overinflation of time due to officers 'flocking' to non-serious incidents, and limitations with the CAD data.
+This was justified for a number of reasons, among them, avoiding the over-inflation of time estimates due to officers 'flocking' to non-serious incidents, and limitations with the CAD data.
 A proportional breakdown of the disposition (assist only, report written or arrest citation) was provided for each of the fourteen call types.
@@
 For each incident, the police response time and time on the scene are reported.
-The former measures the response time from dispatch to when the first officer arrives on-scene.^[In the raw data, response time is referred to as _traveltime_.
-This figure does not include the intake or dispatch time which we did not deem to be deployment.] The latter mimics the measure used by @lum2021can, namely, the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
-Here, we refer to the sum of the response time and time on scene as the _total deployed time_.
+The former measures the time from dispatch to when the first officer arrives on-scene.^[In the raw data, the time between dispatch and first on-scene arrival is referred to as _travel time_.
+The _intake_ and _dispatch time_ are reported, however, we did not consider these measures to reflect deployment time.
+We return to this point in the discussion.] The latter mimics the measure used by @lum2021can, namely, the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
+Here, the data permit a distinction between the time spent on calls with and without the response time.
+We refer to the sum of the response time and time on scene as the _total deployed time_.
 Calls which had a response time of negative values or missing were dropped from analyses.
 Calls which had a _time on scene_ of zero minutes, negative values or missing were also excluded.
-The data permit a distinction between the time spent with and without the response time.
 As noted, @lum2021can considered the _time on scene_ measure to be preferable.
-Here, for completeness, we report times with and without the response time.
+For completeness, in this study we report times with and without the response time -- defined as travel time.
 
@@
 These crime incidents consume 51% of time on scene throughout the year.
-The remainder of total time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
-The differences in the proportional breakdown between time on scene and total deployment time are marginal.
+The remainder of time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
+The differences in the proportional breakdown between time on scene and total deployment time, which includes response time, are marginal.
+This indicates that the response time plays only a minor role determining the scale and composition of time consumed.
 
@@
 This highlights how mean or aggregate summary statistics of police time as a measure of demand can mask (important) underlying variation.
+
+\newpage
 
@@
 On this point, this study raises useful discussion points around the different measures of officer time.
-First, it is clear that findings on the scale and composition of police demand are sensitive, albeit only slightly, to whether measures of officer time include the response time.
-The degree to which the two estimates differ varies by demand classification, likely as a result of the different call priorities, and in turn, the rapidness of the response times.
-That said, this claim would need verifying through detailed examination.
-On this point, second, we would reiterate that the 'time on scene' measure reported here, in alignment with @lum2021can, refers to the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
-This differs to that of @ratcliffe2021policing which accounted the number of officers attending the scene.
+First, it is clear that findings on the scale and composition of police demand are only marginally sensitive to whether measures of officer time include the response time -- defined as the time from dispatch to when the first officer arrives on-scene.
+The overall findings are almost identical between the two measures.
+Here, we would highlight that our response time measure did not include dispatch time, but rather, travel time in isolation.
+Including dispatch time would increase the gap between the two measures: low-priority calls will be held in the dispatch queue for longer, inflating the overall response time, while high-priority calls will be dispatched immediately.
+Since certain calls will have higher priority responses (e.g., violent crime) this gap is likely to differ between demand classifications.
+We excluded dispatch time on the basis that it does not capture deployment, but a specific investigation into its impact would be straightforward using the open materials provided.
+
+Second, and relatedly, we would reiterate that the 'time on scene' measure reported here, in alignment with @lum2021can, refers to the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
+This differs to that of @ratcliffe2021policing which accounted for the number of officers attending the scene.
 @lum2021can note that including back-up officers in the time calculation might introduces complications.
@@
 We propose that future research investigates these distinctions.
-This would require CAD (or equivalent) data, timestamped at each stage the response (call-handler dispatch, arrival on scene, scene departure), for individual police units nested within calls.
+This would require CAD (or equivalent) data, timestamped at each stage of the response (dispatch, scene arrival, scene departure), for individual police units nested within calls.
 This level of granularity would permit the calculation and comparison of multiple measures of officer time, as well as an investigation into the extent to which back-ups and/or 'flocking' might influence such measures.

### 2022-02-15 paper.Rmd
update paper

@@
 Both @ratcliffe2021policing and @lum2021can were published amidst the backdrop of the killing of George Floyd in May, 2020 and subsequent 'defund the police' protests across the US.
-Evidence which demonstrates that the public rely on the police for social duties beyond law enforcement suggests that a radical reduction in their role and responsibilities (e.g., through defunding) would increase demand on other social services.
+Evidence which demonstrates that the public relies on the police for social duties beyond law enforcement suggests that a radical reduction in their role and responsibilities (e.g., through defunding) would increase demand on other social services.
 Should such radical reform take place without first understanding the scale and composition of public demand for the police, and without preemptively providing appropriate substitutes, there will be no supply to meet this shift in demand.
@@
 Broadly speaking, the results presented here provide robustness to those reported by @ratcliffe2021policing and @lum2021can.
-There is now substantial empirical evidence to demonstrate that, in multiple jurisdictions, the public rely on the police for both social _and_ crime-fighting functions.
+There is now substantial empirical evidence to demonstrate that, in multiple jurisdictions, the public relies on the police for both social _and_ crime-fighting functions.
 A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.

### 2022-02-16 paper.Rmd
removed typo

@@
 This differs to that of @ratcliffe2021policing which accounted for the number of officers attending the scene.
-@lum2021can note that including back-up officers in the time calculation might introduces complications.
+@lum2021can note that including back-up officers in the time calculation might introduce complications.
 For instance, non-serious incidents might receive back-up unnecessarily, particularly during periods of low-demand, which would inflate the time consumed and overstate the resources required to resolve such an incident.

### 2022-02-18 paper.Rmd
inserted 'to' in first sentence
small spelling error
paper complete

@@
 Just 51% of total deployed time responding to 911 calls is consumed by crime incidents.
-The remainder is spent on community (5%), health (7%), proactive (4%), quality of life (16%), and traffic (15%) duties.
+The remainder is spent on quality of life (16%), traffic (15%), health (7%), community (5%) and proactive (4%) duties.
 A small number of calls consume a disproportionately large amount of police officer time.
@@
 It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
-Especially since classifying police demand into these different categories is no straightforward or facile task.
-The difference in categories in Ratcliffe's and Lum's paper attest to this and showcase why they should not have to be taken at face-value.
 Open materials have the additional benefit of being available for re-use in other jurisdictions, helping build a reliable and comparable evidence-base around the scale and composition of police demand across different contexts.
@@
 In this way, the calculation did not account for the additional time consumed by multi-officer responses (e.g., back-up).
-This was justified for a number of reasons, among them, avoiding the over-inflation of time estimates due to officers 'flocking' to non-serious incidents, and limitations with the CAD data.
+This was justified for a number of reasons, among them, avoiding the over-inflation of time estimates due to multiple officers attending non-serious incidents, and limitations with the CAD data.
 A proportional breakdown of the disposition (assist only, report written or arrest citation) was provided for each of the fourteen call types.
@@
 This includes call frequency counts, a proportional breakdown of call counts, and a proportional breakdown of police time.
-Both _total deployed time_ and _time on scene_ measures of police resource are reported.
+Both _total deployed time_ (response time + time on scene) and _time on scene_ measures of police resource are reported.
 The proportional breakdown of _time on scene_ is visualized using a treemap graphic for each incident type nested within the demand classifications.
@@
 A series of descriptive statistics summarize the time on scene consumed by each demand type.
-For comprehensiveness and transparency, we also report the mean, median, standard deviation, and range in the Appendix, next to a histogram of the underlying distribution in minutes.
+For comprehensiveness and transparency we report the mean, median, standard deviation and range, and in the Appendix, a histogram of the underlying distribution.
 This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.
@@
 2% of incidents are excluded from the spatial visualizations due to incomplete geographic information.
-To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year, for each demand type classification.
+To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year for each demand type classification.
 The equivalent raw count maps are available in the Appendix.
@@
 The frequency counts and proportional breakdowns for each demand type classification are reported in Table 1.
-This demonstrates that in Detroit, a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
+This demonstrates that in Detroit, a considerable proportion of calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
 Just 44% of emergency calls for service related to crime incidents.
@@
 The differences in the proportional breakdown between time on scene and total deployment time, which includes response time, are marginal.
-This indicates that the response time plays only a minor role determining the scale and composition of time consumed.
+This indicates that the response time--as measured by travel time--plays only a minor role determining the scale and composition of time consumed.
 
@@
 While many calls are resolved in minutes, a handful consume hours of officer time on the scene, resulting in a highly skewed distribution.
-This highlights how mean or aggregate summary statistics of police time as a measure of demand can mask (important) underlying variation.
+This also highlights how mean or aggregate summary statistics of police time as a measure of demand can mask (important) underlying variation.
 
@@
 Broadly speaking, the results presented here provide robustness to those reported by @ratcliffe2021policing and @lum2021can.
-There is now substantial empirical evidence to demonstrate that, in multiple jurisdictions, the public relies on the police for both social _and_ crime-fighting functions.
+There is substantial empirical evidence to demonstrate that, in multiple jurisdictions, the public relies on the police for both social _and_ crime-fighting functions.
 A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
@@
 
-In this study, calls were categorized according the six classes used by @ratcliffe2021policing.
+In this study, calls were categorized according to the six classes used by @ratcliffe2021policing.
 The overall composition of police demand in terms of call volume and proportional time spent are remarkably similar.
@@
 Here, we would highlight that our response time measure did not include dispatch time, but rather, travel time in isolation.
-Including dispatch time would increase the gap between the two measures: low-priority calls will be held in the dispatch queue for longer, inflating the overall response time, while high-priority calls will be dispatched immediately.
+Including dispatch time would increase the gap between the two measures: low-priority calls will be held in the dispatch queue for longer, inflating the overall response time, while high-priority calls will be dispatched rapidly.
 Since certain calls will have higher priority responses (e.g., violent crime) this gap is likely to differ between demand classifications.
@@
 This level of granularity would permit the calculation and comparison of multiple measures of officer time, as well as an investigation into the extent to which back-ups and/or 'flocking' might influence such measures.
+In sum, further investigations into different measures of 'police time' and the corresponding impact on our assessment of the scale and composition of police demand are necessary before we can claim that the evidence-base is sufficient.
 

### 2024-10-28 results_summary.Rmd
reprod check
Adding a results RMD and minor edits to code to avoid errors during re-run

--- 2022-02-15 paper.Rmd
@@
 ---
-title: 'Describing the scale and composition of calls for police service: a replication and extension using open data'
-author: Samuel Langton\dag, Stijn Ruiter\dag\*, Tim Verlaan\dag
+title: 'Results: Describing the scale and composition of calls for police service: a replication and extension using open data'
+author: Samuel Langton, Stijn Ruiter, Tim Verlaan
 bibliography: refs.bib
@@
   - \usepackage{lineno}
-  - \linenumbers
 csl: apa.csl
-abstract: "This paper describes the scale and composition of emergency demand for police services in Detroit, United States.
-The contribution is made in replication and extension of analyses reported by Ratcliffe (2021) and Lum et al.
-(2021).
-Findings indicate that police spend a considerable proportion of time performing a social service function.
-Just 51% of total deployed time responding to 911 calls is consumed by crime incidents.
-The remainder is spent on community (5%), health (7%), proactive (4%), quality of life (16%), and traffic (15%) duties.
-A small number of calls consume a disproportionately large amount of police officer time.
-Emergency demand is concentrated in time and space, and can differ between types of demand.
-Findings further highlight the potential implications of radically reforming police forces in the United States.
-The data and code used here are openly available for reproduction, re-use and scrutiny."
 ---
 
-\begin{center}
-
-Keywords: police, calls for service, 911, demand
-
-\end{center}
-
-\begin{flushleft}
-
-\dag Netherlands Institute for the Study of Crime and Law Enforcement (NSCR), Amsterdam, Netherlands. \linebreak
-*Department of Sociology, Utrecht University, Utrecht, Netherlands \linebreak \linebreak \linebreak \linebreak
-
-\end{flushleft}
-
-**ORCID** Samuel Langton (0000-0002-1322-1553). Stijn Ruiter (0000-0003-2872-2710). Tim Verlaan (0000-0002-3048-5817).
-
-**Acknowledgments** To be completed.
-
-**Disclosure statement** The authors declare that they have no competing interests.
-
-**Funding details** This research is part of the NSCR research program _What Works in Policing. Towards	Evidence-Based Policing in the Netherlands_, which has been funded by Netherlands Police.
-
-**Availability of data** The data and code used in this study are openly available (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
+**THIS DOC HAS BEEN CREATED PURELY SUMMARIZE THE RESULTS (WITH NO TEXT) IN ONE PLACE. ALL RESULTS PRESENTED HERE ARE GENERATED IN THE R SCRPTS.**
 
 \newpage
+
+# Main results
 
@@
 ```
-
-# Background
-
-Amidst austerity measures, growing public expectation, and scrutiny, understanding the public demand for police services has become a priority among evidence-based policing researchers and practitioners [@boulton2017calls].
-Without a grasp on the scale ('how much?') and composition ('what type?') of police demand, we are likely to observe sub-optimal and inequitable outcomes for the public, the inappropriate distribution of public funds, and unnecessary strain on officers [@ellison2021understanding; @lum2021can].
-Understanding the characteristics of public demand for the police has become particularly pertinent following recent calls to rethink, and in some cases, radically reform, the role and reach of contemporary police forces [@lum2021can].
-
-A recent study from @ratcliffe2021policing sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States (US), during 2019.
-The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside of the traditional crime-fighting role of the police.
-Instead, the origins of public demand for the police can often be attributed to a lack of supply and/or accessibility failure in other organisations, such as those providing support for people requiring (mental) health assistance [@wood2021can; @van2017law].
-Ratcliffe reported that just 40% of public emergency calls for service during the year were crime incidents.
-These calls consumed 55% of total officer shift time.
-The remaining public demand from calls involved community issues (10% of calls, 7% of time), medical/public health incidents (7% of calls, 9% of time), proactive policing (8% of calls, 5% of time), quality of life (24% of calls, 14% of time) and traffic duties (11% of calls and 11% of time).
-Comparable results were subsequently reported across multiple anonymized police jurisdictions in the US [@lum2021can; see also @lumwash in _The Washington Post_].
-
-These descriptive findings provided data-driven evidence to suggest that the police indeed spend considerable amounts of time dealing with incidents in a social rather than solely crime-fighting capacity.
-This has implications to and beyond policing practice.
-Findings directly contribute to the fierce and politically-charged debate around the role and responsibilities of police officers in contemporary society.
-Both @ratcliffe2021policing and @lum2021can were published amidst the backdrop of the killing of George Floyd in May, 2020 and subsequent 'defund the police' protests across the US.
-Evidence which demonstrates that the public relies on the police for social duties beyond law enforcement suggests that a radical reduction in their role and responsibilities (e.g., through defunding) would increase demand on other social services.
-Should such radical reform take place without first understanding the scale and composition of public demand for the police, and without preemptively providing appropriate substitutes, there will be no supply to meet this shift in demand.
-Thus, @ratcliffe2021policing and @lum2021can contributed important empirical data to the (ongoing) discussion on police reform in the US.
-
-Here, we propose an extension of @ratcliffe2021policing and @lum2021can.
-This extension has two principal components.
-First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials.
-To date, investigations have been conducted on data which are not publicly available and without the materials (e.g., code) to reproduce the results.
-Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete and replicable pool of evidence [@openedp2019].
-In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of the debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation.
-It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy.
-Open materials have the additional benefit of being available for re-use in other jurisdictions, helping build a reliable and comparable evidence-base around the scale and composition of police demand across different contexts.
-
-<!-- Second, and relatedly, the materials provided here can be re-used as a tool to conduct further replications for the purposes of comparison and generalization.
-One way in which our materials facilitate this is through an open classification scheme, namely, the recoding of raw call descriptions into meaningful groups (e.g., 'animal incident'), and the subsequent classification into demand types (e.g., 'quality of life').
-As both @ratcliffe2021policing and @lum2021can point out, categorizing call types is challenging and subject to a degree of interpretation.
-But, without a systematic application of such a classification, we cannot compare jurisdictions, make generalizations, or make iterative improvements to previous studies.
-Thus, we encourage practitioners and academics to use these materials to help build a reliable and valid evidence-base around the scale and composition of demand for police services.
--->
-
-Second, we extend the analyses used by @ratcliffe2021policing and @lum2021can.
-This includes a comparison between measures of officer time with and without the response time.
-While concerns have been raised about the inclusion of the response time in measures of officer time, largely due to low-priority calls not necessitating a rapid response, the precise impact of their usage on the same data is unclear.
-We also provide a comprehensive account of the variability in the time consumed by different demand classifications (e.g., crime, community, health).
-These statistics shed light on the degree to which a disproportionately large amount of officer time is consumed by just a small number of calls - insight which can be used to assess how a reduction in demand, or a shift away from calls for police service towards other social services, might free-up officer time.
-In addition, we report on the spatial and temporal patterning of each demand classification.
-Distinctions in these patterns across demand types would be indicative of similarly distinct demand-generating mechanisms (e.g., built environment).
-Such concentrations in time and space might also inform deployment decisions based on officer expertise and experience, and/or collaboration with other services (e.g., mental health professionals).
-
-With these points in mind, this paper describes the emergency demand for police services in a large city, Detroit, US, using open data and open code.
-All materials are publicly available (see https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
-It does so in replication, and where appropriate extension, of the main findings reported by @ratcliffe2021policing and later @lum2021can.
-In this way, the contribution serves to test the robustness of existing findings, provides the materials required for re-use, reproduction and replication, and provides additional insight through new analyses.
-
-# Describing demand
-
-Given the aims of this study, this section first provides a brief outline of the data and methods used by @ratcliffe2021policing and @lum2021can to describe the scale and composition of emergency reactive police demand.
-
-## Ratcliffe (2021)
-
-Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for police service during 2019.
-Calls were subset to exclude those which did not require officer dispatch or did not originate from the public.
-CAD codes, discrete descriptions of incidents based on information available to the dispatcher, were categorized into six broad classifications: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties.
-The classification scheme was used to report on the frequency and proportional breakdown of calls relating to each type of demand throughout 2019.
-
-Recognizing that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by public demand, raw counts were augmented with the time spent by officers on each of the six classifications of demand.
-This information permits the calculation of a proportional breakdown of police time consumed by the different demand classifications throughout the year.
-The raw numbers were reported in a table.
-A visual representation of call types nested within each demand classification was summarized in a tree map graphic.
-The time committed by officers was calculated from when the call was dispatched to an officer to when the officer indicated that the event had been closed.
-Ratcliffe accounted for the number of officers attending each incident when calculating the amount of time consumed.
-The aggregate time consumed by different demand classifications was reported using a single count of officer shifts (7.5 hours) over the year.
-
-The spatio-temporal patterning of this demand was summarized in two visuals, focusing specifically on those calls relating to medical/public health.
-A heatmap showed the mean call volume during hours of the day and days of the week, identifying peak times for demand.
-A geographic map visualized the spatial patterning of these call count concentrations using kernel density estimation in order to identify hotspots of police demand involving medical/public health issues.
-To supplement these findings, Ratcliffe also reported on the differences between the initial CAD incident type classification and the final disposition using a Sankey diagram -- again focusing specifically on medical and public health calls.
-
-## Lum et al (2021)
-
-To describe the scale and composition of emergency police demand across the US more broadly, @lum2021can used CAD data from nine different police agencies.
-The agencies themselves were anonymized.
-Individual incidents were categorized into fourteen different demand classifications, namely, alarms, disorders, vice, domestic, follow-ups and service requests, mental, medical, missing persons, violence, interpersonal-other, property, suspicions, traffic, and admin, agency and non-crime.
-The authors reported a proportional breakdown of call counts and police time spent across the nine agencies according to the fourteen demand classifications, including the average (mean).
-For each agency, the authors also reported the average time spent on each demand classification in minutes, and the average across the nine agencies.
-No descriptive statistics on the range or spread of these figures was reported.
-
-The 'time spent' measure used was the _time on scene_ (i.e., excluding the response time).
-The authors considered this to be preferable, since low priority calls may have a non-urgent (slow) response due to the nature of the call, and officers may be reassigned to other duties while on an assignment.
-Lum and colleagues calculated _time on scene_ as the time from the first officer's arrival to when the call was reported as resolved.
-In this way, the calculation did not account for the additional time consumed by multi-officer responses (e.g., back-up).
-This was justified for a number of reasons, among them, avoiding the over-inflation of time estimates due to officers 'flocking' to non-serious incidents, and limitations with the CAD data.
-A proportional breakdown of the disposition (assist only, report written or arrest citation) was provided for each of the fourteen call types.
-No spatial or temporal findings were reported.
-
-# Data
-
-## Detroit calls for service
-
-The data used for this study cover the city of Detroit in Michigan, US.
-The City of Detroit authority publish 911 emergency calls for service data through their Open Data Portal.^[Accessible via https://data.detroitmi.gov/.] In alignment with the two existing studies, the data used here exclude those calls initiated by an officer in order to capture public-initiated calls.
-Each individual incident has a corresponding timestamp and location coordinates to define the call in time and space.
-Data was subset for the year 2019.
-
-For each incident, the police response time and time on the scene are reported.
-The former measures the time from dispatch to when the first officer arrives on-scene.^[In the raw data, the time between dispatch and first on-scene arrival is referred to as _travel time_.
-The _intake_ and _dispatch time_ are reported, however, we did not consider these measures to reflect deployment time.
-We return to this point in the discussion.] The latter mimics the measure used by @lum2021can, namely, the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
-Here, the data permit a distinction between the time spent on calls with and without the response time.
-We refer to the sum of the response time and time on scene as the _total deployed time_.
-Calls which had a response time of negative values or missing were dropped from analyses.
-Calls which had a _time on scene_ of zero minutes, negative values or missing were also excluded.
-As noted, @lum2021can considered the _time on scene_ measure to be preferable.
-For completeness, in this study we report times with and without the response time -- defined as travel time.
-
-Each incident has a _calldescription_ variable which describes the nature of the call.
-Call descriptions involving administrative duties (e.g., "start of shift information"), completely unknown problems, and non-deployment (e.g., "employee call in / time off") were removed.
-This left 207 unique call description categories.
-In the interests of parsimony and ease of interpretation these categories were recoded into 99 broader call descriptions.^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (https://anonymous.4open.science/r/demand_detroit-FCD3/README.md).
-Readers unfamiliar with R code can review the reference table entitled 'categorization summary'.] This included the removal of incidents involving transporting prisoners and executing warrants, which were deemed to have been erroneously flagged as not being officer-initiated.
-Each call description was then categorized into the six demand type classifications used in @ratcliffe2021policing, namely, crime, community, medical/public health, proactive, quality of life and traffic.
-2.8% of calls were deemed unclassifiable.
-For each of the six demand classifications an 'other' incident type was generated for those incidents which consumed less than 0.2% of total deployed time during the year.
-This left a total of 49 unique categories for the reported breakdowns, capturing ~260,000 individual public calls for service in Detroit during 2019.
-No information on the final disposition of calls is provided in the open data.
-
-# Methods
-
-The methods used largely replicate those used by @ratcliffe2021policing and @lum2021can, and where appropriate, extensions.
-First, descriptive statistics and a treemap graphic provide an account of the scale and composition of police demand during the year.
-Second, visualizations summarize the spatial and temporal patterning of this demand.
-
-## Descriptive statistics
-
-Summary statistics describe the overall composition of demand according to the six demand type classifications.
-This includes call frequency counts, a proportional breakdown of call counts, and a proportional breakdown of police time.
-Both _total deployed time_ and _time on scene_ measures of police resource are reported.
-The proportional breakdown of _time on scene_ is visualized using a treemap graphic for each incident type nested within the demand classifications.
-The equivalent graphic for _total deployed time_ is available in the corresponding GitHub repository.
-
-A series of descriptive statistics summarize the time on scene consumed by each demand type.
-For comprehensiveness and transparency, we also report the mean, median, standard deviation, and range in the Appendix, next to a histogram of the underlying distribution in minutes.
-This sheds on light on how long a 'typical' call might take while noting a probable skew in the distribution.
-Such a skew would suggest that averages of time [e.g., @lum2021can] or aggregate measures of time [e.g., @ratcliffe2021policing] in isolation are insufficient and potentially mask important underlying variation, namely, that a small number of high-demand calls drag up the mean and inflate aggregate measures of time.
-To investigate this distribution, we plot Lorenz curves for each demand classification.
-Lorenz curves plot the cumulative distribution of a variable (in this case, time on scene) against the cumulative distribution of a unit (in this case, calls) [@bernasco2017more].
-By way of example, we report on these figures according to arbitrary thresholds of 25% and 50% of time spent.
-This sheds light on the degree to which a disproportionately large amount of officer time might be consumed by just a small number of calls.
-
-<!-- As an aggregate measure, we report the sum time spent on each demand classification by the number of officer shifts.^[Consistent with Ratcliffe (2021) a shift is defined as 7.5 hours.
-The sum total of hours for each demand classification over the year is divided by 7.5.]  -->
-
-## Spatio-temporal patterning
-
-The temporal patterning of the six demand type classifications are visualized using a heatmap of mean call counts, aggregated by day of the week, and hour of the day.
-To summarize the spatial patterning, call location coordinates are aggregated to synthetic grid cells using spatial joins.
-Grid cells are defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localized variation and interpretable visualizations.
-2% of incidents are excluded from the spatial visualizations due to incomplete geographic information.
-To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year, for each demand type classification.
-The equivalent raw count maps are available in the Appendix.
-In this way, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, but here, we provide a comparison across different demand classifications and two different measures of demand.
-
-<!-- ## Data and methods summary -->
-
-<!-- To summarize the various descriptions of police demand replicated and/or extended in this study, a checklist is outlined in Table 1.
-Broadly, this demonstrates that this study provides a comprehensive replication of the analyses reported by @ratcliffe2021policing and @lum2021can.
-Either identical descriptive statistics are reported in replication of one or both existing studies, or additional analyses are provided for enhanced insight and transparency.
-Here, we note two exceptions: we do not report an aggregate summary (total shifts) because our time measure does not  -->
-
-<!-- \newpage -->
-
-<!-- ```{r table1} -->
-<!-- options(knitr.kable.NA = '') -->
-
-<!-- table1_df <- read_xlsx("results/table1_methods_comparison.xlsx") -->
-
-<!-- kable(table1_df, caption = "Summary of the various descriptions of demand replicated or extended in this study. Measures were reported for demand classification unless otherwise stated.", -->
-<!--       booktabs = T, longtable = T, linesep = "", format = "latex") %>% -->
-<!--     kable_styling(latex_options = c("striped"), -->
-<!--                   font_size = 10, full_width = F) %>% -->
-<!--   column_spec(column = c(2,3,4), width = c("2cm", "2cm", "2cm") ) %>% -->
-<!--   row_spec(row = 14, hline_after = 14) %>% -->
-<!--   row_spec(row = 15, italic = T) -->
-<!-- ``` -->
-
-# Results
-
-## Scale and composition
-
-The frequency counts and proportional breakdowns for each demand type classification are reported in Table 1.
-This demonstrates that in Detroit, a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of (often non-criminal) issues.
-Just 44% of emergency calls for service related to crime incidents.
-These crime incidents consume 51% of time on scene throughout the year.
-The remainder of time on scene is spent dealing with traffic duties (16%), quality of life incidents (15%), public and mental health (7%), community issues (5%), and proactive policing (4%).
-The differences in the proportional breakdown between time on scene and total deployment time, which includes response time, are marginal.
-This indicates that the response time plays only a minor role determining the scale and composition of time consumed.
-
-<!-- These crime incidents consumed 51% of total deployed time throughout the year, although the figure increases to 51% when considering time on scene in isolation.
-This likely reflects the fact that crime-related calls have higher priority, and in turn, a swifter response time.^[Crime calls for service had the highest proportion of Priority 1 (highest priority) responses at 52%, followed by traffic (24%), quality of life (17%), community (16%), health (14%), and proactive (3%).] -->
 
@@
 
-The proportional breakdown of time spent on scene for each incident type, grouped by the six demand classifications, is visualized in Figure 1.^[The purpose of the graphic is to provide a broad summary, in replication of Ratcliffe's visual.
-The numbers underlying the graphic are reported in the corresponding GitHub repository.] This figure demonstrates the diversity and complexity of public demand for police services in Detroit.
-Time consumed on the scene of crime incidents is largely comprised of domestic incidents, burglary and assault.
-For traffic incidents, police time was mostly spent attending to accidents, while for quality of life, attending to disturbances consumed the most time on scene.
-Calls relating to public health involved mental health incidents, overdoses, (threats of) suicide and well-being checks.
-Proactive incidents largely involved investigations, while time consumed on community issues consisted mostly of missing/found persons and animal incidents (e.g., aggressive dogs).
+\newpage
 
-## Time distributions
-
-To further quantify the resources consumed, Table 2 summarizes the time spent on the scene for each demand classification in minutes.
-Histograms of the underlying distributions are reported in the Appendix.
-Table 2 demonstrates a number of things.
-First, there is a modest amount of variation in the typical time at scene of different types of emergency call.
-For instance, the mean time spent on the scene of traffic incidents (55 minutes) and crime (51 minutes) is considerably more that quality of life issues (31 minutes) or proactive activity (28 minutes).
-These differences will be a result of a combination of factors, including the complexity and severity of incidents.
-Second, there is considerable variation in the time consumed by incidents _within_ each demand type.
-While many calls are resolved in minutes, a handful consume hours of officer time on the scene, resulting in a highly skewed distribution.
-This highlights how mean or aggregate summary statistics of police time as a measure of demand can mask (important) underlying variation.
+```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of police time consumed over the year, defined as time spent on the scene, for each call category.", out.width = "\\textwidth"}
+knitr::include_graphics("visuals/fig1_time_tos.png")
+```
 
@@
 
-\newpage
-
-```{r fig1, fig.align = "center", fig.cap = "Proportional breakdown of police time consumed over the year, defined as time spent on the scene, for each call category.", out.width = "\\textwidth"}
-knitr::include_graphics("visuals/fig1_time_tos.png")
-```
-
-\newpage
-
-Figure 2 visualizes the disproportionate share of time spent on a small number of calls explicitly using Lorenz curves for each demand classification.
-These can be compared against the reference line of perfect equality.
-Here, the line of perfect equality represents a situation in which officer time is distributed evenly across all calls.
-In reality, we can see that this is not the case: the Lorenz curves for each demand classification deviate considerably from perfect equality.
-The curves for each classification are almost indistinguishable, so we focus on the curve for all calls (in blue).
-This shows that 50% of the time officers spent on the scene during 2019 was consumed by just 15% of calls, and 25% of time on scene was consumed by just 5% of all calls.
-In other words, a small amount of emergency calls for service account for a disproportionately large amount of police resources out on the street.
-This finding is consistent (and comparable) across all six classifications of demand.
-
 ```{r fig2, fig.align="center", fig.cap = "Lorenz curve for time spent on scene across all calls (in blue). Lorenz curves in grey for each of the six demand classifications."}
@@
 ```
-
-## Spatio-temporal patterning
-
-Mean incident counts by day of the week and by hour of day, for each demand classification, are visualized in Figure 3.
-This demonstrates that the different types of public demand for police services, as indicated by the six demand classifications, can have a distinct temporal patterning.
-Community calls appear to concentrate on weekdays, around lunchtime, when other classes of demand are low.
-Crime and quality of life-related calls both appear to occur in the evening and early hours of the morning.
-Traffic incidents concentrate in the evenings, particularly on a Friday.
-Proactive policing does not have a distinct hourly patterning, but this could be attributed to the low hourly counts.
-Calls relating to health issues, the primary interest of @ratcliffe2021policing, appear to largely take place in the evening and early morning.
-
-The spatial patterning of police demand in Detroit, defined as the aggregate time spent on scene throughout the year, is visualized for each demand classification in Figure 4.^[Noting that the aggregate time spent on scene may be influenced by outliers, the raw count maps and underlying distributions are reported in the Appendix.
-These indicate that 'high time' outliers do not seriously misrepresent the aggregate figures in Figure 4.] This demonstrates that the time police spent attending to incidents is highly concentrated in space, and can vary between different demand types.
-The police spend a considerable amount of time around the Wayne State University (WSU) and Midtown attending to community, proactive and quality of life calls for service.
-Officers spend vast amounts of time on crime-related calls in or near hospitals, most notably, the Henry Ford Hospital.
-Police time spent attending to health-related calls concentrates in and around WSU and Midtown, but also among grid cells containing mental health and well-being facilities outside of the city center.
-Time on traffic incidents have a distinct spatial patterning along arterial roads into Downtown.
 
@@
 \newpage
-
-# Discussion
-
-In replication and extension of the main findings reported by @ratcliffe2021policing and @lum2021can this contribution has described the scale and composition of calls for police service in Detroit, US.
-Broadly speaking, the results presented here provide robustness to those reported by @ratcliffe2021policing and @lum2021can.
-There is now substantial empirical evidence to demonstrate that, in multiple jurisdictions, the public relies on the police for both social _and_ crime-fighting functions.
-A reduction in the capacity of the police to meet this emergency demand, without first providing appropriate, capable and accessible replacement emergency services, could have considerable implications for public safety and well-being.
-This contribution provides the first such demonstration using open data and open code, opening up analyses for scrutiny, re-use, and extension by the public, academics and/or police practitioners.
-We encourage usage of the materials provided with this study to replicate analyses across other jurisdictions.
-In doing so, we hope to see a consistent and reliable evidence-base across different contexts for the purposes of comparison and generalization.
-
-In this study, calls were categorized according the six classes used by @ratcliffe2021policing.
-The overall composition of police demand in terms of call volume and proportional time spent are remarkably similar.
-A considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police.
-Instead, the police fulfill a variety of other social service functions, including the resolution of public (mental) health incidents and resolving community issues.
-That said, descriptive statistics which drill down into the underlying distribution of officer time consumed by each demand type provide additional insight which would otherwise be overlooked.
-First, it is clear that there is considerable variation in the amount of time consumed both between and within demand classifications.
-On a methodological point, this highlights how average (e.g., mean times) or aggregate (e.g., sum times) might mask underlying variation.
-Second, and relatedly, these extensions demonstrate that a disproportionately large amount of police time is spent on a small number of emergency calls.
-50% of total time on scene in consumed by just 15% of calls.
-High-demand calls will likely be complex and/or life-threatening incidents or calls which require officers to secure a scene for extended periods of time.
-A targeted examination of these calls might shed light on the extent to which such high-intensity demands could be reduced or better resolved through or in concert with other public services.
-
-On this point, this study raises useful discussion points around the different measures of officer time.
-First, it is clear that findings on the scale and composition of police demand are only marginally sensitive to whether measures of officer time include the response time -- defined as the time from dispatch to when the first officer arrives on-scene.
-The overall findings are almost identical between the two measures.
-Here, we would highlight that our response time measure did not include dispatch time, but rather, travel time in isolation.
-Including dispatch time would increase the gap between the two measures: low-priority calls will be held in the dispatch queue for longer, inflating the overall response time, while high-priority calls will be dispatched immediately.
-Since certain calls will have higher priority responses (e.g., violent crime) this gap is likely to differ between demand classifications.
-We excluded dispatch time on the basis that it does not capture deployment, but a specific investigation into its impact would be straightforward using the open materials provided.
-
-Second, and relatedly, we would reiterate that the 'time on scene' measure reported here, in alignment with @lum2021can, refers to the amount of time which elapses between when the first officer arrives and when the last unit clears the scene.
-This differs to that of @ratcliffe2021policing which accounted for the number of officers attending the scene.
-@lum2021can note that including back-up officers in the time calculation might introduces complications.
-For instance, non-serious incidents might receive back-up unnecessarily, particularly during periods of low-demand, which would inflate the time consumed and overstate the resources required to resolve such an incident.
-Although the overall composition of police demand is comparable between Philadelphia [@ratcliffe2021policing] and Detroit, the true impact of these measures remains unclear.
-We propose that future research investigates these distinctions.
-This would require CAD (or equivalent) data, timestamped at each stage of the response (dispatch, scene arrival, scene departure), for individual police units nested within calls.
-This level of granularity would permit the calculation and comparison of multiple measures of officer time, as well as an investigation into the extent to which back-ups and/or 'flocking' might influence such measures.
-
-The spatial patterning of demand for police services in Detroit--as measured through the aggregate time spent on scene and raw counts--demonstrates that different demands often originate from different places.
-Given the specialized training and experience required to deal with certain forms of police demand (e.g., well being checks, suicide threats), forces might consider accounting for this variation when assigning officers to shifts and patrol areas, or organizing co-patrols with other social service professionals.
-This might lead to more favorable outcomes for the public and less strain on officers.
-The spatio-temporal findings also suggests that the underlying environmental factors generating different forms of police demand are (or can be) distinct.
-Recent research has found that the explanatory frameworks underpinning environmental criminology (e.g., place attractors) are relevant for non-criminal forms of demand, such as those calls involving emotionally disturbed persons [@vaughan2016exploring].
-That said, we know little about the nuances of how different environmental characteristics might drive different forms of demand in the same study area.
-The study materials made available here could be extended using open data resources, such as Open Street Map, to provide a theoretically-driven investigation into the urban character (e.g., place attractors) of high-demand areas according to the six demand classifications.
-Based on the distinct temporal patterning of different demand types, such analyses should be time-sensitive, capturing, for instance, the ambient population flows of different places throughout the day.
-To date, ambient population analyses have focused on crime-specific calls for service [@andresen2007hot] which as we have seen, constitute only part of the emergency demands placed on police.
-
-The study is not without shortcomings.
-First, while the Detroit data permits an open and reproducible case site, the raw incident data do not make a distinction between the initial call classification and the final disposition.
-This would permit a comparison, as conducted by @ratcliffe2021policing, into the extent to which the call handler's initial incident classification changes following the attendance of officers at the scene.
-Here, we cannot conduct such an investigation.
-Second, findings presented here, and those reported by @ratcliffe2021policing and @lum2021can, rely on the reliability and validity of police incident classifications.
-These are imperfect.
-Evidence from the United Kingdom, for instance, indicates that police under-record the proportion of emergency calls involving persons with mental ill-health, which in turn underestimates the amount of police time consumed by such incidents [@langton2021policing].
-Investigations into recording practices and call handler decision-making will be necessary to further understand this issue across US police jurisdictions [see @lum2020constrained; @simpson2021calling].
 

### 2024-10-28 scripts/calls_detroit.R

@@ -292,7 +292,7 @@ unclas_vec    <- c("ADMIT OR E/E",
 # Add broad category.
 # We now also remove prisoner and warrant calls, as these are deemed to officer-initiated.
 detroit_19_times_df <- detroit_19_times_df %>% 
-  mutate(type = if_else(calldescription2 %in% comm_vec     , "community"      , NULL),
+  mutate(type = if_else(calldescription2 %in% comm_vec     , "community"      , NA),
          type = if_else(calldescription2 %in% crime_vec    , "crime"          , type),
          type = if_else(calldescription2 %in% health_vec   , "health"         , type),
          type = if_else(calldescription2 %in% proactive_vec, "proactive"      , type),
@@ -602,7 +602,7 @@ grid_maps_list <- lapply(detroit19_grid_list, function(x){
                           limits = c(0,max(x$call_count))) +
     labs(fill = NULL) +
     guides(fill = guide_colourbar(barwidth = 9, barheight = 0.6, draw.ulim = FALSE,
-                                  ticks.colour = "black", ticks.linewidth = 2)) +
+                                  ticks.colour = "black", ticks.linewidth = 0.5)) +
     theme_void() +
     theme(legend.text = element_text(size = 11),
           legend.position = c(0.7,0.24),

### 2024-10-28 scripts/calls_detroit_time_on_scene.R

@@ -9,7 +9,8 @@ library(purrr)
 library(forcats)
 library(stringr)
 library(lubridate)
-library(lorenzgini) # requires install from github.
+# remotes::install_github("wsteenbeek/lorenzgini") 
+library(lorenzgini) 
 library(treemapify)
 library(ggplot2)
 library(sf)
@@ -294,7 +295,7 @@ unclas_vec    <- c("ADMIT OR E/E",
 # Add broad category.
 # We now also remove prisoner and warrant calls, as these are deemed to officer-initiated.
 detroit_19_times_df <- detroit_19_times_df %>% 
-  mutate(type = if_else(calldescription2 %in% comm_vec     , "community"      , NULL),
+  mutate(type = if_else(calldescription2 %in% comm_vec     , "community"      , NA),
          type = if_else(calldescription2 %in% crime_vec    , "crime"          , type),
          type = if_else(calldescription2 %in% health_vec   , "health"         , type),
          type = if_else(calldescription2 %in% proactive_vec, "proactive"      , type),
@@ -808,7 +809,7 @@ grid_maps_list <- lapply(detroit19_grid_list, function(x){
     scale_fill_viridis_c(guide = "colourbar", n.breaks = 3, alpha = 0.9) +
     labs(fill = NULL) +
     guides(fill = guide_colourbar(barwidth = 9, barheight = 0.6, draw.ulim = FALSE,
-                                  ticks.colour = "black", ticks.linewidth = 2)) +
+                                  ticks.colour = "black", ticks.linewidth = 0.5)) +
     theme_void() +
     theme(legend.text = element_text(size = 11),
           legend.position = c(0.7,0.24),
@@ -908,4 +909,4 @@ maps_gg <-  plot_grid(plotlist = grid_maps_list,
 ggsave(filename = "visuals/fig4_maps_tos.png", height = 48, width = 40, unit = "cm", dpi = 300)
 
 # Save workspace.
-save.image(file = "data/workspace_complete.RData")
+# save.image(file = "data/workspace_complete.RData")

### 2024-10-28 scripts/calls_detroit_time_on_scene.R
renv snapshot
renv snapshot and include commented install options for lorenzgini

@@ -9,7 +9,8 @@ library(purrr)
 library(forcats)
 library(stringr)
 library(lubridate)
-# remotes::install_github("wsteenbeek/lorenzgini") 
+# Standard install: remotes::install_github("wsteenbeek/lorenzgini") 
+# Renv install: renv::install("wsteenbeek/lorenzgini")
 library(lorenzgini) 
 library(treemapify)
 library(ggplot2)

### 2025-03-18 results_summary.Rmd
repro run

@@
 author: Samuel Langton, Stijn Ruiter, Tim Verlaan
-bibliography: refs.bib
 output:
@@
   - \usepackage{lineno}
-csl: apa.csl
 ---

### 2025-03-18 scripts/calls_detroit_time_on_scene.R

@@ -9,7 +9,7 @@ library(purrr)
 library(forcats)
 library(stringr)
 library(lubridate)
-# Standard install: remotes::install_github("wsteenbeek/lorenzgini") 
+# Standard install: remotes::install_github("wsteenbeek/lorenzgini")
 # Renv install: renv::install("wsteenbeek/lorenzgini")
 library(lorenzgini) 
 library(treemapify)
